### 2020-11-03 manuscript_working/ms_working_RMD.Rmd
initialized methods/results writeup

--- 
title: "working owls methods"
author: "Althea Archer"
date: "11/3/2020"
output:
  word_document:
    fig_caption: yes
  pdf_document:
    includes:
      latex_engine: pdflatex
    keep_tex: yes
csl: ecology.csl
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Abstract text would go here.

# Introduction

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can **Knit** to PDF or to Microsoft word, just chose the appropriate dropdown from the **Knit** button.

# Methods

You can make words **bold** with two asterisks, *italicized* with one.

## Data

Hashtags denote a header, and the hierarchy is defined by the number of hashtags.

## Statistical Analysis

To make a floating equation, use two dollar signs on either side:

$$
\text{logit}(g_l) = \beta_{0g} + \beta_{1g}x_l.
$$

To make an in-line equation, use one dollar sign. As in this: $\alpha_\gamma^2$ for an example.

To make an equation array that shows up in both Word and PDF, you can use the following format:

\[
\begin{aligned}
R  &=  \text{the number of sightability trials (here }R = 124)\\
z_l &= \text{a random variable equal to 1 when the }l\text{th group is detected and 0 otherwise } (l = 1,2, \ldots, R)\\
x_l &= \text{the percent visual obstruction (from 0 to 1) associated with the }l\text{th group } (l = 1,2, \ldots, R)\\
g_l &= \text{the probability of detection of the }l\text{th group } (l = 1,2, \ldots, R)
\end{aligned}
\]

Remember to put ampersands (&) to align equations.

# Results

## Results sub-section

# Discussion

You can cite references using the **citr** add-in. With **citr** you can cite papers in line like @Fieberg:2008 is fond of doing, or parenthetically [@R:2014]. You can also cite multiple sources at once [@Buckland:2015; @Fieberg:2008; @Plummer:2013]. References can be added into the myrefs.bib file directly, or created with BibTeX through LaTeX.

# Figures

Figures can be included with the following code, but they must be in .png format to show up in both Word and PDF. They can be referenced for the PDF with Fig. \ref{figure1}. *Unfortunately, I don't yet know how to make the reference number show up in Word.*

![1. Example figure caption, which can include in-line equations like $y = \alpha + \beta x$ and references like @Fieberg:2008. \label{figure1}](ms_figures/figure1.png)

# References

### 2020-11-03 manuscript_working/ms_working_RMD.Rmd
drafting writeup of results/methods occupancy

@@
 csl: ecology.csl
-bibliography: references.bib
+bibliography: owl_references.bib
 ---
@@
 
-# Abstract
-
-Abstract text would go here.
-
-# Introduction
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can **Knit** to PDF or to Microsoft word, just chose the appropriate dropdown from the **Knit** button.
-
 # Methods
 
-You can make words **bold** with two asterisks, *italicized* with one.
+## Occupancy Model Framework
 
-## Data
-
-Hashtags denote a header, and the hierarchy is defined by the number of hashtags.
-
-## Statistical Analysis
-
-To make a floating equation, use two dollar signs on either side:
-
-$$
-\text{logit}(g_l) = \beta_{0g} + \beta_{1g}x_l.
-$$
-
-To make an in-line equation, use one dollar sign. As in this: $\alpha_\gamma^2$ for an example.
-
-To make an equation array that shows up in both Word and PDF, you can use the following format:
+Let
 
@@
 \begin{aligned}
-R  &=  \text{the number of sightability trials (here }R = 124)\\
-z_l &= \text{a random variable equal to 1 when the }l\text{th group is detected and 0 otherwise } (l = 1,2, \ldots, R)\\
-x_l &= \text{the percent visual obstruction (from 0 to 1) associated with the }l\text{th group } (l = 1,2, \ldots, R)\\
-g_l &= \text{the probability of detection of the }l\text{th group } (l = 1,2, \ldots, R)
+t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\\
+h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\\
+i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
+j  &=  \text{the individual stations along each route (here }j = 1,2,\ldots,10)\\
+k  &=  \text{the broadcast species (here }k = 1,2,\ldots,10)\\
+z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
+y_{t,h,i,j,k} &= \text{a random variable equal to 1 when an owl was detected and 0 otherwise}
 \end{aligned}
@@
 
-Remember to put ampersands (&) to align equations.
+We assumed that survey occupancy ($z_{t,h,i}$) was the outcome of Bernoulli trials with probability $\psi$, which we allowed to vary by year and route:
+
+$$
+z_{t,h,i} \sim \text{Bernoulli}(\psi_{t,h})
+$$
+
+$$
+\psi_{t,h} \sim \text{Beta}(a^\psi_{t,h},b^\psi_{t,h})
+$$
+with hyperpriors for $a^\psi_{t,h}$ and $b^\psi_{t,h}$ such that the expected value of $\psi_{t,h} = \mu^\psi_{t,h}$:
+
+\[
+\begin{aligned}
+\text{E}[\psi_{t,h}] &= \mu^\psi_{t,h} = a^\psi_{t,h}/(a^\psi_{t,h}+b^\psi_{t,h})\\
+\rho^\psi_{t,h} &= a^\psi_{t,h}+b^\psi_{t,h}\\
+\mu^\psi_{t,h} &\sim \text{Uniform}(0.01, 0.99)\\
+\rho^\psi_{t,h} &\sim \text{N}(5,1)\text{, truncated to 0.01,10}\\
+\end{aligned}
+\]
+
+We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the broadcast species:
+
+$$
+y_{t,h,i,j,k} \sim \text{Bernoulli}(z_{t,h,i}*p_{t,h,i,j,k})
+$$
+
+where generally $\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k$, or more specifically:
+
+\[
+\begin{aligned}
+\text{logit}(p_{t,h,i,j,k}) &= \beta_\text{Pre-broadcast}X_\text{Pre-broadcast} + \\
+&= \beta_\text{Mottled}X_\text{Mottled}+\\
+&= \beta_\text{Pacific}X_\text{Pacific}+\\
+&= \beta_\text{Crested}X_\text{Crested}+\\
+&= \beta_\text{Black and White}X_\text{Black and White}+\\
+&= \beta_\text{Spectacled}X_\text{Spectacled}+\\
+&= \beta_\text{Whiskered}X_\text{Whiskered}+\\
+&= \beta_\text{Guat Barred}X_\text{Guat Barred}+\\
+&= \beta_\text{Stygian}X_\text{Stygian}+\\
+&= \beta_\text{Great Horned}X_\text{Great Horned}\\
+\end{aligned}
+\]
+
+This model provides a consistent probability of detection for all surveys in the first two minutes of each survey before the broadcast owl recording was played.
+Then, the probability of detection for each post-broadcast time period would depend on the species of broadcast owl.
+This allowed species-specific behavior in response to the different broadcast owl species [@Baumgardt:2019].
+We used means parameterization such that the coefficients were interpretable as the effect of that specific broadcast species $k$ (including the pre-broadcast time frame).
+The priors for every logistic model coefficient $\beta_k$ were chosen to be near uniform as recommended in @Gelman:2008.
+Specifically, we used the Cauchy prior as such: $\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)$.
+
+The broadcast species were consistent at each station across years, but varied by route.
+For stations $j = 1, 2, \ldots, 10$ at routes EI1, EI2, M2, N1, and N2, the following broadcast species were used, respectively: Pacific, Mottled, Crested, Black and White, Spectacled, Mottled, Crested, Black and White, and Spectacled.
+At the route M1, the following species were called at stations $j = 1, 2, \ldots, 10$, respectively: Whiskered, Mottled, Gaut Barred, Stygian, Great Horned, Whiskered, Mottled, Gaut Barred, Stygian, and Great Horned.
+Thus, the effective sample sizes of the different broadcast species varied.
+
+### Occupany Model Implementation
+
+We used the R2jags package in R [@R:2014,@Plummer:2013] to implement the occupancy model for three owl species: Mottled, Spectacled, and FerPy owls.
+These owls had enough positive detections to analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137, and 187 positive detections over the 11 year period, respectively.
+Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed that route from those occupancy models.
+
+For all three species' occupancy models, we ran 3 chains for 10,000 iterations and 1000 iterations discarded as burn-in, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+We visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
 
@@
 
-## Results sub-section
+## Occupany Model Results
 
-# Discussion
+Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for FerPy or Spectacled owls in both routes of El Imposible and in M2 route of Montecristo (Fig.
+1X).
+The probability of occupancy across all three species and both routes of Nancuchiname were relatively high (Fig.
+1XX).
+We found that the probability of occupancy for Mottled owls was relatively consistent across time in El Imposible and in N2 route (Fig.
+2XX), but that it seemed to increase in the last 6 or so years of the surveys in M2 and N1 routes (Fig.
+2XX).
+The probability of occupancy in M1 seemed to have decreased through time for Mottled owls (Fig.
+2XX).
 
-You can cite references using the **citr** add-in.
-With **citr** you can cite papers in line like @Fieberg:2008 is fond of doing, or parenthetically [@R:2014].
-You can also cite multiple sources at once [@Buckland:2015; @Fieberg:2008; @Plummer:2013].
-References can be added into the myrefs.bib file directly, or created with BibTeX through LaTeX.
+The probability of occupancy for FerPy and Spectacled stayed relatively consistently low across time in routes EI1, EI2, and M2 other than a higher probability of occupancy for FerPy in EI1 in the early years of the surveys (Figs.
+3XX and 4XX).
+FerPy probability of occupancy was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
+3XX), and the probability of occupany for the Spectacled owl slightly fluctuated year to year in the Nancuchaname routes (Fig.
+4XX).
+
+The probability of detecting these three owls varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast species were called varied by broadcast species and between Mottled, FerPy, and Spectacled owls (Fig.
+5XX).
+The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owls.
+The probability of detecting FerPy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owls (Fig.
+5XX).
 
@@
 
-Figures can be included with the following code, but they must be in .png format to show up in both Word and PDF.
-They can be referenced for the PDF with Fig.
-\ref{figure1}.
-*Unfortunately, I don't yet know how to make the reference number show up in Word.*
+![1XX.
+Probability of occupancy ($\psi_{t,h}$) averaged by route $h$ for Mottled, FerPy, and Spectacled Owls.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median $\pm$ 90% credible intervals.
+\label{psi_means}](../output/figures/psi_means-1.png)
 
-![1. Example figure caption, which can include in-line equations like $y = \alpha + \beta x$ and references like @Fieberg:2008. \label{figure1}](ms_figures/figure1.png)
+![2XX.
+Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for Mottled Owl.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median $\pm$ 90% credible intervals.
+\label{figure1}](../output/figures/mottd_psi_byYr-1.png)
+
+![3XX.
+Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for FerPy Owl.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median $\pm$ 90% credible intervals.
+\label{figure1}](../output/figures/ferpy_psi_byYr-1.png)
+
+![4XX.
+Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for Spectacled Owl.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median $\pm$ 90% credible intervals.
+\label{figure1}](../output/figures/specd_psi_byYr-1.png)
+
+![5XX.
+The logistic regression parameters that related the probability of detecting owls ($\p_{t,h,i,j,k}$) to the broadcast timeframe (pre-broadcast) or speccies.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median $\pm$ 90% credible intervals.
+\label{figure1}](../output/figures/p_detection-1.png)
 

### 2020-11-03 manuscript_working/ms_working_RMD.tex
drafting writeup of results/methods occupancy

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={working owls methods},
            pdfauthor={Althea Archer},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{working owls methods}
\author{Althea Archer}
\date{11/3/2020}

\begin{document}
\maketitle

\section{Methods}\label{methods}

\subsection{Occupancy Model}\label{occupancy-model}

Let

\[
\begin{aligned}
t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\\
h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\\
i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
j  &=  \text{the individual stations along each route (here }j = 1,2,\ldots,10)\\
k  &=  \text{the broadcast species (here }k = 1,2,\ldots,10)\\
z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
y_{t,h,i,j,k} &= \text{a random variable equal to 1 when an owl was detected and 0 otherwise}
\end{aligned}
\]

We assumed that survey occupancy (\(z_{t,h,i}\)) was the outcome of
Bernoulli trials with probability \(\psi\), which we allowed to vary by
year and route:

\[
z_{t,h,i} \sim \text{Bernoulli}(\psi_{t,h})
\]

\[
\psi_{t,h} \sim \text{Beta}(a^\psi_{t,h},b^\psi_{t,h})
\] with hyperpriors for \(a^\psi_{t,h}\) and \(b^\psi_{t,h}\) such that
the expected value of \(\psi_{t,h} = \mu^\psi_{t,h}\):

\[
\begin{aligned}
\text{E}[\psi_{t,h}] &= \mu^\psi_{t,h} = a^\psi_{t,h}/(a^\psi_{t,h}+b^\psi_{t,h})\\
\rho^\psi_{t,h} &= a^\psi_{t,h}+b^\psi_{t,h}\\
\mu^\psi_{t,h} &\sim \text{Uniform}(0.01, 0.99)\\
\rho^\psi_{t,h} &\sim \text{N}(5,1)\text{, truncated to 0.01,10}\\
\end{aligned}
\]

We assumed that detecting an owl depended on an owl being present during
that survey (\(z_{t,h,i} = 1\)) and the probability of detection, which
was related to the broadcast species:

\[
y_{t,h,i,j,k} \sim \text{Bernoulli}(z_{t,h,i}*p_{t,h,i,j,k})
\]

where generally \(\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k\), or more
specifically:

\[
\begin{aligned}
\text{logit}(p_{t,h,i,j,k}) &= \beta_\text{Pre-broadcast}X_\text{Pre-broadcast} + \\
&= \beta_\text{Mottled}X_\text{Mottled}+\\
&= \beta_\text{Pacific}X_\text{Pacific}+\\
&= \beta_\text{Crested}X_\text{Crested}+\\
&= \beta_\text{Black and White}X_\text{Black and White}+\\
&= \beta_\text{Spectacled}X_\text{Spectacled}+\\
&= \beta_\text{Whiskered}X_\text{Whiskered}+\\
&= \beta_\text{Guat Barred}X_\text{Guat Barred}+\\
&= \beta_\text{Stygian}X_\text{Stygian}+\\
&= \beta_\text{Great Horned}X_\text{Great Horned}\\
\end{aligned}
\]

This model provides a consistent probability of detection for all
surveys in the first two minutes of each survey before the broadcast owl
recording was played. Then, the probability of detection for each
post-broadcast time period would depend on the species of broadcast owl.
This allowed species-specific behavior in response to the different
broadcast owl species (Baumgardt et al. 2019). We used means
parameterization such that the coefficients were interpretable as the
effect of that specific broadcast species \(k\) (including the
pre-broadcast time frame). The broadcast species were consistent at each
station across years, but varied by route:

\begin{table}[]
\centering
\caption{Transformations Associated with the Johnson System}
\begin{tabular}{|l|l|l|l|}
\hline
Johnson Family & Transformation & Parameter Conditions & X Condition \\ \hline
$S_B$ & $Z=\gamma + \eta ln(\frac {X - \epsilon} {\lambda + \epsilon - X})$ & $\eta, \lambda >0, -\infty < \gamma, \epsilon < \infty$ & $\epsilon < X < \epsilon + \lambda$ \\ \hline
$S_L$ & $Z=\gamma + \eta ln(X - \epsilon)$ & $\eta >0, -\infty < \gamma, \epsilon < \infty$ & $X > \epsilon$ \\ \hline
$S_U$ & $Z=\gamma + \eta \sinh^{-1}(\frac {X - \epsilon} {\lambda})$ & $\eta, \lambda >0, -\infty < \gamma, \epsilon < \infty$ & $-\infty < X < \infty$ \\ \hline
\end{tabular}
\end{table}

The priors for every logistic model coefficient \(\beta_k\) were chosen
to be near uniform as recommended in Gelman et al. (2008). Specifically,
we used the Cauchy prior as such:
\(\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)\).

\subsubsection{Implementation of Occupany
Model}\label{implementation-of-occupany-model}

We used the R2jags package in R (R Development Core Team 2014,Plummer
(2013)) to implement the occupancy model for three owl species: Mottled,
Spectacled, and FerPy owls. These owls had enough positive detections to
analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137,
and 187 positive detections over the 11 year period, respectively. Based
on our understanding of owl ecology, we assumed that Spectacled and
FerPy owls would not occupy route M1 in Montecristo, so we removed that
route from those occupancy models.

For all three species' occupancy models, we ran 3 chains for 10,000
iterations and 1000 iterations discarded as burn-in, for a total of
27,000 iterations comprising the posterior distributions for each model
parameter. We visually inspected traceplots to verify that chains mixed
well (Supplemental information S1-3xx).

\section{Results}\label{results}

\subsection{Results sub-section}\label{results-sub-section}

\section{Discussion}\label{discussion}

You can cite references using the \textbf{citr} add-in. With
\textbf{citr} you can cite papers in line like ({\textbf{???}}) is fond
of doing, or parenthetically (R Development Core Team 2014). You can
also cite multiple sources at once ({\textbf{???}}, {\textbf{???}},
Plummer 2013). References can be added into the myrefs.bib file
directly, or created with BibTeX through LaTeX.

\section{Figures}\label{figures}

Figures can be included with the following code, but they must be in
.png format to show up in both Word and PDF. They can be referenced for
the PDF with Fig. \ref{figure1}. \emph{Unfortunately, I don't yet know
how to make the reference number show up in Word.}

\begin{figure}[htbp]
\centering
\includegraphics{../output/figures/ferpy_psi_byYr-1.png}
\caption{1. Example figure caption, which can include in-line equations
like \(y = \alpha + \beta x\) and references like ({\textbf{???}}).
\label{figure1}}
\end{figure}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Baumgardt:2019}{}
Baumgardt, J. A., M. L. Morrison, L. A. Brennan, and T. A. Campbell.
2019. Effects of broadcasting calls on detection probability in
occupancy analyses of multiple raptor species. Western North American
Naturalist 79:185--194.

\hypertarget{ref-Gelman:2008}{}
Gelman, A., A. Jakulin, M. G. Pittau, and Y.-S. Su. 2008. A weakly
informative default prior distribution for logistic and other regression
models. The Annals of Applied Statistics 2:1360--1383.

\hypertarget{ref-Plummer:2013}{}
Plummer, M. 2013. JAGS Version 3.4.0 user manual.

\hypertarget{ref-R:2014}{}
R Development Core Team. 2014. R: A Language and Environment for
Statistical Computing. R Foundation for Statistical Computing, Vienna,
Austria.

\end{document}

### 2020-11-03 programs/e01_processing_results_ferpy.R

@@ -74,7 +74,7 @@ psi.post.ferpy <- MCMCsummary(ferpy.jagsout,
                               Rhat = TRUE,
                               n.eff = TRUE,
                               probs = c(0.05, 0.5, 0.95))
-psi.post.ferpy$Year <- rep(1:length(year.index), each = length(route.index))
+psi.post.ferpy$Year <- rep(year.names, each = length(route.index))
 psi.post.ferpy$Route <- rep(c("EI.1", "EI.2", "M.2", "N.1", "N.2"), length(year.index))
 psi.post.ferpy$Species <- "FerPy"
 psi.post.ferpy <- psi.post.ferpy[include.byrow,]

### 2020-11-03 programs/e02_processing_results_Mottd.R

@@ -74,7 +74,7 @@ psi.post.mottd <- MCMCsummary(mottd.jagsout,
                               Rhat = TRUE,
                               n.eff = TRUE,
                               probs = c(0.05, 0.5, 0.95))
-psi.post.mottd$Year <- rep(1:length(year.index), each = length(route.names))
+psi.post.mottd$Year <- rep(year.names, each = length(route.names))
 psi.post.mottd$Route <- rep(c("EI.1", "EI.2", "M.1", "M.2", "N.1", "N.2"), length(year.index))
 psi.post.mottd$Species <- "Mottd"
 psi.post.mottd <- psi.post.mottd[include.byrow,]

### 2020-11-03 programs/e03_processing_results_specd.R

@@ -74,7 +74,7 @@ psi.post.specd <- MCMCsummary(specd.jagsout,
                               Rhat = TRUE,
                               n.eff = TRUE,
                               probs = c(0.05, 0.5, 0.95))
-psi.post.specd$Year <- rep(1:length(year.index), each = length(route.index))
+psi.post.specd$Year <- rep(year.names, each = length(route.index))
 psi.post.specd$Route <- rep(c("EI.1", "EI.2", "M.2", "N.1", "N.2"), length(year.index))
 psi.post.specd$Species <- "Specd"
 psi.post.specd <- psi.post.specd[include.byrow,]

### 2020-11-03 programs/f01_graphing_results_ferpy.R

@@ -1,4 +1,4 @@
-#' # Graphing Occupancy Analysis
+#' # Graphing Occupancy Analysis: FerPy
 #' 
 #' Description: This program graphs occupancy analysis results for owls of El 
 #' Salvador. Plots show the posteriors of Psi and P(detection). 
@@ -57,7 +57,7 @@ ggplot(data = psi.post.ferpy,
   theme_minimal()+
   xlab("Year")+
   ylab("Probability of Occupancy")+
-  ggtitle("FerPy")
+  scale_x_continuous(breaks = 2003:2013)
 
 
 

### 2020-11-03 programs/f02_graphing_results_Mottd.R

@@ -1,4 +1,4 @@
-#' # Graphing Occupancy Analysis
+#' # Graphing Occupancy Analysis: Mottled
 #' 
 #' Description: This program graphs occupancy analysis results for owls of El 
 #' Salvador. Plots show the posteriors of Psi and P(detection). 
@@ -57,7 +57,7 @@ ggplot(data = psi.post.mottd,
   theme_minimal()+
   xlab("Year")+
   ylab("Probability of Occupancy")+
-  ggtitle("Mottd")
+  scale_x_continuous(breaks = 2003:2013)
 
 
 

### 2020-11-03 programs/f03_graphing_results_specd.R

@@ -1,4 +1,4 @@
-#' # Graphing Occupancy Analysis
+#' # Graphing Occupancy Analysis: Spectacled
 #' 
 #' Description: This program graphs occupancy analysis results for owls of El 
 #' Salvador. Plots show the posteriors of Psi and P(detection). 
@@ -57,7 +57,7 @@ ggplot(data = psi.post.specd,
   theme_minimal()+
   xlab("Year")+
   ylab("Probability of Occupancy")+
-  ggtitle("Specd")
+  scale_x_continuous(breaks = 2003:2013)
 
 
 

### 2020-11-04 manuscript_working/ms_working_RMD.Rmd
broadcast results wrt detection

@@
 The broadcast species were consistent at each station across years, but varied by route.
-For stations $j = 1, 2, \ldots, 10$ at routes EI1, EI2, M2, N1, and N2, the following broadcast species were used, respectively: Pacific, Mottled, Crested, Black and White, Spectacled, Mottled, Crested, Black and White, and Spectacled.
-At the route M1, the following species were called at stations $j = 1, 2, \ldots, 10$, respectively: Whiskered, Mottled, Gaut Barred, Stygian, Great Horned, Whiskered, Mottled, Gaut Barred, Stygian, and Great Horned.
+For stations $j = 1, 2, \ldots, 10$ at El Imposible and Nancachaname, the following broadcast species were used, respectively: Pacific, Mottled, Crested, Black and White, Spectacled, Mottled, Crested, Black and White, and Spectacled.
+At Montecristo, the following species were called at stations $j = 1, 2, \ldots, 10$, respectively: Whiskered, Mottled, Gaut Barred, Stygian, Great Horned, Whiskered, Mottled, Gaut Barred, Stygian, and Great Horned.
 Thus, the effective sample sizes of the different broadcast species varied.

### 2020-11-04 programs/b_richness_model_global.R
ready to run richness analysis

@@ -49,7 +49,8 @@ model.richness <- function(){
         b.psi[hh,tt,ss] <- rho.psi-(mu.psi[hh,tt]*rho.psi)
         
         
-        # Prior for Psi, which will vary by route (hh) and year (tt) and species (ss)
+        # Prior for Psi, which will vary by route (hh) and year (tt) and species (ss),
+        #   but based on shared mean probability of occupancy for each route/year (mu.psi)
         psi[hh,tt,ss] ~ dbeta(a.psi[hh,tt,ss], b.psi[hh,tt,ss])%_%T(0.0001,0.99)
 
         

### 2020-11-04 programs/c04_process_data_richness.R

@@ -93,7 +93,7 @@ n.species <- length(species.names)
 #' 
 #' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401)
 #' 
-n.aug <- 5 #number of species that may have been present but were undetected
+n.aug <- 4 #number of species that may have been present but were undetected
 n.species.aug <- n.species + n.aug
 
 
@@ -140,11 +140,11 @@ for(hh in 1:n.route){
 }
 #' Check a couple examples:
 #' 
-ks.crested[1,,] #EI1
-ks.mottled[1,,] #EI1
-ks.mottled[3,,] #M1
-ks.stygian[3,,] #M1
-ks.stygian[5,,] #N1
+ks.crested[1,,] #EI1, Crested
+ks.mottled[1,,] #EI1, Mottled
+ks.mottled[3,,] #M1, Mottled
+ks.stygian[3,,] #M1, Stygian
+ks.stygian[5,,] #N1, Stygian
 
 
 
@@ -232,7 +232,7 @@ for(hh in 1:n.route){
 #' 
 #' Example data
 ys.aug[1,1,1,1,,] #EI1, 2003, Mottd, Survey 1
-ys.aug[1,1,3,1,,]
+ys.aug[1,1,3,1,,] #EI1, 2003, FerPy, Survey 1
 # Corresponds with these data:
 tab.owls[tab.owls$Stations_ID %in% c(92:98,100:101),2:7]
 
@@ -248,7 +248,7 @@ tab.owls[tab.owls$Stations_ID %in% c(615:617),2:7]
 
 #' Results for augmented species
 ys.aug[1,1,11,1,,] #EI1, 2003, Augmented Species 1, Survey 1
-ys.aug[1,8,11,1,,]
+ys.aug[1,8,11,1,,] #EI1, 2010, Augmented Species 1, Survey 1 
 
 
 

### 2020-11-04 programs/d04_richness_analysis.R

@@ -71,7 +71,7 @@ n.species <- length(species.names)
 #' 
 #' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401)
 #' 
-n.aug <- 5 #number of species that may have been present but were undetected
+n.aug <- 4 #number of species that may have been present but were undetected
 n.species.aug <- n.species + n.aug
 
 
@@ -159,7 +159,7 @@ richness.jagsout <- jags(data = richness.jags.data,
                       n.chains = 3,
                       n.iter = 10000, #10000
                       n.burnin = 1000, #1000
-                      n.thin = 1)
+                      n.thin = 10)
 
 
 

### 2020-11-04 programs/d04_richness_analysis.R
richness output

@@ -168,7 +168,7 @@ richness.jagsout <- jags(data = richness.jags.data,
 #' ## Save files
 #' 
 #' Jagsout Mottd
-save(mottd.jagsout, file = "data/output_data/mottd_jagsout.Rdata")
+save(richness.jagsout, file = "data/output_data/richness_jagsout.Rdata")
 
 
 

### 2020-11-10 programs/b_richness_model_global.R
processing richness output from jags

@@ -126,14 +126,14 @@ model.richness <- function(){
   
   # Derived quantities
   for(ss in 1:n.species.aug){
-    spp.occ[ss] <- sum(z[,,,ss]) # number of occupied routes/years among sampled
+    spp.occ[ss] <- sum(z[,,ss,]) # number of occupied routes/years among sampled
     species.present[ss] <- ifelse(spp.occ[ss]>0, 1, 0) # if each species exists
   }
   Nsmall <- sum(species.present) # small estimate of number of species present
   
   for(hh in 1:n.route){
     for(tt in 1:n.year){
-      richness[hh,tt] <- sum(z[hh,tt,,], na.rm = T) # number of species at each route/year
+      richness[hh,tt] <- sum(z[hh,tt,,]) # number of species at each route/year
     }
   }
 }

### 2020-11-10 programs/d04_richness_analysis.R

@@ -139,7 +139,7 @@ z.init[5,6,,] # rows are different species, N1, 2008
 
 #' ## Run models
 #' 
-#' Mottd
+#' Richness
 richness.jagsout <- jags(data = richness.jags.data, 
                       inits = function(){list(z = z.init, w = w.init)}, 
                       parameters.to.save = c("mu.psi",
@@ -160,9 +160,10 @@ richness.jagsout <- jags(data = richness.jags.data,
                                              "beta.ghorned"), 
                       model.file = model.richness, 
                       n.chains = 3,
-                      n.iter = 100000, #10000
-                      n.burnin = 10000, #1000
+                      n.iter = 20000, #10000
+                      n.burnin = 2000, #1000
                       n.thin = 20)
+
 
 
 

### 2020-11-10 programs/e04_processing_results_richness.R

@@ -34,11 +34,19 @@ set.seed(258854)
 #' Jagsout Richness Model
 load(file = "data/output_data/richness_jagsout.Rdata")
 
+#' Access tables, in Rdata format
+#' 
+load(file = "data/processed_data/tables_global.Rdata")
+
 #' Survey list
 (route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
 route.index <- 1:length(route.names)
 (year.names <- 2003:2013)
 year.index <- 1:length(year.names)
+
+#' #' There were 10 observed owls (if we pool all three unknown observations)
+(species.names <- unique(tab.owls$Owl_Species_ID))
+n.species <- length(species.names)
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy
@@ -67,95 +75,69 @@ all.rows <- 1:(length(route.names)*length(year.index))
 (include.byrow <- all.rows[!all.rows %in% exclude.byrow])
 #' 
 
-#' Specd
+#' Richness
 #' 
-mupsi.post.richness <- MCMCsummary(richness.jagsout, 
-                              params = "mu.psi", 
+richness.post <- MCMCsummary(richness.jagsout, 
+                              params = "richness", 
                               Rhat = TRUE,
                               n.eff = TRUE,
                               probs = c(0.05, 0.5, 0.95))
-mupsi.post.richness$Year <- rep(year.names, each = length(route.index))
-mupsi.post.richness$Route <- 
-  rep(c("EI.1", "EI.2", "M.1", "M.2", "N.1", "N.2"), length(year.index))
-mupsi.post.richness <- mupsi.post.richness[include.byrow,]
+richness.post$Year <- rep(year.names, each = length(route.index))
+richness.post$Route <- 
+  rep(c("EI1", "EI2", "M1", "M2", "N1", "N2"), length(year.index))
+richness.post <- richness.post[include.byrow,]
 
 
 
 
 #' Populate Region
 #' 
-psi.post.specd$Region[psi.post.specd$Route == "EI.1" |
-                        psi.post.specd$Route == "EI.2"] <- "El Imposible"
-psi.post.specd$Region[psi.post.specd$Route == "N.1" |
-                        psi.post.specd$Route == "N.2"] <- "Nancuchiname"
-psi.post.specd$Region[psi.post.specd$Route == "M.1" |
-                        psi.post.specd$Route == "M.2"] <- "Montecristo"
-colnames(psi.post.specd)
[…]

### 2020-11-10 programs/e04_processing_results_richness.R
graphing richness results

@@ -70,8 +70,8 @@ n.species <- length(species.names)
 #' and year
 #' 
 #' For all analysis, remove years and routes that were not surveyed
-(exclude.byrow <- c(3,16:20,33,37,47))
-all.rows <- 1:(length(route.names)*length(year.index))
+(exclude.byrow <- c(3,4,9,15,19:24,27,33,40,44,56,63,64))
+all.rows <- 1:(length(route.index)*length(year.index))
 (include.byrow <- all.rows[!all.rows %in% exclude.byrow])
 #' 
 

### 2020-11-10 programs/g_graphing_results_global.R

@@ -33,6 +33,7 @@ set.seed(587453)
 load(file = "data/output_data/specd_psi_posteriors_RtYr.Rdata")
 load(file = "data/output_data/ferpy_psi_posteriors_RtYr.Rdata")
 load(file = "data/output_data/mottd_psi_posteriors_RtYr.Rdata")
+load(file = "data/output_data/richness_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
 load(file = "data/output_data/specd_psi_posteriors_RtSpp.Rdata")
@@ -44,6 +45,11 @@ load(file = "data/output_data/mottd_psi_posteriors_RtSpp.Rdata")
 load(file = "data/output_data/specd_p_detection_posteriors.Rdata")
 load(file = "data/output_data/ferpy_p_detection_posteriors.Rdata")
 load(file = "data/output_data/mottd_p_detection_posteriors.Rdata")
+load(file = "data/output_data/richness_p_detection_posteriors.Rdata")
+
+#' Species accounts
+#' 
+load(file = "data/output_data/richness_species_accounts.Rdata")
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy
@@ -104,10 +110,92 @@ ggplot(data = p.det.post, aes(y = median.plogis, x = Broadcast, group = Species)
         strip.text.x  = element_blank())
 
 #' _____________________________________________________________________________
-#' ## Save files
+#' ## Richness p = Probability of detection
 #' 
+#' Probability of detection was a function of what broadcast species was used, 
+#' with a constant probability of detection for all pre-broadcast time periods.
+#' 
+#' 
+#+ p_detection_richness, fig.height = 4
+ggplot(data = p.det.post.richness, aes(y = median.plogis, x = Broadcast))+
+  geom_bar(stat = "identity", position = position_dodge() )+
+  geom_linerange(aes(ymin = LL05.plogis, ymax = UL95.plogis), position = position_dodge(0.9))+
+  scale_fill_manual(values = c("#7fc97f", "#beaed4", "#fdc086"))+
+  geom_hline(data = p.det.post.richness[p.det.post.richness$broadcast.param == "beta.prebroad",], 
+             aes(yintercept = median.plogis))+
[…]

### 2020-11-10 programs/d04_richness_analysis.R
spun documents for richness model

@@ -139,7 +139,8 @@ z.init[5,6,,] # rows are different species, N1, 2008
 
 #' ## Run models
 #' 
-#' Richness
+#' Richness, 
+#+ jags, eval = FALSE
 richness.jagsout <- jags(data = richness.jags.data, 
                       inits = function(){list(z = z.init, w = w.init)}, 
                       parameters.to.save = c("mu.psi",
@@ -172,6 +173,7 @@ richness.jagsout <- jags(data = richness.jags.data,
 #' ## Save files
 #' 
 #' Jagsout Mottd
+#+ savejags, eval = F
 save(richness.jagsout, file = "data/output_data/richness_jagsout.Rdata")
 
 
@@ -183,5 +185,5 @@ save(richness.jagsout, file = "data/output_data/richness_jagsout.Rdata")
 devtools::session_info()
 #' This document was "spun" with:
 #' 
-#' ezspin(file = "programs/d02_occupancy_analysis_mottd.R", out_dir = "output", fig_dir = "figures", keep_md = F)
+#' ezspin(file = "programs/d04_richness_analysis.R", out_dir = "output", fig_dir = "figures", keep_md = F)
 #' 

### 2020-11-10 manuscript_working/ms_working_RMD.Rmd
updating methods with richness model

@@
 author: "Althea Archer"
-date: "11/3/2020"
+date: "11/10/2020"
 output:
@@
 For all three species' occupancy models, we ran 3 chains for 10,000 iterations and 1000 iterations discarded as burn-in, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+Initial values for $z_{t,h,i}$ were equivalent to one if an owl was detected in that survey $i$ and otherwise zero.
 We visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
+
+## Richness Model Framework
+
+To model species richness at each route in each year, we assumed that there were 14 possible species present in El Salvador.
+This included the 10 species observed during our surveys (including unknown owl species) and an additional 4 unidentified species.
+This limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus one species we identified (Stygian owl), which was previously undocumented in El Salvador.
+We augmented our owl detection data with 4 additional potential species with all-zero detection [@Royle:2007].
+For each species $s$ (where $s = 1,2, \ldots,14$), $w_{t,h,s}$ is a binary indicator of that owl species being present in each route and year:
+
+$$
+w_{t,h,s} \sim \text{Bernoulli}(\Omega_h)
+$$
+where $\Omega_h$ is the probability of species presence in each route $h$.
+Now, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, is conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belong to that route's community ($w_{t,h,s}$):
+
+$$
+z_{t,h,i,s} \sim \text{Bernoulli}(w_{t,h,s}*\psi_{t,h,s})
+$$
+
+Richness is conditional on each species presence and was estimated as the sum of owl species present in each route and year:
+
+$$
+\text{Richness}_{t,h} = \sum_{s=1}^{14} w_{t,h,s} 
+$$
+
+To share information about occupancy amongst species, we incorporated random effects into the parameterization of $\psi_{t,h,s}$:
+
+\[
+\begin{aligned}
+\psi_{t,h,s} &\sim \text{Beta}(a^\psi_{t,h,s},b^\psi_{t,h,s})\\ 
+\text{logit}(\mu^\psi_{t,h}) &= \text{E}(\psi_{t,h,s}) \sim N(0,2.25)\\
+\ln(\rho^\psi) &\sim N(5,1)\text{, truncated to (0.01,10)}\\
+a^\psi_{t,h,s} &= \mu^\psi_{t,h}\rho^\psi\\
+b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
+\end{aligned}
+\]
+
+This parameterization allowed us to borrow strength across species in determining the mean probability of occupancy ($\mu^\psi_{t,h}$) for each route and year [@Guillera-Arroita:2019].
+
+Consistent with our approach for the occupancy models, we assumed that detecting an owl depended on an owl of that species being present during that survey ($z_{t,h,i,s} = 1$) and the probability of detection $p_{t,h,i,j,k,s})$, which was related to the broadcast species:
+
+$$
+y_{t,h,i,j,k,s} \sim \text{Bernoulli}(z_{t,h,i,s}*p_{t,h,i,j,k,s})
+$$
+
+where generally $\text{logit}(p_{t,h,i,j,k,s}) = \beta_kX_k$ for each broadcast species $k$ as described above.
+The coefficients of the detection model were thus assumed to be consistent for all owl species, which allowed us to borrow information about the detection process for undetected owls from those that were detected.
+
+### Richness Model Implementation
+
+We used the R2jags package in R [@R:2014,@Plummer:2013] to implement the richness model. 
+We ran 3 chains for 20,000 iterations and 2000 iterations discarded as burn-in and a thinning rate of 2, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl $s$ was detected in that survey $i$ and otherwise zero.
+Similarly, initial values for $w_{t,h,s}$ were equivalent to one if that species of owl was detected in that route and year.
+We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
 

### 2020-11-10 programs/b_richness_model_global.R

@@ -133,7 +133,7 @@ model.richness <- function(){
   
   for(hh in 1:n.route){
     for(tt in 1:n.year){
-      richness[hh,tt] <- sum(z[hh,tt,,]) # number of species at each route/year
+      richness[hh,tt] <- sum(w[hh,tt,]) # number of species at each route/year
     }
   }
 }

### 2020-11-10 programs/d04_richness_analysis.R

@@ -163,7 +163,7 @@ richness.jagsout <- jags(data = richness.jags.data,
                       n.chains = 3,
                       n.iter = 20000, #10000
                       n.burnin = 2000, #1000
-                      n.thin = 20)
+                      n.thin = 2)
 
 
 

### 2020-11-11 programs/d04_richness_analysis.R
results, processed, from richness analysis of sum(w)

@@ -166,16 +166,152 @@ richness.jagsout <- jags(data = richness.jags.data,
                       n.thin = 2)
 
 
+#' _____________________________________________________________________________
+#' ## Post-process Psi = Probability of occupancy
+#' 
+#' Survey list
+(route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
+route.index <- 1:length(route.names)
+(year.names <- 2003:2013)
+year.index <- 1:length(year.names)
+#' ### Calculate median, 95% quantile and 5% quantile of psi posteriors by route
+#' and year
+#' 
+#' For all analysis, remove years and routes that were not surveyed
+(exclude.byrow <- c(3,4,9,15,19:24,27,33,40,44,56,63,64))
+all.rows <- 1:(length(route.index)*length(year.index))
+(include.byrow <- all.rows[!all.rows %in% exclude.byrow])
+#' 
+
+#' Richness
+#' 
+richness.post <- MCMCsummary(richness.jagsout, 
+                             params = "richness", 
+                             Rhat = TRUE,
+                             n.eff = TRUE,
+                             probs = c(0.05, 0.5, 0.95))
+richness.post$Year <- rep(year.names, each = length(route.index))
+richness.post$Route <- 
+  rep(c("EI1", "EI2", "M1", "M2", "N1", "N2"), length(year.index))
+richness.post <- richness.post[include.byrow,]
 
 
+
+
+#' Populate Region
+#' 
+richness.post$Region[richness.post$Route == "EI1" |
+                       richness.post$Route == "EI2"] <- "El Imposible"
+richness.post$Region[richness.post$Route == "N1" |
+                       richness.post$Route == "N2"] <- "Nancuchiname"
+richness.post$Region[richness.post$Route == "M1" |
+                       richness.post$Route == "M2"] <- "Montecristo"
+colnames(richness.post)
+colnames(richness.post) <- c("Richness.mean", "Richness.sd", "Richness.LL05", 
+                             "Richness.median",
+                             "Richness.UL95", "Richness.Rhat", "Richness.neff", 
+                             "Year", "Route", "Region")
+
+
+
+#' _____________________________________________________________________________
+#' ## Calculate number of species detected in each year/route
+#' 
+#' Create a table to determine what species were detected in what routes/years
+species.accounts <- as.data.frame(matrix(
+  NA, nrow = nrow(richness.post), ncol = n.species
+))
[…]

### 2020-11-11 programs/e04_processing_results_richness.R

@@ -54,11 +54,11 @@ n.species <- length(species.names)
 #' Create traceplots of parameters (creates PDF)
 #'
 #' Richness
-# MCMCtrace(richness.jagsout,
-#           pdf = TRUE,
-#           open_pdf = FALSE,
-#           filename = "richnessTrace",
-#           wd = "output")
+MCMCtrace(richness.jagsout,
+          pdf = TRUE,
+          open_pdf = FALSE,
+          filename = "richnessTrace",
+          wd = "output")
 
 
 

### 2020-11-11 programs/e04_processing_results_richness.R
processed jagsout rather than saving it (>100MB)

@@ -44,7 +44,8 @@ route.index <- 1:length(route.names)
 (year.names <- 2003:2013)
 year.index <- 1:length(year.names)
 
-#' #' There were 10 observed owls (if we pool all three unknown observations)
+#' 
+#' There were 10 observed owls (if we pool all three unknown observations)
 (species.names <- unique(tab.owls$Owl_Species_ID))
 n.species <- length(species.names)
 
@@ -54,151 +55,22 @@ n.species <- length(species.names)
 #' Create traceplots of parameters (creates PDF)
 #'
 #' Richness
-MCMCtrace(richness.jagsout,
-          pdf = TRUE,
-          open_pdf = FALSE,
-          filename = "richnessTrace",
-          wd = "output")
+# MCMCtrace(richness.jagsout,
+#           pdf = TRUE,
+#           open_pdf = FALSE,
+#           filename = "richnessTrace",
+#           wd = "output")
 
 
 
 
-#' _____________________________________________________________________________
-#' ## Psi = Probability of occupancy
-#' 
-#' ### Calculate median, 95% quantile and 5% quantile of psi posteriors by route
-#' and year
-#' 
-#' For all analysis, remove years and routes that were not surveyed
-(exclude.byrow <- c(3,4,9,15,19:24,27,33,40,44,56,63,64))
-all.rows <- 1:(length(route.index)*length(year.index))
-(include.byrow <- all.rows[!all.rows %in% exclude.byrow])
-#' 
-
-#' Richness
-#' 
-richness.post <- MCMCsummary(richness.jagsout, 
-                              params = "richness", 
-                              Rhat = TRUE,
-                              n.eff = TRUE,
-                              probs = c(0.05, 0.5, 0.95))
-richness.post$Year <- rep(year.names, each = length(route.index))
-richness.post$Route <- 
-  rep(c("EI1", "EI2", "M1", "M2", "N1", "N2"), length(year.index))
-richness.post <- richness.post[include.byrow,]
 
 
-
-
-#' Populate Region
-#' 
-richness.post$Region[richness.post$Route == "EI1" |
-                       richness.post$Route == "EI2"] <- "El Imposible"
-richness.post$Region[richness.post$Route == "N1" |
-                       richness.post$Route == "N2"] <- "Nancuchiname"
-richness.post$Region[richness.post$Route == "M1" |
-                       richness.post$Route == "M2"] <- "Montecristo"
-colnames(richness.post)
[…]

### 2020-11-11 programs/b_occupancy_model_global.R
htmls with updated hyperpriors for occ analysis

@@ -27,7 +27,8 @@ model.occ <- function(){
   for(hh in 1:n.route){
     for(tt in 1:n.year){
       #Hyperpriors for Psi ensure a flexible and uninformed prior
-      mu.psi[hh,tt] ~ dunif(0.01,0.99)
+      logit.mu.psi[hh,tt] ~ dnorm(0, 0.2)
+      mu.psi[hh,tt] <- exp(logit.mu.psi[hh,tt])/(1+exp(logit.mu.psi[hh,tt]))
       lnrho.psi[hh,tt] ~ dnorm(5,1)%_%T(0.01,10)
       rho.psi[hh,tt] <- exp(lnrho.psi[hh,tt])
       a.psi[hh,tt] <- mu.psi[hh,tt]*rho.psi[hh,tt]

### 2020-11-11 programs/c02_process_data_Mottd.R

@@ -93,7 +93,7 @@ ks.prebroad <- ks.mottled <- ks.pacific <- ks.crested <- ks.bw <- ks.spectacled 
 
 broadcast.array <- array(c(
   rep(c("pacific", "mottled", "crested", "bw", "spectacled"),4), # routes EI1/EI2
-  rep(c("whiskered", "mottled", "gbarred", "stygian", "ghorned"),2), # routes M1/M2
+  rep(c("whiskered", "mottled", "gbarred", "stygian", "ghorned"),4), # routes M1/M2
   rep(c("pacific", "mottled", "crested", "bw", "spectacled"),4)), # routes N1/N2
   dim = c(n.station, n.route))
 broadcast.array

### 2020-11-11 manuscript_working/ms_working_RMD.Rmd
updating manuscript with richness results/methods

@@
 author: "Althea Archer"
-date: "11/10/2020"
+date: "11/11/2020"
 output:
@@
 # Methods
+
+``` {r}
+load("../data/output_data/richness_species_accounts_byRt.Rdata")
+library(pander)
+pander(species.accounts.byRt)
+```
 
@@
 \rho^\psi_{t,h} &= a^\psi_{t,h}+b^\psi_{t,h}\\
-\mu^\psi_{t,h} &\sim \text{Uniform}(0.01, 0.99)\\
-\rho^\psi_{t,h} &\sim \text{N}(5,1)\text{, truncated to 0.01,10}\\
+\text{logit}(\mu^\psi_{t,h}) &\sim \text{N}(0, 2.25)\\
+\ln(\rho^\psi_{t,h}) &\sim \text{N}(5,1)\text{, truncated to 0.01,10}\\
 \end{aligned}
@@
 
-We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the broadcast species:
+We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the species of owl played for broadcast ($k$):
 
@@
 
-where generally $\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k$, or more specifically:
+where, generally, $\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k$, or more specifically:
 
@@
 
-This model provides a consistent probability of detection for all surveys in the first two minutes of each survey before the broadcast owl recording was played.
+This model provides a consistent probability of detection for all surveys in the first two minutes of each survey before the broadcast owl recording was played ($k = \text{Pre-broadcast}$).
 Then, the probability of detection for each post-broadcast time period would depend on the species of broadcast owl.
@@
 These owls had enough positive detections to analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137, and 187 positive detections over the 11 year period, respectively.
-Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed that route from those occupancy models.
+Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy models.
 
@@
 We augmented our owl detection data with 4 additional potential species with all-zero detection [@Royle:2007].
-For each species $s$ (where $s = 1,2, \ldots,14$), $w_{t,h,s}$ is a binary indicator of that owl species being present in each route and year:
+For each species $s$ (where $s = 1,2, \ldots,14$), $w_{t,h,s}$ was a binary indicator of that owl species being present in each route and year:
 
@@
 $$
-where $\Omega_h$ is the probability of species presence in each route $h$.
-Now, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, is conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belong to that route's community ($w_{t,h,s}$):
+where $\Omega_h$ was the probability of species presence in each route $h$.
+Now, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belong to that route's community ($w_{t,h,s}$):
 
@@
 
-Richness is conditional on each species presence and was estimated as the sum of owl species present in each route and year:
+Richness was conditional on each species presence and was estimated as the sum of owl species present in each route and year:
 
@@
 \psi_{t,h,s} &\sim \text{Beta}(a^\psi_{t,h,s},b^\psi_{t,h,s})\\ 
+a^\psi_{t,h,s} &= \mu^\psi_{t,h}\rho^\psi\\
+b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
 \text{logit}(\mu^\psi_{t,h}) &= \text{E}(\psi_{t,h,s}) \sim N(0,2.25)\\
 \ln(\rho^\psi) &\sim N(5,1)\text{, truncated to (0.01,10)}\\
-a^\psi_{t,h,s} &= \mu^\psi_{t,h}\rho^\psi\\
-b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
 \end{aligned}
@@
 
-Consistent with our approach for the occupancy models, we assumed that detecting an owl depended on an owl of that species being present during that survey ($z_{t,h,i,s} = 1$) and the probability of detection $p_{t,h,i,j,k,s})$, which was related to the broadcast species:
+Consistent with our approach for the occupancy models, we assumed that detecting an owl depended on an owl of that species being present during that survey ($z_{t,h,i,s} = 1$) and the probability of detection $p_{t,h,i,j,k,s}$, which was related to the broadcast species:
 
@@
 Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl $s$ was detected in that survey $i$ and otherwise zero.
-Similarly, initial values for $w_{t,h,s}$ were equivalent to one if that species of owl was detected in that route and year.
+Similarly, initial values for $w_{t,h,s}$ were equivalent to one if that species of owl was detected in that route and year and otherwise zero.
 We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
@@
 4XX).
+
+## Richness Model Results
+
+## Detection Results
 

### 2020-11-11 programs/e04_processing_results_richness.R

@@ -32,11 +32,15 @@ set.seed(258854)
 #' 
 #' 
 #' Jagsout Richness Model
-load(file = "data/output_data/richness_jagsout.Rdata")
+#load(file = "data/output_data/richness_jagsout.Rdata")
 
 #' Access tables, in Rdata format
 #' 
 load(file = "data/processed_data/tables_global.Rdata")
+
+#' Species accounts to summarize by route
+#' 
+load("../data/output_data/richness_species_accounts.Rdata")
 
 #' Survey list
 (route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
@@ -62,15 +66,30 @@ n.species <- length(species.names)
 #           wd = "output")
 
 
-
-
-
-
+#' Species presence by route
+#' 
+species.accounts.byRt <- as.data.frame(matrix(
+  NA, nrow = length(unique(species.accounts$Route)), 
+  ncol = ncol(species.accounts)
+  ))
+colnames(species.accounts.byRt) <- colnames(species.accounts)
+species.accounts.byRt$Route <- unique(species.accounts$Route)
+species.accounts.byRt$Year <- c("2003-2013","2003-2013",
+                                "2003-2013","2003-2013",
+                                "2009-2012","2004-2012")
+species.accounts.byRt$Region <- rep(unique(species.accounts$Region),each=2)
+for(hh in 1:nrow(species.accounts.byRt)){
+  temp.route <- species.accounts.byRt$Route[hh]
+  temp.accounts <- species.accounts[species.accounts$Route == temp.route,]
+  for(ss in 1:10){ #across all 10 species
+    species.accounts.byRt[hh,ss] <- ifelse(sum(temp.accounts[,ss],na.rm = T)>0,1,0)
+  }
+}
 
 #' _____________________________________________________________________________
 #' ## Save files
 #' 
-
+save(species.accounts.byRt, file = "data/output_data/richness_species_accounts_byRt.Rdata")
 
 #' _____________________________________________________________________________
 #' ### Footer

### 2020-11-11 programs/g_graphing_results_global.R

@@ -86,28 +86,32 @@ ggplot(data = psi.means, aes(x = Route, y = Psi.median, group = Species))+
 #' ## p = Probability of detection
 #' 
 #' Merge data
-p.det.post <- rbind(p.det.post.ferpy, p.det.post.mottd, p.det.post.specd)
+p.det.post.richness$Species <- "All Species"
+#' 
+p.det.post <- rbind(p.det.post.ferpy, p.det.post.mottd, p.det.post.specd,p.det.post.richness)
 #' Probability of detection was a function of what broadcast species was used, 
 #' with a constant probability of detection for all pre-broadcast time periods.
 #' 
 #' 
-#+ p_detection, fig.height = 8
+#+ p_detection, fig.height = 10
 ggplot(data = p.det.post, aes(y = median.plogis, x = Broadcast, group = Species))+
   geom_bar(stat = "identity", position = position_dodge(), aes(fill = Species))+
-  geom_linerange(aes(ymin = LL05.plogis, ymax = UL95.plogis), position = position_dodge(0.9))+
-  scale_fill_manual(values = c("#7fc97f", "#beaed4", "#fdc086"))+
+  geom_linerange(aes(ymin = LL05.plogis, ymax = UL95.plogis), 
+                 position = position_dodge(0.9))+
+  scale_fill_manual(values = c("darkgrey", "#7fc97f", "#beaed4", "#fdc086"))+
   geom_hline(data = p.det.post[p.det.post$broadcast.param == "beta.prebroad",], 
              aes(yintercept = median.plogis))+
   geom_hline(data = p.det.post[p.det.post$broadcast.param == "beta.prebroad",], 
              aes(yintercept = LL05.plogis), color = "#ffff99")+
   geom_hline(data = p.det.post[p.det.post$broadcast.param == "beta.prebroad",], 
              aes(yintercept = UL95.plogis), color = "#ffff99")+
-  facet_wrap(~Species, nrow = 3)+
+  facet_wrap(~Species, nrow = 4)+
   ylab("Probability of Detection")+
   xlab("Broadcast Species")+
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        strip.text.x  = element_blank())
+        #strip.text.x  = element_blank(),
+        legend.position = "none")
 
 #' _____________________________________________________________________________
 #' ## Richness p = Probability of detection

### 2020-11-11 programs/c02_process_data_Mottd.R
inserting table of broadcast species for manuscript

@@ -213,6 +213,13 @@ save(ks.prebroad, ks.pacific, ks.mottled, ks.crested, ks.bw, ks.spectacled,
   ks.whiskered, ks.gbarred, ks.stygian, ks.ghorned,
   file = "data/processed_data/mottd_jags_ks.Rdata")
 
+#' Save broadcast array for manuscript
+#' 
+broadcast.array <- as.data.frame(broadcast.array)
+colnames(broadcast.array) <- c("EI1", "EI2", "M1", "M2", "N1", "N2")
+broadcast.array$Station <- 1:10
+save(broadcast.array, file = "data/output_data/broadcast_array.Rdata")
+
 
 #' _____________________________________________________________________________
 #' ### Footer

### 2020-11-11 manuscript_working/ms_working_RMD.Rmd
inserting table of broadcast species for manuscript
preliminary manuscript results/methods

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 # Methods
+
+*Table 1XX.
+Species detection records by route.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route (Year range in table).*
 
@@
 
-The broadcast species were consistent at each station across years, but varied by route.
-For stations $j = 1, 2, \ldots, 10$ at El Imposible and Nancachaname, the following broadcast species were used, respectively: Pacific, Mottled, Crested, Black and White, Spectacled, Mottled, Crested, Black and White, and Spectacled.
-At Montecristo, the following species were called at stations $j = 1, 2, \ldots, 10$, respectively: Whiskered, Mottled, Gaut Barred, Stygian, Great Horned, Whiskered, Mottled, Gaut Barred, Stygian, and Great Horned.
-Thus, the effective sample sizes of the different broadcast species varied.
+The broadcast species were consistent at each station across years, but varied by route (Table 2XX).  Thus, the effective sample sizes of the different broadcast species varied.
+
+*Table 2XX.
+Broadcast species by route and station.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Pre-recorded owl calls were broadcast after 2 minutes of silent listening.*
+
+``` {r, echo = F}
+load(file = "../data/output_data/broadcast_array.Rdata")
+pander(broadcast.array)
+```
 
@@
 
+Median species richness varied from 1 to 5, depending on route and year (Fig.
+5XXX).
+At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species in several years of EI1 surveys and two years of M2 surveys (Fig.
+5XXX).
+The shared information through random effects by route tended to smooth out the year-to-year variation in predicted richness, which in a few instances meant that the median predicted richness was less than the number of observed species in that route and year (Fig.
+5XXX)
+
 ## Detection Results
 
-The probability of detecting these three owls varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast species were called varied by broadcast species and between Mottled, FerPy, and Spectacled owls (Fig.
-5XX).
+Adding in all detected species to one richness model increased the precision of the estimates of the effect of broadcast species on detection probability (i.e., compare All Species with individual occupany results in Fig.
+6XXX).
+When incorporating all detected species, broadcasting calls from Mottled, Pacific, Crested, Black and White, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned owl calls decreased detection (Fig.
+6XXX).
+
+The probability of detection when predicted through independent occupancy analyses varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast species were called varied by broadcast species and between Mottled, FerPy, and Spectacled owls (Fig.
+6XX).
 The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owls.
 The probability of detecting FerPy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owls (Fig.
-5XX).
+6XX).
 
@@
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure1}](../output/figures/mottd_psi_byYr-1.png)
+\label{figure2}](../output/figures/mottd_psi_byYr-1.png)
 
@@
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure1}](../output/figures/ferpy_psi_byYr-1.png)
+\label{figure3}](../output/figures/ferpy_psi_byYr-1.png)
 
@@
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure1}](../output/figures/specd_psi_byYr-1.png)
+\label{figure4}](../output/figures/specd_psi_byYr-1.png)
 
 ![5XX.
-The logistic regression parameters that related the probability of detecting owls ($\p_{t,h,i,j,k}$) to the broadcast timeframe (pre-broadcast) or speccies.
+Species richness by route and year.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Posteriors were summarized with median (black dot) $\pm$ 90% credible intervals (black lines).
+The number of species detected at each route and year are indicated with a red dot.
+\label{figure5}](../output/figures/richness_byRtYr-1.png)
+
+![6XX.
+The logistic regression parameters that related the probability of detecting owls ($\p_{t,h,i,j,k}$) to the broadcast timeframe (pre-broadcast) or species.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure1}](../output/figures/p_detection-1.png)
+\label{figure6}](../output/figures/p_detection-1.png)
 

### 2020-11-13 manuscript_working/ms_working_RMD.Rmd
manuscript draft to send to Jane

@@
 Species detection records by route.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route (Year range in table).*
+Ones indicate that that species was detected during that route and year, and zeros indicated that those owls were not detected, although they may have been present and undetected.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route, which was given in column "Year"*
 
@@
 \begin{aligned}
-t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\\
-h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\\
+t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\text{ corresponding with 2003-2013}\\
+h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\text{ corresponding with EI1, EI2, M1, M2, N1, N2}\\
 i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
@@
 The priors for every logistic model coefficient $\beta_k$ were chosen to be near uniform as recommended in @Gelman:2008.
-Specifically, we used the Cauchy prior as such: $\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)$.
+Specifically, we used the Cauchy prior such that: $\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)$.
 
@@
 
-### Occupany Model Implementation
+### Occupancy Model Implementation
 
@@
 These owls had enough positive detections to analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137, and 187 positive detections over the 11 year period, respectively.
-Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy models.
+Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
 
@@
 where $\Omega_h$ was the probability of species presence in each route $h$.
-Now, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belong to that route's community ($w_{t,h,s}$):
+In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community ($w_{t,h,s}$):
 
@@
 
-## Occupany Model Results
+## Occupancy Model Results
 
@@
 FerPy probability of occupancy was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
-3XX), and the probability of occupany for the Spectacled owl slightly fluctuated year to year in the Nancuchaname routes (Fig.
+3XX), and the probability of occupancy for the Spectacled owl slightly fluctuated year to year in the Nancuchaname routes (Fig.
 4XX).
@@
 
-Adding in all detected species to one richness model increased the precision of the estimates of the effect of broadcast species on detection probability (i.e., compare All Species with individual occupany results in Fig.
+Adding in all detected species to one richness model increased the precision of the estimates of the effect of broadcast species on detection probability (i.e., compare All Species with individual occupancy results in Fig.
 6XXX).
-When incorporating all detected species, broadcasting calls from Mottled, Pacific, Crested, Black and White, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned owl calls decreased detection (Fig.
+When incorporating all owl species, broadcasting calls from Mottled, Pacific, Crested, Black and White, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned owl calls decreased detection (Fig.
 6XXX).
@@
 ![1XX.
-Probability of occupancy ($\psi_{t,h}$) averaged by route $h$ for Mottled, FerPy, and Spectacled Owls.
+Probability of occupancy ($\psi_{t,h}$) averaged by route for Mottled, FerPy, and Spectacled Owls.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 ![2XX.
-Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for Mottled Owl.
+Probability of occupancy by route and year for Mottled Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 ![3XX.
-Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for FerPy Owl.
+Probability of occupancy by route and year for FerPy Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 ![4XX.
-Probability of occupancy ($\psi_{t,h}$) by route $h$ and year $t$ for Spectacled Owl.
+Probability of occupancy by route and year for Spectacled Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 ![6XX.
-The logistic regression parameters that related the probability of detecting owls ($\p_{t,h,i,j,k}$) to the broadcast timeframe (pre-broadcast) or species.
+The logistic regression parameters that related the probability of detecting owls to the broadcast timeframe (pre-broadcast) or species.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2020-11-18 manuscript_working/ms_working_RMD.Rmd
updating manuscript wrt MRE edits

@@
 author: "Althea Archer"
-date: "11/11/2020"
+date: "11/17/2020"
 output:
@@
 
-*Table 1XX.
-Species detection records by route.
-Ones indicate that that species was detected during that route and year, and zeros indicated that those owls were not detected, although they may have been present and undetected.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route, which was given in column "Year"*
+*Table 1XX. Species detection records by route. Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route.*
 
@@
 library(pander)
-pander(species.accounts.byRt)
+pander(species.counts.long[c(3:9)])
 ```
 
-## Occupancy Model Framework
+## Single-Species Occupancy Model Framework
 
-Let
+We modeled occupancy within each route, year, and survey assuming that the probability of occupancy would be closed across surveys of a given route in a given year.
+In other words, for any single species of owl, we assumed that the probability of a route being occupied by that owl species would not change between surveys within a year but that the probability of occupancy would vary between years.
+Let:
 
@@
 \begin{aligned}
-t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\text{ corresponding with 2003-2013}\\
-h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\text{ corresponding with EI1, EI2, M1, M2, N1, N2}\\
+t  &=  \text{the year each survey was conducted (here }t = 1,2,\ldots,11)\text{ corresponding with 2003-2013}\\
+h  &=  \text{the individual route (here }h = 1,2,\ldots,6)\text{ corresponding with EI1, EI2, M1, M2, N1, N2}\\
 i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
 j  &=  \text{the individual stations along each route (here }j = 1,2,\ldots,10)\\
-k  &=  \text{the broadcast species (here }k = 1,2,\ldots,10)\\
+k  &=  \text{the broadcast call species (here }k = 1,2,\ldots,10)\\
 z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
@@
 
-We assumed that survey occupancy ($z_{t,h,i}$) was the outcome of Bernoulli trials with probability $\psi$, which we allowed to vary by year and route:
+We assumed that occupancy in each individual survey ($z_{t,h,i}$) was the outcome of Bernoulli trials with a probability of occupancy, $\psi_{t,h}$, which we allowed to vary by year and route:
 
@@
 
-We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the species of owl played for broadcast ($k$):
+We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the species-specific broadcast call ($k$):
 
@@
 
-This model provides a consistent probability of detection for all surveys in the first two minutes of each survey before the broadcast owl recording was played ($k = \text{Pre-broadcast}$).
-Then, the probability of detection for each post-broadcast time period would depend on the species of broadcast owl.
-This allowed species-specific behavior in response to the different broadcast owl species [@Baumgardt:2019].
-We used means parameterization such that the coefficients were interpretable as the effect of that specific broadcast species $k$ (including the pre-broadcast time frame).
-The priors for every logistic model coefficient $\beta_k$ were chosen to be near uniform as recommended in @Gelman:2008.
-Specifically, we used the Cauchy prior such that: $\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)$.
+This model provides a consistent probability of detection for all surveys in the first two minutes of each observation period, before the broadcast call was played ($k = \text{Pre-broadcast}$).
+Then, the probability of detection for each post-broadcast time period depended which call had been played.
+This allowed species-specific behavior in response to the different broadcast calls [@Baumgardt:2019].
+We used means parameterization for the logistic regression model such that the coefficients were interpretable as the effect of that specific broadcast call $k$ (including the pre-broadcast time frame).
+The priors for every logistic model coefficient $\beta_k$ were chosen to be near uniform as recommended in @Gelman:2008: $\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)$.
 
-The broadcast species were consistent at each station across years, but varied by route (Table 2XX).  Thus, the effective sample sizes of the different broadcast species varied.
+The broadcast calls varied between routes, but were consistent at each route and station across years (Table 2XX).  Thus, the effective sample sizes of the different broadcast calls varied.
 
 *Table 2XX.
-Broadcast species by route and station.
+Broadcast calls by route and station.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 
-### Occupancy Model Implementation
+### Single-Species Occupancy Model Implementation
 
-We used the R2jags package in R [@R:2014,@Plummer:2013] to implement the occupancy model for three owl species: Mottled, Spectacled, and FerPy owls.
-These owls had enough positive detections to analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137, and 187 positive detections over the 11 year period, respectively.
-Based on our understanding of owl ecology, we assumed that Spectacled and FerPy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
+We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the occupancy model for three owl species: Mottled, Spectacled, and Ferruginous Pygmy owls.
+These species had sufficient positive detections to analyze occupancy (Table 1XX).
+Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
 
@@
 
-To model species richness at each route in each year, we assumed that there were 14 possible species present in El Salvador.
-This included the 10 species observed during our surveys (including unknown owl species) and an additional 4 unidentified species.
-This limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus one species we identified (Stygian owl), which was previously undocumented in El Salvador.
-We augmented our owl detection data with 4 additional potential species with all-zero detection [@Royle:2007].
-For each species $s$ (where $s = 1,2, \ldots,14$), $w_{t,h,s}$ was a binary indicator of that owl species being present in each route and year:
+To model species richness at each route in each year, we assumed that there were 15 possible species present in El Salvador.
+This included the 10 species observed during our surveys (including unknown owl species) and an additional 5 species that we never observed during our surveys (Table 1XX).
+This upper limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus our unknown species and one species we identified (Stygian owl), which was previously undocumented in El Salvador.
+We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
+For each species $s$ (where $s = 1,2, \ldots,15$), $w_{t,h,s}$ was a binary indicator of whether that owl species was present in each route and year:
 
@@
 $$
-where $\Omega_h$ was the probability of species presence in each route $h$.
-In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community ($w_{t,h,s}$):
+where $\Omega_h$ was the probability of species presence in each route $h$ [@Broms:2016].
+We used prior recommended by @Broms:2016, @Guillera-Arroita:2019, and others for $\Omega_h$: $\Omega_h \sim \text{Beta}(0.001, 1)$.
+
+In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community that year ($w_{t,h,s}$):
 
@@
 z_{t,h,i,s} \sim \text{Bernoulli}(w_{t,h,s}*\psi_{t,h,s})
-$$
-
-Richness was conditional on each species presence and was estimated as the sum of owl species present in each route and year:
-
-$$
-\text{Richness}_{t,h} = \sum_{s=1}^{14} w_{t,h,s} 
 $$
@@
 b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
-\text{logit}(\mu^\psi_{t,h}) &= \text{E}(\psi_{t,h,s}) \sim N(0,2.25)\\
+\text{logit}(\mu^\psi_{t,h}) &= \sim N(0,2.25)\\
 \ln(\rho^\psi) &\sim N(5,1)\text{, truncated to (0.01,10)}\\
@@
 
-Consistent with our approach for the occupancy models, we assumed that detecting an owl depended on an owl of that species being present during that survey ($z_{t,h,i,s} = 1$) and the probability of detection $p_{t,h,i,j,k,s}$, which was related to the broadcast species:
+Richness was conditional on each species' presence and was estimated as the sum of owl species present in each route and year:
+
+$$
+\text{Richness}_{t,h} = \sum_{s=1}^{14} w_{t,h,s} 
+$$
+
+Consistent with our approach for the occupancy models, we assumed that detecting an owl depended on an individual of that species being present during that survey ($z_{t,h,i,s} = 1$) and the species-specific probability of detection $p_{t,h,i,j,k,s}$, which was related to the broadcast call:
 
@@
 
-where generally $\text{logit}(p_{t,h,i,j,k,s}) = \beta_kX_k$ for each broadcast species $k$ as described above.
-The coefficients of the detection model were thus assumed to be consistent for all owl species, which allowed us to borrow information about the detection process for undetected owls from those that were detected.
+where generally $\text{logit}(p_{t,h,i,j,k,s}) = \beta_kX_k$ for each broadcast call $k$ as described above.
+The coefficients of the detection model were thus assumed to be consistent for all owl species, which allowed us to borrow information about the detection process for undetected owl species from those that were detected.
 
@@
 
-We used the R2jags package in R [@R:2014,@Plummer:2013] to implement the richness model. 
-We ran 3 chains for 20,000 iterations and 2000 iterations discarded as burn-in and a thinning rate of 2, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
-Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl $s$ was detected in that survey $i$ and otherwise zero.
-Similarly, initial values for $w_{t,h,s}$ were equivalent to one if that species of owl was detected in that route and year and otherwise zero.
+We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the richness model.
+We ran 3 chains for 20,000 iterations with 2000 iterations discarded as burn-in and a thinning rate of 2, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl was detected in that survey and otherwise zero.
+Similarly, initial values for whether a species belonged to a route's community ($w_{t,h,s}$) were equivalent to one if that species of owl was detected in that route and year and otherwise zero.
 We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
@@
 
-## Occupancy Model Results
+## Single-Species Occupancy Model Results
 
-Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for FerPy or Spectacled owls in both routes of El Imposible and in M2 route of Montecristo (Fig.
+Mottled, Spectacled, and Ferruginous Pygmy owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 1XX).
+No other owl species had enough detections to analyze species-specific occupancy (Table 1XX).
+Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for Ferruginous Pygmy or Spectacled owls in both routes of El Imposible and in M2 route of Montecristo (Fig.
 1X).
@@
 We found that the probability of occupancy for Mottled owls was relatively consistent across time in El Imposible and in N2 route (Fig.
-2XX), but that it seemed to increase in the last 6 or so years of the surveys in M2 and N1 routes (Fig.
+2XX), but that it seemed to increase in the last ~6 years of the surveys in M2 and N1 routes (Fig.
 2XX).
@@
 
-The probability of occupancy for FerPy and Spectacled stayed relatively consistently low across time in routes EI1, EI2, and M2 other than a higher probability of occupancy for FerPy in EI1 in the early years of the surveys (Figs.
+For routes EI1, EI2, and M2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI1 in the early years of the surveys (Figs.
 3XX and 4XX).
-FerPy probability of occupancy was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
+The probability of occupancy for Ferruginous Pygmy owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
 3XX), and the probability of occupancy for the Spectacled owl slightly fluctuated year to year in the Nancuchaname routes (Fig.
@@
 
-Adding in all detected species to one richness model increased the precision of the estimates of the effect of broadcast species on detection probability (i.e., compare All Species with individual occupancy results in Fig.
+Including all detected species into one richness model increased the precision of the estimates of the effect of broadcast calls on detection probability (i.e., compare All Species with individual occupancy results in Fig.
 6XXX).
@@
 
-The probability of detection when predicted through independent occupancy analyses varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast species were called varied by broadcast species and between Mottled, FerPy, and Spectacled owls (Fig.
+The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by broadcast call and between Mottled, Ferruginous Pygmy, and Spectacled owls (Fig.
 6XX).
-The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owls.
-The probability of detecting FerPy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owls (Fig.
+The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owl calls.
+The probability of detecting Ferruginous Pygmy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owl calls (Fig.
 6XX).
@@
 ![1XX.
-Probability of occupancy ($\psi_{t,h}$) averaged by route for Mottled, FerPy, and Spectacled Owls.
+Probability of occupancy ($\psi_{t,h}$) averaged by route for Mottled, Ferruginous Pygmy, and Spectacled Owls.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
@@
 ![3XX.
-Probability of occupancy by route and year for FerPy Owl.
+Probability of occupancy by route and year for Ferruginous Pygmy Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2020-11-18 programs/b_richness_model_global.R
updating Maug, W_{h,s} not by yr, recalc Richness ests

@@ -89,17 +89,21 @@ model.richness <- function(){
   
   # Likelihood
   for(hh in 1:n.route){ # 6 routes
+    
+
+    
+    
     for(tt in 1:n.year){ # all years
       for(ss in 1:n.species.aug){ # all augmented species
         
-        # Is each species present in that route and year?
-        w[hh,tt,ss] ~ dbern(omega[hh])
+        # Is each species present in that route's community?
+        w[hh,ss] ~ dbern(omega[hh])
         
         for(ii in 1:n.survey){ # 1 to 3 surveys per year
           
           # Occupancy by route and year and survey based on 
           #     psi, probability of occupancy for each route/year
-          z[hh,tt,ss,ii] ~ dbern(psi[hh,tt,ss]*w[hh,tt,ss]) 
+          z[hh,tt,ss,ii] ~ dbern(psi[hh,tt,ss]*w[hh,ss]) 
           
           
           
@@ -125,15 +129,29 @@ model.richness <- function(){
   } # end likelihood
   
   # Derived quantities
-  for(ss in 1:n.species.aug){
-    spp.occ[ss] <- sum(z[,,ss,]) # number of occupied routes/years among sampled
-    species.present[ss] <- ifelse(spp.occ[ss]>0, 1, 0) # if each species exists
+  for(hh in 1:n.route){
+    for(ss in 1:n.species.aug){
+      # number of occupied years
+      spp.occ[hh,ss] <- sum(z[hh,,ss,]) 
+      
+      # if each species exists by route
+      species.present[hh,ss] <- ifelse(spp.occ[hh,ss]>0, 1, 0) 
+    }
   }
-  Nsmall <- sum(species.present) # small estimate of number of species present
+
   
   for(hh in 1:n.route){
+    
+    richness[hh] <- sum(w[hh,]) # number of species at each route
+    
     for(tt in 1:n.year){
-      richness[hh,tt] <- sum(w[hh,tt,]) # number of species at each route/year
+      for(ss in 1:n.species.aug){
+        # If each species was present in each route/year
+        occ.RtYrSpp[hh,tt,ss] <- max(z[hh,tt,ss,])
+        
+      }
+      # Sum up if each species was present, to get richness estimates by route/year
+      richness.RtYr[hh,tt] <- sum(occ.RtYrSpp[hh,tt,])
     }
   }
 }

### 2020-11-18 programs/d04_richness_analysis.R

@@ -71,7 +71,7 @@ n.species <- length(species.names)
 #' 
 #' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401)
 #' 
-n.aug <- 4 #number of species that may have been present but were undetected
+n.aug <- 5 #number of species that may have been present but were undetected
 n.species.aug <- n.species + n.aug
 
 
@@ -110,12 +110,12 @@ richness.jags.data <- list(
 z.init <- array(NA, 
                 dim = c(n.route, n.year, n.species.aug, n.survey))
 w.init <- array(NA,
-                dim = c(n.route, n.year, n.species.aug))
+                dim = c(n.route, n.species.aug))
 
 for(hh in 1:n.route){
   for(tt in 1:n.year){
     for(ss in 1:n.species.aug){
-      w.init[hh,tt,ss] <- ifelse(ss <= n.species, yes = 1, no = 0)
+
       
       for(ii in 1:n.survey){
         if(sum(is.na(ys.aug[hh,tt,ss,ii,,])==FALSE)==0){ # no survey conducted here
@@ -127,10 +127,22 @@ for(hh in 1:n.route){
     }
   }
 }
+for(hh in 1:n.route){
+    for(ss in 1:n.species){
+      w.init[hh,ss] <- max(z.init[hh,,ss,], na.rm = T)
+    }
+    for(ss in (n.species+1):n.species.aug){
+      w.init[hh,ss] <- 0
+    }
+}
+
 #' Check initials (should be summary of ys)
+#'
 #' 
-#' Owl names correspond with 1:10 rows
+#' Owl names correspond with 1:10 rows (z) or 1:10 cols (w)
 species.names
+#' 
+w.init
 #' 
 z.init[4,3,,] # rows are different species, M2, 2005
 z.init[3,9,,] # rows are different species, M1, 2011
@@ -149,6 +161,7 @@ richness.jagsout <- jags(data = richness.jags.data,
                                              "spp.occ",
                                              "Nsmall",
                                              "richness",
+                                             "richness.RtYr",
                                              "beta.prebroad",
                                              "beta.pacific",
                                              "beta.mottled",
@@ -158,7 +171,8 @@ richness.jagsout <- jags(data = richness.jags.data,
                                              "beta.whiskered",
                                              "beta.gbarred",
                                              "beta.stygian",
-                                             "beta.ghorned"), 
+                                             "beta.ghorned",
[…]

### 2020-11-18 programs/a_process_data_global.R
update tables without "None" records

@@ -111,12 +111,19 @@ summary(tab.owls$Owl_Number)
 #' How many of each species of owl are there?
 table(tab.owls$Owl_Species_ID)
 
-#' Make "None" and "NoID" match
+#' Remove "None" owls - these are records that don't have associated owls IDed at those surveys
 #' 
-tab.owls$Owl_Species_ID <- ifelse(test = tab.owls$Owl_Species_ID == "None" |
-                                    tab.owls$Owl_Species_ID == "NoID",
-                                  yes = "Unk",
-                                  no = tab.owls$Owl_Species_ID)
+tab.owls <- tab.owls[tab.owls$Owl_Species_ID != "None",]
+
+#' There is one NoID owl with no positive records, remove
+#' 
+tab.owls[tab.owls$Owl_Species_ID == "NoID",1:7]
+tab.owls <- tab.owls[tab.owls$Owl_ID != 468,]
+
+#' Rename NoID as "Unk"
+#' 
+tab.owls$Owl_Species_ID[tab.owls$Owl_Species_ID=="NoID"] <- "Unk"
+#' 
 table(tab.owls$Owl_Species_ID)
 
 

### 2020-11-18 programs/c04_process_data_richness.R

@@ -93,7 +93,7 @@ n.species <- length(species.names)
 #' 
 #' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401)
 #' 
-n.aug <- 4 #number of species that may have been present but were undetected
+n.aug <- 5 #number of species that may have been present but were undetected
 n.species.aug <- n.species + n.aug
 
 
@@ -106,9 +106,9 @@ ks.prebroad <- ks.mottled <- ks.pacific <- ks.crested <- ks.bw <- ks.spectacled 
   ks.whiskered <- ks.gbarred <- ks.stygian <- ks.ghorned <- ks
 
 broadcast.array <- array(c(
-  rep(c("pacific", "mottled", "crested", "bw", "spectacled"),4), # routes EI1/EI2
-  rep(c("whiskered", "mottled", "gbarred", "stygian", "ghorned"),4), # routes M1/M2
-  rep(c("pacific", "mottled", "crested", "bw", "spectacled"),4)), # routes N1/N2
+  rep(c("PacSc", "Mottd", "Crested", "BW", "Specd"),4), # routes EI1/EI2
+  rep(c("Whisk", "Mottd", "Guat Barred", "Styg", "GrHor"),4), # routes M1/M2
+  rep(c("PacSc", "Mottd", "Crested", "BW", "Specd"),4)), # routes N1/N2
   dim = c(n.station, n.route))
 broadcast.array
 
@@ -117,23 +117,23 @@ for(hh in 1:n.route){
     ks.prebroad[hh,jj,1] <- 1
     ks.prebroad[hh,jj,2] <- 0
     
-    if(broadcast.array[jj,hh] == "pacific"){
+    if(broadcast.array[jj,hh] == "PacSc"){
       ks.pacific[hh,jj,2] <- 1
-    }else if(broadcast.array[jj,hh] == "mottled"){
+    }else if(broadcast.array[jj,hh] == "Mottd"){
       ks.mottled[hh,jj,2] <- 1
-    }else if(broadcast.array[jj,hh] == "crested"){
+    }else if(broadcast.array[jj,hh] == "Crested"){
       ks.crested[hh,jj,2] <- 1
-    }else if(broadcast.array[jj,hh] == "bw"){
+    }else if(broadcast.array[jj,hh] == "BW"){
       ks.bw[hh,jj,2] <- 1
[…]

### 2020-11-18 programs/e04_processing_results_richness.R

@@ -40,7 +40,7 @@ load(file = "data/processed_data/tables_global.Rdata")
 
 #' Species accounts to summarize by route
 #' 
-load("../data/output_data/richness_species_accounts.Rdata")
+load("data/output_data/richness_species_accounts.Rdata")
 
 #' Survey list
 (route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
@@ -82,14 +82,89 @@ for(hh in 1:nrow(species.accounts.byRt)){
   temp.route <- species.accounts.byRt$Route[hh]
   temp.accounts <- species.accounts[species.accounts$Route == temp.route,]
   for(ss in 1:10){ #across all 10 species
-    species.accounts.byRt[hh,ss] <- ifelse(sum(temp.accounts[,ss],na.rm = T)>0,1,0)
+    species.accounts.byRt[hh,ss] <- sum(temp.accounts[,ss],na.rm = T)
   }
 }
+
+(full.species.list <- c(colnames(species.accounts)[c(1,3:4,2,5:9)], 
+                       "Crested","Burrowing", 
+                       "BW","Striped", 
+                       "Saw-whet", "Unk"))
+
+sci.names.list <- c("Ciccaba virgata",#mottled
+                    "Glaucidium brasilianum",#ferpy
+                    "Pulsatrix perspicillata",#specd
+                    "Megascops cooperi", #pacsc
+                    "Tyto alba", #barn
+                    "Asio stygius", #stygian
+                    "M. trichopsis",#whiskered
+                    "Bubo virginianus",#grhor
+                    "Strix fulvescens",#fulvous
+                    "Lophostrix cristata",#crested
+                    "Athene cunicularia",#burrowing
+                    "Ciccaba nigrolineata", #bw
+                    "Pseudoscops clamator",#striped
+                    "Aegolius ridgwayi",#unspotted saw-whet
+                    "Unknown"
+                    )
+common.names.list <- c("Mottled",#mottled
+                       "Ferruginous Pygmy",#ferpy
+                       "Spectacled",#specd
+                       "Pacific Screech", #pacsc
+                       "Barn", #barn
+                       "Stygian", #stygian
+                       "Whiskered Screech",#whiskered
+                       "Great Horned",#grhor
+                       "Fulvous",#fulvous
+                       "Lophostrix cristata",#crested
+                       "Burrowing",#burrowing
[…]

### 2020-11-20 programs/b_richness_model_global.R
results from richness model with w(h,s) not w(h,t,s)

@@ -91,13 +91,15 @@ model.richness <- function(){
   for(hh in 1:n.route){ # 6 routes
     
 
-    
+    for(ss in 1:n.species.aug){
+      # Is each species present in that route's community?
+      w[hh,ss] ~ dbern(omega[hh])
+    }
     
     for(tt in 1:n.year){ # all years
       for(ss in 1:n.species.aug){ # all augmented species
         
-        # Is each species present in that route's community?
-        w[hh,ss] ~ dbern(omega[hh])
+        
         
         for(ii in 1:n.survey){ # 1 to 3 surveys per year
           

### 2020-11-20 programs/c04_process_data_richness.R

@@ -267,6 +267,13 @@ save(ks.prebroad, ks.pacific, ks.mottled, ks.crested, ks.bw, ks.spectacled,
   file = "data/processed_data/augmented_jags_ks.Rdata")
 
 
+#' Processed broadcast array
+#' 
+broadcast.array <- as.data.frame(broadcast.array)
+colnames(broadcast.array) <- c("EI1", "EI2", "M1", "M2", "N1", "N2")
+broadcast.array$Station <- 1:10
+save(broadcast.array, file = "data/output_data/broadcast_array.Rdata")
+
 #' _____________________________________________________________________________
 #' ### Footer
 #' 

### 2020-11-20 programs/d04_richness_analysis.R

@@ -159,7 +159,6 @@ richness.jagsout <- jags(data = richness.jags.data,
                                              "rho.psi",
                                              "omega",
                                              "spp.occ",
-                                             "Nsmall",
                                              "richness",
                                              "richness.RtYr",
                                              "beta.prebroad",
@@ -179,51 +178,41 @@ richness.jagsout <- jags(data = richness.jags.data,
                       n.burnin = 2000, #1000
                       n.thin = 2)
 
-
 #' _____________________________________________________________________________
-#' ## Post-process Psi = Probability of occupancy
+#' ## Psi = Probability of occupancy
 #' 
-#' Survey list
-(route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
-route.index <- 1:length(route.names)
-(year.names <- 2003:2013)
-year.index <- 1:length(year.names)
-#' ### Calculate median, 95% quantile and 5% quantile of psi posteriors by route
-#' and year
-#' 
-#' For all analysis, remove years and routes that were not surveyed
-(exclude.byrow <- c(3,4,9,15,19:24,27,33,40,44,56,63,64))
-all.rows <- 1:(length(route.index)*length(year.index))
-(include.byrow <- all.rows[!all.rows %in% exclude.byrow])
-#' 
-
+#' Create traceplots of parameters (creates PDF)
+#'
 #' Richness
+# MCMCtrace(richness.jagsout,
+#           pdf = TRUE,
+#           open_pdf = FALSE,
+#           filename = "richnessTrace",
+#           wd = "output")
+
+
+#' ## Posteriors for saving
 #' 
-richness.post <- MCMCsummary(richness.jagsout, 
-                             params = "richness", 
+#' Mu(psi)
+#' 
+mu.psi.post <- MCMCsummary(richness.jagsout, 
+                             params = "mu.psi", 
                              Rhat = TRUE,
                              n.eff = TRUE,
                              probs = c(0.05, 0.5, 0.95))
-richness.post$Year <- rep(year.names, each = length(route.index))
-richness.post$Route <- 
-  rep(c("EI1", "EI2", "M1", "M2", "N1", "N2"), length(year.index))
-richness.post <- richness.post[include.byrow,]
+mu.psi.chains <- MCMCchains(richness.jagsout, 
+                            params = "mu.psi")
 
-
-
-#' Populate Region
+#' Rho(psi)
 #' 
-richness.post$Region[richness.post$Route == "EI1" |
[…]

### 2020-11-20 programs/e04_processing_results_richness.R

@@ -31,8 +31,37 @@ set.seed(258854)
 #' ## Load Data
 #' 
 #' 
-#' Jagsout Richness Model
-#load(file = "data/output_data/richness_jagsout.Rdata")
+#' Jagsout Mottd
+#load(file = "data/output_data/richness_jagsout_posteriors.Rdata")
+
+#' Richness Posteriors 
+load(file = "data/output_data/richness_psi_posteriors.Rdata")
+load(file = "data/output_data/richness_psi_posteriors_RtYr.Rdata")
+
+#' Save posteriors
+#' 
+#' Mu(psi)
+load(file = "data/output_data/richness_mu_psi_posteriors.Rdata")
+
+#' Rho(psi)
+#' 
+load(file = "data/output_data/richness_rho_psi_posteriors.Rdata")
+
+#' Omega
+#' #' 
+load(file = "data/output_data/richness_omega_posteriors.Rdata")
+
+#' Spp.occ
+#' #' 
+load(file = "data/output_data/richness_spp_occ_posteriors.Rdata")
+
+#' Probability of detection by broadcast species and species of analysis
+#' 
+load(file = "data/output_data/richness_p_detection_posteriors.Rdata")
+
+#' Species present
+#' 
+load(file = "data/output_data/richness_species_present_posteriors.Rdata")
 
 #' Access tables, in Rdata format
 #' 
@@ -49,21 +78,116 @@ route.index <- 1:length(route.names)
 year.index <- 1:length(year.names)
 
 #' 
-#' There were 10 observed owls (if we pool all three unknown observations)
+#' There were 10 observed owls (if we include unknown observations)
 (species.names <- unique(tab.owls$Owl_Species_ID))
 n.species <- length(species.names)
 
 #' _____________________________________________________________________________
-#' ## Psi = Probability of occupancy
+#' ## Post-process Psi = Probability of occupancy
 #' 
-#' Create traceplots of parameters (creates PDF)
-#'
+#' Survey list
+(route.names <- c("EI1", "EI2", "M1", "M2",  "N1",  "N2") )
+route.index <- 1:length(route.names)
+(year.names <- 2003:2013)
+year.index <- 1:length(year.names)
+#' ### Calculate median, 95% quantile and 5% quantile of psi posteriors by route
+#' and year
+#' 
+#' For all analysis, remove years and routes that were not surveyed
[…]

### 2020-11-20 programs/g_graphing_results_global.R

@@ -143,7 +143,7 @@ ggplot(data = p.det.post.richness, aes(y = median.plogis, x = Broadcast))+
 #' 
 #' All 6 routes
 #+ richness_byRtYr
-ggplot(data = richness.post, 
+ggplot(data = richness.RtYr.post, 
        aes(x = Year, y = Richness.median, group = Route))+
   geom_pointrange(aes(ymin = Richness.LL05, ymax = Richness.UL95))+
   geom_line()+
@@ -159,7 +159,7 @@ ggplot(data = richness.post,
 #' El Imposible
 #' 
 #+ richness_ElImposible
-ggplot(data = richness.post[richness.post$Region == "El Imposible",], 
+ggplot(data = richness.RtYr.post[richness.RtYr.post$Region == "El Imposible",], 
        aes(x = Year, y = Richness.median))+
   geom_pointrange(aes(ymin = Richness.LL05, ymax = Richness.UL95))+
   geom_line()+
@@ -174,7 +174,7 @@ ggplot(data = richness.post[richness.post$Region == "El Imposible",],
 #' Nancuchiname
 #' 
 #+ richness_Nancuchiname
-ggplot(data = richness.post[richness.post$Region == "Nancuchiname",], 
+ggplot(data = richness.RtYr.post[richness.RtYr.post$Region == "Nancuchiname",], 
        aes(x = Year, y = Richness.median))+
   geom_pointrange(aes(ymin = Richness.LL05, ymax = Richness.UL95))+
   geom_line()+
@@ -189,7 +189,7 @@ ggplot(data = richness.post[richness.post$Region == "Nancuchiname",],
 #' Montecristo
 #' 
 #+ richness_Montecristo
-ggplot(data = richness.post[richness.post$Region == "Montecristo",], 
+ggplot(data = richness.RtYr.post[richness.RtYr.post$Region == "Montecristo",], 
        aes(x = Year, y = Richness.median))+
   geom_pointrange(aes(ymin = Richness.LL05, ymax = Richness.UL95))+
   geom_line()+

### 2020-11-20 manuscript_working/ms_working_RMD.Rmd
ready for MRE v2

@@
 We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
-For each species $s$ (where $s = 1,2, \ldots,15$), $w_{t,h,s}$ was a binary indicator of whether that owl species was present in each route and year:
+For each species $s$ (where $s = 1,2, \ldots,15$), $w_{h,s}$ was a binary indicator of whether that owl species was present in each route:
 
 $$
-w_{t,h,s} \sim \text{Bernoulli}(\Omega_h)
+w_{h,s} \sim \text{Bernoulli}(\Omega_h)
 $$
@@
 
-In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community that year ($w_{t,h,s}$):
+In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community that year ($w_{h,s}$):
 
 $$
-z_{t,h,i,s} \sim \text{Bernoulli}(w_{t,h,s}*\psi_{t,h,s})
+z_{t,h,i,s} \sim \text{Bernoulli}(w_{h,s}*\psi_{t,h,s})
 $$
@@
 
-Richness was conditional on each species' presence and was estimated as the sum of owl species present in each route and year:
+Richness of each route was conditional on each species' presence and was estimated as the sum of owl species present in that route:
 
 $$
-\text{Richness}_{t,h} = \sum_{s=1}^{14} w_{t,h,s} 
+\text{Richness}_{h} = \sum_{s=1}^{15} w_{h,s} 
+$$
+
+We also estimated year-to-year richness ($\text{Richness}^*_{t,h}$) by route by adding up the number of unique species that were present in each route, year, and survey.
+We first determined if a species was present in any of that route and year's set of surveys with a species-specific binary indicator $\vartheta_{t,h,s}$.
+Then, we summed up the number of species present in each year and route:
+$$
+\text{Richness}^*_{t,h} = \sum_{s=1}^{15} \vartheta_{t,h,s} 
 $$
@@
 Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl was detected in that survey and otherwise zero.
-Similarly, initial values for whether a species belonged to a route's community ($w_{t,h,s}$) were equivalent to one if that species of owl was detected in that route and year and otherwise zero.
+Similarly, initial values for whether a species belonged to a route's community ($w_{h,s}$) were equivalent to one if that species of owl was detected in that route and otherwise zero.
 We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
@@
 
-Median species richness varied from 1 to 5, depending on route and year (Fig.
+Our estimates of each route's community richness (Richness$_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
+These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those deteced.
+
+When analyzed by year, median species richness (Richness$^*_{t,h}$) varied from 0 to 4, depending on route and year (Fig.
 5XXX).
-At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species in several years of EI1 surveys and two years of M2 surveys (Fig.
+At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 3 new undetected species 2004 in M2 (Fig.
 5XXX).
-The shared information through random effects by route tended to smooth out the year-to-year variation in predicted richness, which in a few instances meant that the median predicted richness was less than the number of observed species in that route and year (Fig.
-5XXX)
+.
 
@@
 ![5XX.
-Species richness by route and year.
+Species richness (Richness*) by route and year.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2020-11-25 programs/a_process_data_global.R
ran richness model without "unknown" species

@@ -113,17 +113,10 @@ table(tab.owls$Owl_Species_ID)
 
 #' Remove "None" owls - these are records that don't have associated owls IDed at those surveys
 #' 
-tab.owls <- tab.owls[tab.owls$Owl_Species_ID != "None",]
+tab.owls <- tab.owls[tab.owls$Owl_Species_ID != "None" &
+                       tab.owls$Owl_Species_ID != "NoID",]
 
-#' There is one NoID owl with no positive records, remove
-#' 
-tab.owls[tab.owls$Owl_Species_ID == "NoID",1:7]
-tab.owls <- tab.owls[tab.owls$Owl_ID != 468,]
 
-#' Rename NoID as "Unk"
-#' 
-tab.owls$Owl_Species_ID[tab.owls$Owl_Species_ID=="NoID"] <- "Unk"
-#' 
 table(tab.owls$Owl_Species_ID)
 
 

### 2020-11-25 programs/c04_process_data_richness.R

@@ -91,7 +91,7 @@ n.species <- length(species.names)
 
 #' We should augment with M >> n, where n = 10
 #' 
-#' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401)
+#' Owls of El Salvador lists 13 total species recorded in El Salvador (Table 11.1, p401), so with Stygian, this should be 14
 #' 
 n.aug <- 5 #number of species that may have been present but were undetected
 n.species.aug <- n.species + n.aug
@@ -102,8 +102,8 @@ n.species.aug <- n.species + n.aug
 ks <-  array(0, dim = c(n.route, 
                         n.station, 
                         n.broadcast))
-ks.prebroad <- ks.mottled <- ks.pacific <- ks.crested <- ks.bw <- ks.spectacled <-
-  ks.whiskered <- ks.gbarred <- ks.stygian <- ks.ghorned <- ks
+ks.prebroad <- ks.mottled <- ks.pacific <- ks.crested <- ks.bw <- 
+  ks.spectacled <- ks.whiskered <- ks.gbarred <- ks.stygian <- ks.ghorned <- ks
 
 broadcast.array <- array(c(
   rep(c("PacSc", "Mottd", "Crested", "BW", "Specd"),4), # routes EI1/EI2

### 2020-11-25 programs/d04_richness_analysis.R

@@ -16,6 +16,7 @@ library(devtools) # documentation-related
 
 # analysis-related
 library(R2jags)
+library(MCMCvis)
 
 #' Clear environment and set seed
 #' 
@@ -292,7 +293,7 @@ species.present.chains <- MCMCchains(richness.jagsout,
 #+ savejags, eval = F
 #save(richness.post, file = "data/output_data/richness_jagsout_posteriors.Rdata")
 #' Richness Posteriors 
-# save(richness.post, richness.chains, 
+# save(richness.post, richness.chains,
 #      file = "data/output_data/richness_psi_posteriors.Rdata")
 # save(richness.RtYr.chains, richness.RtYr.post,
 #      file = "data/output_data/richness_psi_posteriors_RtYr.Rdata")
@@ -300,22 +301,22 @@ species.present.chains <- MCMCchains(richness.jagsout,
 #' Save posteriors
 #' 
 #' Mu(psi)
-# save(mu.psi.post, mu.psi.chains, 
+# save(mu.psi.post, mu.psi.chains,
 #      file = "data/output_data/richness_mu_psi_posteriors.Rdata")
 
 #' Rho(psi)
 #' 
-# save(rho.psi.chains, rho.psi.post, 
+# save(rho.psi.chains, rho.psi.post,
 #      file = "data/output_data/richness_rho_psi_posteriors.Rdata")
 
 #' Omega
 #' #' 
-#' save(omega.chains, omega.post, 
-#'      file = "data/output_data/richness_omega_posteriors.Rdata")
+# save(omega.chains, omega.post,
+#      file = "data/output_data/richness_omega_posteriors.Rdata")
 
 #' Spp.occ
 #' #' 
-# save(spp.occ.post, spp.occ.chains, 
+# save(spp.occ.post, spp.occ.chains,
 #      file = "data/output_data/richness_spp_occ_posteriors.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
@@ -325,15 +326,12 @@ species.present.chains <- MCMCchains(richness.jagsout,
 
 #' Species present
 #' 
-# save(species.present.chains, species.present.post, 
+# save(species.present.chains, species.present.post,
 #     file = "data/output_data/richness_species_present_posteriors.Rdata")
 
 
 
 
-#' Species Accounts
-#' 
-save(species.accounts, file = "data/output_data/richness_species_accounts.Rdata")
 
 
 #' _____________________________________________________________________________

### 2020-11-25 programs/e04_processing_results_richness.R

@@ -81,6 +81,7 @@ year.index <- 1:length(year.names)
 #' There were 10 observed owls (if we include unknown observations)
 (species.names <- unique(tab.owls$Owl_Species_ID))
 n.species <- length(species.names)
+n.species.aug <- n.species + 5
 
 #' _____________________________________________________________________________
 #' ## Post-process Psi = Probability of occupancy
@@ -90,6 +91,9 @@ n.species <- length(species.names)
 route.index <- 1:length(route.names)
 (year.names <- 2003:2013)
 year.index <- 1:length(year.names)
+n.route <- length(route.index)
+
+
 #' ### Calculate median, 95% quantile and 5% quantile of psi posteriors by route
 #' and year
 #' 
@@ -205,7 +209,7 @@ species.accounts.byRt$Region <- rep(unique(species.accounts$Region),each=2)
 for(hh in 1:nrow(species.accounts.byRt)){
   temp.route <- species.accounts.byRt$Route[hh]
   temp.accounts <- species.accounts[species.accounts$Route == temp.route,]
-  for(ss in 1:10){ #across all 10 species
+  for(ss in 1:n.species){ #across all 9 species
     species.accounts.byRt[hh,ss] <- sum(temp.accounts[,ss],na.rm = T)
   }
 }
@@ -213,7 +217,7 @@ for(hh in 1:nrow(species.accounts.byRt)){
 (full.species.list <- c(colnames(species.accounts)[c(1,3:4,2,5:9)], 
                        "Crested","Burrowing", 
                        "BW","Striped", 
-                       "Saw-whet", "Unk"))
+                       "Saw-whet"))
 
 sci.names.list <- c("Ciccaba virgata",#mottled
                     "Glaucidium brasilianum",#ferpy
@@ -228,8 +232,7 @@ sci.names.list <- c("Ciccaba virgata",#mottled
                     "Athene cunicularia",#burrowing
                     "Ciccaba nigrolineata", #bw
                     "Pseudoscops clamator",#striped
-                    "Aegolius ridgwayi",#unspotted saw-whet
-                    "Unknown"
+                    "Aegolius ridgwayi"#unspotted saw-whet
                     )
 common.names.list <- c("Mottled",#mottled
                        "Ferruginous Pygmy",#ferpy
@@ -244,13 +247,12 @@ common.names.list <- c("Mottled",#mottled
                        "Burrowing",#burrowing
[…]

### 2020-11-25 manuscript_working/ms_working_RMD.Rmd
incorporated MRE edits, sending to Jane

@@
 We modeled occupancy within each route, year, and survey assuming that the probability of occupancy would be closed across surveys of a given route in a given year.
-In other words, for any single species of owl, we assumed that the probability of a route being occupied by that owl species would not change between surveys within a year but that the probability of occupancy would vary between years.
+In other words, for any single species of owl, we assumed that the probability of a route being occupied would not change between surveys within a year but that the probability of occupancy would vary between years.
 Let:
@@
 \begin{aligned}
-t  &=  \text{the year each survey was conducted (here }t = 1,2,\ldots,11)\text{ corresponding with 2003-2013}\\
-h  &=  \text{the individual route (here }h = 1,2,\ldots,6)\text{ corresponding with EI1, EI2, M1, M2, N1, N2}\\
-i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
-j  &=  \text{the individual stations along each route (here }j = 1,2,\ldots,10)\\
-k  &=  \text{the broadcast call species (here }k = 1,2,\ldots,10)\\
+t  &=  \text{the year each survey was conducted (}t = 1,2,\ldots,11)\text{ representing 2003-2013}\\
+h  &=  \text{the individual route (}h = 1,2,\ldots,6)\text{ representing EI1, EI2, M1, M2, N1, N2}\\
+i  &=  \text{the individual surveys conducted in each year and route (}i = 1,2,3)\\
+j  &=  \text{the individual stations along each route (}j = 1,2,\ldots,10)\\
+k  &=  \text{the broadcast call species (}k = 1,2,\ldots,10)\\
 z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
@@
 
-This model provides a consistent probability of detection for all surveys in the first two minutes of each observation period, before the broadcast call was played ($k = \text{Pre-broadcast}$).
+This model provided a consistent probability of detection for all surveys in the first two minutes of each observation period, before the broadcast call was played ($k = \text{Pre-broadcast}$).
 Then, the probability of detection for each post-broadcast time period depended which call had been played.
@@
 
-We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the occupancy model for three owl species: Mottled, Spectacled, and Ferruginous Pygmy owls.
-These species had sufficient positive detections to analyze occupancy (Table 1XX).
+We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the occupancy model for three owl species that had sufficient positive detections for analysis: Mottled, Spectacled, and Ferruginous Pygmy owls (Table 1XX).
 Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
@@
 
-To model species richness at each route in each year, we assumed that there were 15 possible species present in El Salvador.
-This included the 10 species observed during our surveys (including unknown owl species) and an additional 5 species that we never observed during our surveys (Table 1XX).
-This upper limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus our unknown species and one species we identified (Stygian owl), which was previously undocumented in El Salvador.
+To model species richness at each route in each year, we assumed that there were 14 possible species present in El Salvador.
+This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 1XX).
+This upper limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus one species we identified (Stygian owl), which was previously undocumented in El Salvador.
 We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
@@
 where $\Omega_h$ was the probability of species presence in each route $h$ [@Broms:2016].
-We used prior recommended by @Broms:2016, @Guillera-Arroita:2019, and others for $\Omega_h$: $\Omega_h \sim \text{Beta}(0.001, 1)$.
+We used prior distributions recommended by @Broms:2016, @Guillera-Arroita:2019, and others for $\Omega_h$: $\Omega_h \sim \text{Beta}(0.001, 1)$.
 
-In the richness model, the presence of an owl species in any route, year, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community that year ($w_{h,s}$):
+In the richness model, the presence of an owl species in any year, route, and survey, $z_{t,h,i,s}$, was conditional on the probability of occupancy for that route, year, and species $\psi_{t,h,s}$ *and* that species belonging to that route's community that year ($w_{h,s}$):
 
@@
 b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
-\text{logit}(\mu^\psi_{t,h}) &= \sim N(0,2.25)\\
+\text{logit}(\mu^\psi_{t,h}) &\sim N(0,2.25)\\
 \ln(\rho^\psi) &\sim N(5,1)\text{, truncated to (0.01,10)}\\
@@
 Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl was detected in that survey and otherwise zero.
-Similarly, initial values for whether a species belonged to a route's community ($w_{h,s}$) were equivalent to one if that species of owl was detected in that route and otherwise zero.
+Similarly, initial values for whether a species belonged to a route's community ($w_{h,s}$) were equivalent to one if that species of owl was ever detected in that route and otherwise zero.
 We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
@@
 1X).
-The probability of occupancy across all three species and both routes of Nancuchiname were relatively high (Fig.
+Occupancy probabilities were relatively high across all three species for both routes of Nancuchiname (Fig.
 1XX).
-We found that the probability of occupancy for Mottled owls was relatively consistent across time in El Imposible and in N2 route (Fig.
-2XX), but that it seemed to increase in the last ~6 years of the surveys in M2 and N1 routes (Fig.
+We found that the probability of occupancy for Mottled owls was relatively consistent across time in the El Imposible routes and in N2 (Fig.
+2XX), but that it seemed to increase in the last ~6 years in M2 and N1 (Fig.
 2XX).
@@
 The probability of occupancy for Ferruginous Pygmy owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
-3XX), and the probability of occupancy for the Spectacled owl slightly fluctuated year to year in the Nancuchaname routes (Fig.
+3XX), and the probability of occupancy for the Spectacled owl  fluctuated slightly from year to year in the Nancuchaname routes (Fig.
 4XX).
@@
 
-Our estimates of each route's community richness (Richness$_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
-These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those deteced.
+Our estimates of each route's community richness ($\text{Richness}^*_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
+These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
 
-When analyzed by year, median species richness (Richness$^*_{t,h}$) varied from 0 to 4, depending on route and year (Fig.
+When analyzed by year, median species richness ($\text{Richness}^*_{h}$) varied from 0 to 4, depending on route and year (Fig.
 5XXX).
-At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 3 new undetected species 2004 in M2 (Fig.
+At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 3 new undetected species in M2 during 2004 surveys (Fig.
 5XXX).
-.
 
@@
 
-The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by broadcast call and between Mottled, Ferruginous Pygmy, and Spectacled owls (Fig.
+The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled, Ferruginous Pygmy, and Spectacled owls (Fig.
 6XX).

### 2021-01-03 manuscript_working/ms_working_RMD.Rmd
new introduction draft

@@
 --- 
-title: "working owls methods"
-author: "Althea Archer"
-date: "11/17/2020"
+title: "Owl occupancy in three of El Salvador's protected areas from 2003 to 2013"
+author: "Jane West and Althea Archer"
+date: "1/2/2021"
 output:
@@
 ```
+
+# Abstract
+
+# Introduction
+
+Owls are a top predator in various habitats and as such, their role as a keystone species makes them an important, yet cryptic, bioindicator of ecosystem health.
+Top predators also often have long-lasting and wide-ranging effects through trophic cascades.
+The spotted owl species of North America have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owl have complex and contrasting behavior responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
+
+Neotropical owls are less well-studied than temperate species, such as the spotted owls of North America, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017].
+Yet, neotropical owls are reliant upon ecosystems under tremendous anthropogenic pressure and rapid and unpredictable responses to climate change [@Corlett:2012].
+In the tropics, particularly in rain and cloud forests, owl studies are more difficult than in temperate regions, although there are more species living in the same habitat [@Konig:1999].
+In general, while our knowledge about the conservation status of neotropical populations is limited, we do know that owl abundance and distribution is decreasing in several regions, where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
+
+The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration of plumage as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
+Owls vocalize to communicate with the same species and to delimit territory [@Johnsgard:1988].
+Spontaneous and provocation listening are two methods for auditory surveys that have been used with increasing frequency in surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
+Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls" that were previously recorded.
+One benefit of auditory surveys (including both spontaneous and broadcast) is being able to identify multiple species simultaneously.
+Adding in provocation to auditory surveys can increase the detection probability of owls, although the intra- and interspecific responses to broadcast calls could play a role in response patterns [@Enriquez:1997;@Baumgardt:2019].
+
+The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and species community composition.
+We surveyed owls on foot using passive listening and broadcast vocalizations in two transects each within the three largest protected areas of El Salvador from 2003 through 2013.
 

### 2021-01-03 manuscript_working/ms_working_RMD.Rmd
updated tables

@@
 
-Owls are a top predator in various habitats and as such, their role as a keystone species makes them an important, yet cryptic, bioindicator of ecosystem health.
+Owls, as a keystone species, are an important, yet cryptic, bioindicator of ecosystem health.
 Top predators also often have long-lasting and wide-ranging effects through trophic cascades.
@@
 Spontaneous and provocation listening are two methods for auditory surveys that have been used with increasing frequency in surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
-Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls" that were previously recorded.
-One benefit of auditory surveys (including both spontaneous and broadcast) is being able to identify multiple species simultaneously.
+Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls." One benefit of auditory surveys (including both spontaneous and broadcast) is being able to identify multiple species simultaneously.
 Adding in provocation to auditory surveys can increase the detection probability of owls, although the intra- and interspecific responses to broadcast calls could play a role in response patterns [@Enriquez:1997;@Baumgardt:2019].
@@
 
-*Table 1XX. Species detection records by route. Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route.*
+*Table 1XX. Route locations and habitats.*
+
+```{r}
+library(pander)
+Habitat <- c("Deciduous Forest with Reforestation",
+             "Semi-Deciduous Forest",
+             "Cloud Forest",
+             "Pine/Oak Forest",
+             "Alluvial Forest with Secondary Growth",
+             "Alluvial Forest")
+Latitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
+Longitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
+Route <- c("EI-1", "EI-2", "M-1", "M-2", "N-1", "N-2")
+Area <- c("El Imposible", "NP", 
+          "Montecristo", "NP", "Nancuchiname", "Forest")
+#parkTable <- as.data.frame(matrix(
+#  c(NamePark, NameRoute, Latitude, Longitude, Habitat), byrow = T, nrow = 5, ncol = 7))
+parkTable <- data.frame(Area, Route, Latitude, Longitude, Habitat)
+pander(parkTable)
+
+```
+
+*Table 2XX. Species detection records by route. Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route.*
 
@@
 library(pander)
-pander(species.counts.long[c(3:9)])
+updated <- species.counts.long[,c(3:9)]
+updated$Common[updated$Common == "Lophostrix cristata"] <- "Crested"
+scientific <- species.counts.long[,c(2,4:9)]
+scientific[,c(2:7)] <- " "
+colnames(updated) <- names(scientific) <- c("NameBoth", "EI-1", "EI-2",
+                                            "M-1", "M-2", "N-1", "N-2")
+
+combo <- rbind(updated, scientific)
+
+combo$Name <- factor(x = combo$NameBoth,
+                           levels = c("Barn", "Tyto alba",
+                                      "Whiskered Screech", "M. trichopsis",
+                                      "Pacific Screech", "Megascops cooperi",
+                                      "Crested", "Lophostrix cristata",
+                                      "Spectacled", "Pulsatrix perspicillata",
+                                      "Great Horned", "Bubo virginianus",
+                                      "Ferruginous Pygmy", "Glaucidium brasilianum",
+                                      "Burrowing", "Athene cunicularia",
+                                      "Mottled", "Ciccaba virgata",
+                                      "Black-and-white", "Ciccaba nigrolineata",
+                                      "Fulvous", "Strix fulvescens",
+                                      "Stygian", "Asio stygius",
+                                      "Striped", "Pseudoscops clamator",
+                                      "Unspotted Saw-whet", "Aegolius ridgwayi"),
+                           labels = c("Barn Owl", "(Tyto alba)",
+                                      "Whiskered Screech-Owl", "(Megascops trichopsis)",
+                                      "Pacific Screech-Owl", "(Megascops cooperi)",
+                                      "Crested Owl", "(Lophostrix cristata)",
+                                      "Spectacled Owl**", "(Pulsatrix perspicillata)",
+                                      "Great Horned Owl", "(Bubo virginianus)",
+                                      "Ferruginous Pygmy-Owl", "(Glaucidium brasilianum)",
+                                      "Burrowing Owl", "(Athene cunicularia)",
+                                      "Mottled Owl", "(Ciccaba virgata)",
+                                      "Black-and-white Owl**", "(Ciccaba nigrolineata)",
+                                      "Fulvous Owl*", "(Strix fulvescens)",
+                                      "c.f. Stygian Owl", "(Asio stygius)",
+                                      "Striped Owl", "(Pseudoscops clamator)",
+                                      "Unspotted Saw-whet Owl", "(Aegolius ridgwayi)"
+                           ))
+
+row.names(combo) <- NULL
+final <- combo[order(combo$Name),c(8,2:7)]
+row.names(final) <- NULL
+pander(final)
 ```
+*Note: \*\* = Endangered and \* = Threatened [@MARN:2015]*
 
@@
 
-The broadcast calls varied between routes, but were consistent at each route and station across years (Table 2XX).  Thus, the effective sample sizes of the different broadcast calls varied.
-
-*Table 2XX.
-Broadcast calls by route and station.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
-Pre-recorded owl calls were broadcast after 2 minutes of silent listening.*
-
-``` {r, echo = F}
-load(file = "../data/output_data/broadcast_array.Rdata")
-pander(broadcast.array)
-```
+The broadcast calls varied between routes, but were consistent at each route and station across years.
+Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible (EI-1 and EI-2) and Nancuchiname (N-1) and N-2).
+We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo (M-1 and M-2).
+Thus, the effective sample sizes of the different broadcast calls varied.
 

### 2021-01-03 manuscript_working/ms_working_RMD.Rmd
updated results

@@
 
+This project was carried out during the breeding season (March-May) for three weeks each year over a 10-year period in the three previously mentioned protected natural areas.
+These areas were selected to represent different ecosystems found in the country and the transect routes were selected to represent the habitats in the protected natural areas.
+
 # Methods
+
+## Study Area
+
+El Salvador is the smallest and most densely populated (approximately 6.5 million) Central American country with 13.6% of its land forested [@CIA:2020].
+It is bordered by the Pacific Ocean, Guatemala and Honduras (Fig.
+1XXX).
+El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot (Conservation International 2009 in Komar and Ibarra-Portillo 2009).
+This project includes three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that are representative of several of the country’s protected natural areas (Table 1XXX).
+
+El Imposible National Park (NP) is located 119 km southwest of San Salvador (Fig.
+1XXX, Table 1XXX).
+Its elevation ranges from 250 to 1425 m above sea level.
+It is the largest NP in the country, covering 3792 ha.
+The topography is extremely steep and broken, with many cliffs [@Alverez:2003].
+The area contains deciduous and semi-deciduous forest, secondary growth vegetation on the edges and former pasture areas that were reforested with native species in 1997.
+In the western portion of this NP, one survey transect (EI-1) was set up in the lower elevation deciduous in an area of reforestation and the other (EI-2) was located in semi-deciduous forest.
+
+Montecristo NP is located 125 km northwest of San Salvador (Fig.
+1XXX; Table 1XXX).
+Its altitude ranges from 730 to 2418 m above sea level and covers 1973 ha [@MARN:2010].
+The higher elevations contain cloud forest, middle elevations contain pine/oak forest (with some pine and cypress plantations) and the lower elevations contain a semi-deciduous tropical forest [@MARN:2015b].
+This NP is part of the Montecristo Tri-national Protected Area which includes extensive adjoining natural areas in Guatemala and Honduras [@Komar:2010].
+In this NP, one transect (M-1) was set up in a cloud forest and another (M-2) was located at a lower elevation in a pine/oak forest.
+
+Nancuchiname Forest, a Protected Natural Area, is located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
+1XXX).
+Its elevation ranges from 5 to 12 m above sea level and covers 797 ha.
+Nancuchiname Forest represents the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001].
+The area contains deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike.
+In Nancuchiname Forest, one survey transect (N-1) was set up in alluvial forest along a dike surrounded by second growth vegetation and another (N-2) in a semi-deciduous alluvial forest.
 
@@
 library(pander)
-Habitat <- c("Deciduous Forest with Reforestation",
+Habitat <- c("Reforested Deciduous Forest",
              "Semi-Deciduous Forest",
@@
 Longitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
+Elevation <- c("xxx", "xxx", "xxx", "xxx", "xxx", "xxx")
 Route <- c("EI-1", "EI-2", "M-1", "M-2", "N-1", "N-2")
@@
 #  c(NamePark, NameRoute, Latitude, Longitude, Habitat), byrow = T, nrow = 5, ncol = 7))
-parkTable <- data.frame(Area, Route, Latitude, Longitude, Habitat)
+parkTable <- data.frame(Area, Route, Latitude, Longitude, Elevation, Habitat)
 pander(parkTable)
@@
 ```
+
+## Survey Methods
+
+We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001] and a similar protocol developed for Costa Rica [@Enriquez:2001].
+This study's survey protocol was an extension of previous research completed in the western half of El Imposible NP [@West:1988].
+For our surveys, we created six 2-km long survey routes (transects) that each included 10 permanent survey stations at 200 m intervals.
+Two routes were established in each protected area (Table 1xxx).
+
+Surveys were begun at local twilight and took approximately five hours to complete.
+At each station, we measured and recorded environmental conditions (Brunton Sherpa, Brunton Incorporated) including temperature, precipitation, cloud cover, fog cover, wind speed, barometric pressure, moon phase, and noise level.
+The survey then began with passive listening for two minutes, followed by a three-minute broadcast call, and then seven minutes of silent listening.
+We recorded local owl vocalizations and used them for broadcast calls whenever possible.
+If an owl vocalized during broadcast playback, we stopped the broadcast and recorded the owl's vocalization.
+When we detected an owl, we noted its  location and angle relative to the station, and noted when the same owl was identified at consecutive stations.
+The broadcast calls varied between routes, but were consistent at each route and station across years.
+Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible (EI-1 and EI-2) and Nancuchiname (N-1) and N-2).
+We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo (M-1 and M-2).
+Surveys were repeated at each route up to three times a year from 2003 through 2013, depending on site access and weather.
+We did not conduct any surveys in 2006.
+We georeferenced the owl detections for each survey and route  to eliminate double counting of the owl observations.
 
@@
 We modeled occupancy within each route, year, and survey assuming that the probability of occupancy would be closed across surveys of a given route in a given year.
-In other words, for any single species of owl, we assumed that the probability of a route being occupied would not change between surveys within a year but that the probability of occupancy would vary between years.
+In other words, for any single species of owl, we assumed that the probability of a route being occupied would not change between surveys within a year but could vary between years.
 Let:
@@
 k  &=  \text{the broadcast call species (}k = 1,2,\ldots,10)\\
-z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
+z_{t,h,i} &= \text{a random variable equal to 1 when a survey route was occupied and 0 otherwise}\\
 y_{t,h,i,j,k} &= \text{a random variable equal to 1 when an owl was detected and 0 otherwise}
@@
 
-We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, which was related to the species-specific broadcast call ($k$):
+We assumed that detecting an owl depended on an owl being present during that survey ($z_{t,h,i} = 1$) and the probability of detection, $p_{t,h,i,j,k}$, which was related to the species-specific broadcast call ($k$):
 
@@
 
-The broadcast calls varied between routes, but were consistent at each route and station across years.
-Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible (EI-1 and EI-2) and Nancuchiname (N-1) and N-2).
-We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo (M-1 and M-2).
-Thus, the effective sample sizes of the different broadcast calls varied.
-
-### Single-Species Occupancy Model Implementation
-
-We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the occupancy model for three owl species that had sufficient positive detections for analysis: Mottled, Spectacled, and Ferruginous Pygmy owls (Table 1XX).
-Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
-
-For all three species' occupancy models, we ran 3 chains for 10,000 iterations and 1000 iterations discarded as burn-in, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
-Initial values for $z_{t,h,i}$ were equivalent to one if an owl was detected in that survey $i$ and otherwise zero.
-We visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
-
 ## Richness Model Framework
@@
 We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
-For each species $s$ (where $s = 1,2, \ldots,15$), $w_{h,s}$ was a binary indicator of whether that owl species was present in each route:
+For each species $s$ (where $s = 1,2, \ldots,15$), $w_{h,s}$ was a binary indicator of whether that owl species was present in each route based on the probability of species presence in that route, $\Omega_h$ [@Broms:2016]:
 
@@
 $$
-where $\Omega_h$ was the probability of species presence in each route $h$ [@Broms:2016].
-We used prior distributions recommended by @Broms:2016, @Guillera-Arroita:2019, and others for $\Omega_h$: $\Omega_h \sim \text{Beta}(0.001, 1)$.
+We used prior distributions for $\Omega_h$: $\Omega_h \sim \text{Beta}(0.001, 1)$ [@Broms:2016; @Guillera-Arroita:2019].
 
@@
 
-### Richness Model Implementation
+### Model Implementation
 
-We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the richness model.
-We ran 3 chains for 20,000 iterations with 2000 iterations discarded as burn-in and a thinning rate of 2, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the Bayesian models.
+We verified that R-hats were lower than XXXX and visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
+
+We applied the single-species occupancy model for three owl species that had sufficient positive detections for analysis: Mottled Owl, Spectacled Owl, and Ferruginous Pygmy-Owl (Table 2XX).
+Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
+For all three of the single-species occupancy models, we ran 3 chains for 10,000 iterations and 1000 iterations discarded as burn-in, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
+Initial values for $z_{t,h,i}$ were equivalent to one if an owl was detected in that survey $i$ and otherwise zero.
+
+To implement the richness model, we ran 3 chains for 20,000 iterations with 2000 iterations discarded as burn-in and a thinning rate of 2, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
 Initial values for $z_{t,h,i,s}$ were equivalent to one if that species of owl was detected in that survey and otherwise zero.
 Similarly, initial values for whether a species belonged to a route's community ($w_{h,s}$) were equivalent to one if that species of owl was ever detected in that route and otherwise zero.
-We visually inspected traceplots to verify that chains mixed well (Supplemental information S4xx).
 
@@
 
-Mottled, Spectacled, and Ferruginous Pygmy owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 1XX).
-No other owl species had enough detections to analyze species-specific occupancy (Table 1XX).
+We detected nine species of owls during our surveys: Barn Owl, Whiskered Screech-Owl, Pacific Screech-Owl, Spectacled Owl, Great Horned Owl, Ferruginous Pygmy-Owl, Mottled Owl, Fulvous Owl, and c.f.
+Stygian Owl (Table 2xXX).
+We identified Whiskered Screech-Owlsin Montecristo NP, which had previously been undocumented in that protected area.
+We also identified a c.f.
+Stygian Owl in Montecristo NP, which was a species not previously documented in the entire country of El Salvador.
+
+Mottled Owls, Spectacled Owls, and Ferruginous Pygmy owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 2XX).
+No other owl species had enough detections to analyze species-specific occupancy (Table 2XX).
 Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for Ferruginous Pygmy or Spectacled owls in both routes of El Imposible and in M2 route of Montecristo (Fig.
@@
 
-Our estimates of each route's community richness ($\text{Richness}^*_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
+Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
 These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
 
-When analyzed by year, median species richness ($\text{Richness}^*_{h}$) varied from 0 to 4, depending on route and year (Fig.
+When analyzed by year, median species richness ($\text{Richness}^*_{t,h}$) varied from 0 to 5, depending on route and year (Fig.
 5XXX).
-At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 3 new undetected species in M2 during 2004 surveys (Fig.
+At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species (e.g., in M-2 during 2004 where we detected zero species (Fig.
 5XXX).

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
discussion, theme, and updated citations

@@
     fig_caption: yes
+    reference_docx: theme.docx
   pdf_document:
@@
 Top predators also often have long-lasting and wide-ranging effects through trophic cascades.
-The spotted owl species of North America have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owl have complex and contrasting behavior responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
+The various spotted owl sub-species (*Strix occidentalis*) of North America have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owl have complex and contrasting behavior responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
 
@@
 The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration of plumage as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
-Owls vocalize to communicate with the same species and to delimit territory [@Johnsgard:1988].
+*springer and forsman not cited on doc from Jane xxx* Owls vocalize to communicate with the same species and to delimit territory [@Johnsgard:1988].
 Spontaneous and provocation listening are two methods for auditory surveys that have been used with increasing frequency in surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
@@
 The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and species community composition.
+El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot (Conservation International 2009 in Komar and Ibarra-Portillo 2009).
 We surveyed owls on foot using passive listening and broadcast vocalizations in two transects each within the three largest protected areas of El Salvador from 2003 through 2013.
@@
 
-## Study Area
+## Study area
 
@@
 1XXX).
-El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot (Conservation International 2009 in Komar and Ibarra-Portillo 2009).
 This project includes three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that are representative of several of the country’s protected natural areas (Table 1XXX).
@@
 It is the largest NP in the country, covering 3792 ha.
-The topography is extremely steep and broken, with many cliffs [@Alverez:2003].
+The topography is extremely steep and broken, with many cliffs [@Alvarez:2003].
 The area contains deciduous and semi-deciduous forest, secondary growth vegetation on the edges and former pasture areas that were reforested with native species in 1997.
@@
 Its elevation ranges from 5 to 12 m above sea level and covers 797 ha.
-Nancuchiname Forest represents the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001].
+Nancuchiname Forest represents the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001]xxx **we can't site a geocities website**.
 The area contains deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike.
@@
 
-## Survey Methods
+## Survey methods
 
@@
 
-Surveys were begun at local twilight and took approximately five hours to complete.
+Surveys began at local twilight and took approximately five hours to complete.
 At each station, we measured and recorded environmental conditions (Brunton Sherpa, Brunton Incorporated) including temperature, precipitation, cloud cover, fog cover, wind speed, barometric pressure, moon phase, and noise level.
@@
 If an owl vocalized during broadcast playback, we stopped the broadcast and recorded the owl's vocalization.
-When we detected an owl, we noted its  location and angle relative to the station, and noted when the same owl was identified at consecutive stations.
+When we detected an owl, we noted its  location and angle relative to the station, and we tracked when the same owl was identified at consecutive stations.
 The broadcast calls varied between routes, but were consistent at each route and station across years.
@@
 
-## Single-Species Occupancy Model Framework
+## Single-species occupancy model framework
 
@@
 t  &=  \text{the year each survey was conducted (}t = 1,2,\ldots,11)\text{ representing 2003-2013}\\
-h  &=  \text{the individual route (}h = 1,2,\ldots,6)\text{ representing EI1, EI2, M1, M2, N1, N2}\\
+h  &=  \text{the individual route (}h = 1,2,\ldots,6)\text{ representing EI-1, EI-2, M-1, M-2, N-1, N-2}\\
 i  &=  \text{the individual surveys conducted in each year and route (}i = 1,2,3)\\
@@
 
-## Richness Model Framework
+## Richness model framework
 
@@
 
-### Model Implementation
+### Model implementation
 
@@
 We applied the single-species occupancy model for three owl species that had sufficient positive detections for analysis: Mottled Owl, Spectacled Owl, and Ferruginous Pygmy-Owl (Table 2XX).
-Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M1 in Montecristo, so we removed M1 from those species' occupancy analyses.
+Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M-1 in Montecristo, so we removed M-1 from those species' occupancy analyses.
 For all three of the single-species occupancy models, we ran 3 chains for 10,000 iterations and 1000 iterations discarded as burn-in, for a total of 27,000 iterations comprising the posterior distributions for each model parameter.
@@
 
-## Single-Species Occupancy Model Results
+A total of 86 surveys were conducted between March and May from 2003 through 2005 and 2007 through 2013. No surveys were conducted in 2006.
 
@@
 Stygian Owl (Table 2xXX).
-We identified Whiskered Screech-Owlsin Montecristo NP, which had previously been undocumented in that protected area.
+We identified Whiskered Screech-Owls in Montecristo NP, which had previously been undocumented in that protected area.
 We also identified a c.f.
 Stygian Owl in Montecristo NP, which was a species not previously documented in the entire country of El Salvador.
+The c.f.
+Stygian Owl requires further investigation to determine true presence of that owl in El Salvador.
 
-Mottled Owls, Spectacled Owls, and Ferruginous Pygmy owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 2XX).
+## Single-species occupancy model results
+
+Mottled Owls, Spectacled Owls, and Ferruginous Pygmy-Owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 2XX).
 No other owl species had enough detections to analyze species-specific occupancy (Table 2XX).
-Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for Ferruginous Pygmy or Spectacled owls in both routes of El Imposible and in M2 route of Montecristo (Fig.
+Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for Ferruginous Pygmy or Spectacled owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
 1X).
@@
 1XX).
-We found that the probability of occupancy for Mottled owls was relatively consistent across time in the El Imposible routes and in N2 (Fig.
-2XX), but that it seemed to increase in the last ~6 years in M2 and N1 (Fig.
+We found that the probability of occupancy for Mottled owls was relatively consistent across time in the El Imposible routes and in N-2 (Fig.
+2XX), but that it seemed to increase in the last ~6 years in M-2 and N-1 (Fig.
 2XX).
-The probability of occupancy in M1 seemed to have decreased through time for Mottled owls (Fig.
+The probability of occupancy in M-1 seemed to have decreased through time for Mottled owls (Fig.
 2XX).
 
-For routes EI1, EI2, and M2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI1 in the early years of the surveys (Figs.
+For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI-1 in the early years of the surveys (Figs.
 3XX and 4XX).
@@
 
-## Richness Model Results
+## Richness model results
 
-Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI1, EI2, M1, M2, N1, and N2, respectively.
+Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
 These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
@@
 
-## Detection Results
+## Detection results
 
@@
 6XX).
+
+# Discussion
+
+The goal of this long-term project was to complete two survey passes at each route for each year from 2003 through 2013.
+However, that was not always possible due to weather conditions, such as strong winds or rain.
+Access to the routes also precluded us from surveys in some instances.
+Rain and high winds diminish the vocalizations of owls as well as the ability for surveyors to detect any owl calls [@Takats:2001; @Andersen:2007].
+The survey was initially set up for 10 years based on recommendations from Mexican Spotted Owls researchers who stated that a monitoring period of 10 years would provide adequate information about population trends [@USFWS:1995].
+
+In Central America, the Mottled Owl is the most common and widespread owl [@Vallely:2018].
+@Perez-Leon:2017 stated that habitat loss has restricted the distribution of several owl species, but that populations of Barn Owls, Pacific Screech-owls, Mottled Owls, and Ferruginous Pygmy-owls have survived the fragmentation processes in El Salvador relative to other Neotropical species.
+Contrarily, @Enriquez:2001 stated that Mottled Owl populations in La Selva may have decreased during the past 30 years.
+The results of this study indicated that Mottled Owls are still fairly common at each of the protected areas, although there seemed to be a downward trend in occupancy in M-2 (Fig xXX).
+
+Ferruginous Pygmy-owls were the second most-often detected owls in this study.
+Ferruginous Pygmy-owls were often detected at the beginning of survey routes, often closer to sunset than later at night, and Ferruginous Pygmy-owls demonstrated fairly low occupancy rates in El Imposible routes, other than in the first couple years of the survey in EI-1.
+Ferruginous Pygmy-owls prefer open habitats with a lot of edge.
+@West:1988 observed that the lower elevation areas of El Imposible had been grazed and dotted with houses.
+The activities and vegetation structure of the area has changed overtime, and native species re-vegetation efforts began in that area in 1997.
+It is possible that Ferruginous Pygmy-owls are responding to the decrease in edge habitat in this park, and we recommend further surveys to determine the risk to Ferruginous Pygmy-owls in response to revegetation management.
+
+Stygian Owl was listed as an expected owl in El Salvador [@Perez-Leon:2017).
+On 26 March 2002, after listening to recordings and viewing photographs, the Montecristo NP guards stated that they thought c.f.
+Stygian owl was present in the park (personal communication).
+While there was only one c.f.
+Stygian Owl detected during the course of our surveys, we did also hear a c.f.
+Stygian Owl between the survey points on 25 April XXXXyear.
+We also heard c.f.
+Stygian Owl twice on M-2 between station points on 20 March 2005 and once between stations in 2007.
+We did not record any of these vocalizations by c.f.
+Stygian Owls.
+Detections that occurred between stations or away from survey routes were not included in our results.
+So while we are confident that c.f.
+Stygian Owls are present in Montecristo NP, we believe that further documentation of the species would be beneficial for confirmation of the species in El Salvador.
+
+During 1979-1981, in the western portion of El Imposible NP, West (1988) heard or observed six owl species, including two that were not detected during this study (Barn Owls and Black-and-white Owls) in that NP.
+In 1980, Barn Owls nested in a cave of Loma de Paja Mountain (opposite this study's El Imposible transects).
+An 5 April 2002, Barn Owls were detected in that same location (Vidal Campos Aguirre, personal communication).
+Vidal Campos Aguirre also stated that Black-and-white Owls were observed in two areas of the Mistepe River Valley in that same timeframe.
+It is likely, then, that these two owls still occur in El Imposible, though we were not able to detect them.
+We also observed a Barn Owl nesting near the N-1 route; however, again, we did not include any visual or auditory detections of owls outside of the survey stations.
+Barn Owls were detected in both Nancuchiname routes, but Black-and-white Owls were not detected in any routes during our surveys.
+
+Nancuchiname Forest was the richest in diversity of our three protected areas; however, many of the more specialist species were only detected in Montecristo NP (e.g., Whiskered Owl, Great Horned Owl, Fulvous Owl, and c.f.
+Stygian Owl).
+The 2002 Nancuchiname Forest Management Plan only listed one owl, Ferruginous Pygmy-owl, as being present [@Zepeda:1995]xx *why does year not match**, but we confirmed the presence of four additional owl species in that NP.
+Fulvous Owl is one of El Salvador's 18 endemic bird species [@Komar:2002], and it was detected numerous times in the cloud forest of Montecristo NP (M-1).
+Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
+
+@Perez-Leon:2017 stated that there are human activities affecting owl populations in El Salvador, including illegal hunting, trapping, persecution, killing, and wildlife trade.
+They also stated that, of eight species of owls confiscated from local markets in 1995 to 2008, the most frequent were Mottled Owls and Ferruginous Pygmy-owls.
 

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
updated as many references as I could

@@
 author: "Jane West and Althea Archer"
-date: "1/2/2021"
+date: "1/11/2021"
 output:
@@
 
-Neotropical owls are less well-studied than temperate species, such as the spotted owls of North America, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017].
+Neotropical owls are less well-studied than temperate species, such as the spotted owls of North America, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017; @Rangel-Salazar:2017].
 Yet, neotropical owls are reliant upon ecosystems under tremendous anthropogenic pressure and rapid and unpredictable responses to climate change [@Corlett:2012].
@@
 The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and species community composition.
-El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot (Conservation International 2009 in Komar and Ibarra-Portillo 2009).
+El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot [@Myers:2000].
 We surveyed owls on foot using passive listening and broadcast vocalizations in two transects each within the three largest protected areas of El Salvador from 2003 through 2013.
@@
 
-We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001] and a similar protocol developed for Costa Rica [@Enriquez:2001].
+We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and raptor research guidelines [@Fuller:1987].
 This study's survey protocol was an extension of previous research completed in the western half of El Imposible NP [@West:1988].
@@
 This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 1XX).
-This upper limit to species richness relates directly to the 13 species known to inhabit El Salvador (cite Owls of El Salvador here) plus one species we identified (Stygian owl), which was previously undocumented in El Salvador.
+This upper limit to expected species richness relates directly to the 13 species known to inhabit El Salvador [@Dickey:1938] plus one species we identified (c.f.
+Stygian owl), which was previously undocumented in El Salvador.
 We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
@@
 
+An estimated 75% of the nearly 250 surviving species of owls are associated with dense and undisturbed forests [@Mikkola:2012].
+@Perez-Leon:2017 stated that within El Salvador, natural ecosystems were the most diverse, specifically natural habitats of cloud, deciduous, riparian and pine-oak forests.
+In El Salvador, 13.6% of its land is forested [@CIA:2020].
+The country has continued with the traditional shade-growing coffee farming, which helps maintain forest cover.
+Coffee plantations account for around 7% of its forested land because of the intensive use of shade-cover farming practices [@Silva:2016].
+@Silva:2016 also stated that almost all primary or closed forest is surrounded by coffee farms, which act as a buffer zone against land use change.
+@Perez-Leon:2017 stated that coffee plantations are the most important agroecosystem for owl communities’ survival in El Salvador.
+
 In Central America, the Mottled Owl is the most common and widespread owl [@Vallely:2018].
@@
 Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
+
+Distinguishing individuals by non-invasive means, such as vocal traits, is preferable when species are rare, sensitive to handling, elusive, or when other techniques are unfeasible [@Terry:2005].
+Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally recorded owl vocalizations when possible.
+In Guatemala, @Gerhardt:1989 found that a series of vocalizations obtained from a non-local ornithological laboratory, which were identified as vocalizations of the Mottled Owl, could not generate a response from Guatemalan owls.
+We found that the probability of detecting most owls increased after we broadcasted the vocalizations of Mottled Owls, Pacific Screech-owls, Crested Owls, and Spectacled Owls.
+On the other hand, owls were less likely to call or be detected after we played Great Horned Owl calls.
+Because owls, like many birds of prey, are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019], we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
+
+We did not use spectogram analysis to determine any of the specific individuals of owls that we detected.
+Spectograms of acoustic signals, such as owl vocalizations, are capable of not only recognizing differences between species, but even potentially individual owls of the same species or owl gender [@Eakle:1989; @Farquahar:1993; @Galeotii:1993; @Rognan:2009].
+*xx can't find Eakle 1989, Farquahar 1993, Galeotii 1993* Identifying individuals within a population can improve census estimates and provide useful information on demographics, life history traits, and behavior--all of which often influence management decisions [@Terry:2005].
+Spectogram analysis could help future owl researchers in El Salvadar determine more accurate population sizes of the more rare species we detected, such as Whiskered Screech-owls, c.f.
+Stygian Owls, and Great Horned Owls.
 

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
renumbered figures and rewrote fig captions

@@
 To model species richness at each route in each year, we assumed that there were 14 possible species present in El Salvador.
-This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 1XX).
+This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 2XX).
 This upper limit to expected species richness relates directly to the 13 species known to inhabit El Salvador [@Dickey:1938] plus one species we identified (c.f.
@@
 Occupancy probabilities were relatively high across all three species for both routes of Nancuchiname (Fig.
-1XX).
+2XX).
 We found that the probability of occupancy for Mottled owls was relatively consistent across time in the El Imposible routes and in N-2 (Fig.
-2XX), but that it seemed to increase in the last ~6 years in M-2 and N-1 (Fig.
-2XX).
+3XX), but that it seemed to increase in the last ~6 years in M-2 and N-1 (Fig.
+3XX).
 The probability of occupancy in M-1 seemed to have decreased through time for Mottled owls (Fig.
-2XX).
+3XX).
 
-For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI-1 in the early years of the surveys (Figs.
-3XX and 4XX).
+For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI-1 in the early years of the surveys (Fig.
+3XX).
 The probability of occupancy for Ferruginous Pygmy owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
 3XX), and the probability of occupancy for the Spectacled owl  fluctuated slightly from year to year in the Nancuchaname routes (Fig.
-4XX).
+3XX).
 
@@
 Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
-These were equivalent to the number of species detected in each route (Table 1XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
+These were equivalent to the number of species detected in each route (Table 2XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
 
 When analyzed by year, median species richness ($\text{Richness}^*_{t,h}$) varied from 0 to 5, depending on route and year (Fig.
-5XXX).
+4XXX).
 At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species (e.g., in M-2 during 2004 where we detected zero species (Fig.
-5XXX).
+4XXX).
 
@@
 Including all detected species into one richness model increased the precision of the estimates of the effect of broadcast calls on detection probability (i.e., compare All Species with individual occupancy results in Fig.
-6XXX).
+5XXX).
 When incorporating all owl species, broadcasting calls from Mottled, Pacific, Crested, Black and White, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned owl calls decreased detection (Fig.
-6XXX).
+5XXX).
 
 The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled, Ferruginous Pygmy, and Spectacled owls (Fig.
-6XX).
+5XX).
 The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owl calls.
 The probability of detecting Ferruginous Pygmy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owl calls (Fig.
-6XX).
+5XX).
 
@@
 Contrarily, @Enriquez:2001 stated that Mottled Owl populations in La Selva may have decreased during the past 30 years.
-The results of this study indicated that Mottled Owls are still fairly common at each of the protected areas, although there seemed to be a downward trend in occupancy in M-2 (Fig xXX).
+The results of this study indicated that Mottled Owls are still fairly common at each of the protected areas, although there seemed to be a downward trend in occupancy in M-2 (Fig 3XX).
 
@@
 ![1XX.
-Probability of occupancy ($\psi_{t,h}$) averaged by route for Mottled, Ferruginous Pygmy, and Spectacled Owls.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+Map of El Salvador.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
+\label{psi_means}](../output/figures/psi_means-12.png)
+
+![2XX.
+Probability of occupancy ($\psi_{t,h}$) averaged by route for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
 Posteriors were summarized with median $\pm$ 90% credible intervals.
@@
 
-![2XX.
-Probability of occupancy by route and year for Mottled Owl.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
+![3XX.
+Probability of occupancy by route and year for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure2}](../output/figures/mottd_psi_byYr-1.png)
-
-![3XX.
-Probability of occupancy by route and year for Ferruginous Pygmy Owl.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
-Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure3}](../output/figures/ferpy_psi_byYr-1.png)
+\label{figure2}](../output/figures/psi_byYr-1.png)
 
 ![4XX.
-Probability of occupancy by route and year for Spectacled Owl.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
-Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure4}](../output/figures/specd_psi_byYr-1.png)
+Species richness (Richness*) by route and year.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
+Posteriors were summarized with median (black dot) $\pm$ 90% credible intervals (dark grey shading and vertical black lines).
+The number of owl species detected at each route and year is indicated with the light grey area.
+\label{figure5}](../output/figures/richness_byRtYr-1.png)
 
 ![5XX.
-Species richness (Richness*) by route and year.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.
-Posteriors were summarized with median (black dot) $\pm$ 90% credible intervals (black lines).
-The number of species detected at each route and year are indicated with a red dot.
-\label{figure5}](../output/figures/richness_byRtYr-1.png)
-
-![6XX.
-The logistic regression parameters that related the probability of detecting owls to the broadcast timeframe (pre-broadcast) or species.
+The probabiliity of detecting owls during spontaneous surveys (i.e., pre-broadcast) and after broadcast, depending on the species of owl used for broadcast vocalization.
+These logistic regression coefficients were evaluated from the richness model, which combined all detected species (All Species) and the single-species occupancy models for Ferruginous Pygmy-owls (FEPO), Mottled Owls (MOOW), and Spectacled Owls (SPEO).
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
introduction edits

@@
 
-Owls, as a keystone species, are an important, yet cryptic, bioindicator of ecosystem health.
-Top predators also often have long-lasting and wide-ranging effects through trophic cascades.
-The various spotted owl sub-species (*Strix occidentalis*) of North America have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owl have complex and contrasting behavior responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
+Owls are a keystone species in many ecosystems, and as such are an important bioindicator of ecosystem health.
+Top predators such as owls affect communities through their top-down control on biodiversity and their presence or absence can cause trophic cascades.
+However, owls are also cryptic, nocturnal predators that are difficult to accurately census.
+The Mexican Spotted Owl (*Strix occidentalis*) and its sub-species have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owls have complex and often contrasting responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
 
-Neotropical owls are less well-studied than temperate species, such as the spotted owls of North America, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017; @Rangel-Salazar:2017].
+Neotropical owls are less well-studied than temperate species, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017; @Rangel-Salazar:2017].
 Yet, neotropical owls are reliant upon ecosystems under tremendous anthropogenic pressure and rapid and unpredictable responses to climate change [@Corlett:2012].
-In the tropics, particularly in rain and cloud forests, owl studies are more difficult than in temperate regions, although there are more species living in the same habitat [@Konig:1999].
-In general, while our knowledge about the conservation status of neotropical populations is limited, we do know that owl abundance and distribution is decreasing in several regions, where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
+In the tropics, particularly in rain and cloud forests, owl studies are even more difficult than in temperate regions, regardless of denser populations [@Konig:1999].
+In general, while our knowledge about the conservation status of neotropical populations is limited, we do know that owl abundance and distribution is decreasing in several regions where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
 
-The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration of plumage as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
+The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
 *springer and forsman not cited on doc from Jane xxx* Owls vocalize to communicate with the same species and to delimit territory [@Johnsgard:1988].
-Spontaneous and provocation listening are two methods for auditory surveys that have been used with increasing frequency in surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
+Spontaneous and provocation listening are two methods that researchers use with increasing frequency to survey nocturnal birds of prey [e.g., @Baumgardt:2019].
 Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls." One benefit of auditory surveys (including both spontaneous and broadcast) is being able to identify multiple species simultaneously.
-Adding in provocation to auditory surveys can increase the detection probability of owls, although the intra- and interspecific responses to broadcast calls could play a role in response patterns [@Enriquez:1997;@Baumgardt:2019].
+Adding in provocation to auditory surveys can increase the detection probability of owls, although intra- and inter-species interactions can play a role whether owls are more likely or less likely to call in response to a vocalization simulus [@Enriquez:1997;@Baumgardt:2019].
 
-The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and species community composition.
-El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot [@Myers:2000].
-We surveyed owls on foot using passive listening and broadcast vocalizations in two transects each within the three largest protected areas of El Salvador from 2003 through 2013.
-
-This project was carried out during the breeding season (March-May) for three weeks each year over a 10-year period in the three previously mentioned protected natural areas.
-These areas were selected to represent different ecosystems found in the country and the transect routes were selected to represent the habitats in the protected natural areas.
+The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and community composition.
+El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot [@Myers:2000], and we selected the three survey areas to represent ecosystem diversity across El Salvador.
+We set up two routes in each protected area to further represent the habitats diversity of this heterogeneous country.
+We conducted nighttime foot surveys using passive listening and broadcast calls during the breeding season for three weeks each year from from 2003 through 2013.
 

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
updated route and area descriptions to past tense

@@
 
-El Salvador is the smallest and most densely populated (approximately 6.5 million) Central American country with 13.6% of its land forested [@CIA:2020].
+El Salvador is the smallest and most densely populated Central American country with 13.6% of its land forested [@CIA:2020].
 It is bordered by the Pacific Ocean, Guatemala and Honduras (Fig.
 1XXX).
-This project includes three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that are representative of several of the country’s protected natural areas (Table 1XXX).
+We conducted surveys in three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that are representative of the country's diverse ecosystems (Fig.
+1xxx, Table 1XXX).
 
-El Imposible National Park (NP) is located 119 km southwest of San Salvador (Fig.
+El Imposible National Park (NP) was located 119 km southwest of San Salvador (Fig.
 1XXX, Table 1XXX).
-Its elevation ranges from 250 to 1425 m above sea level.
-It is the largest NP in the country, covering 3792 ha.
-The topography is extremely steep and broken, with many cliffs [@Alvarez:2003].
-The area contains deciduous and semi-deciduous forest, secondary growth vegetation on the edges and former pasture areas that were reforested with native species in 1997.
-In the western portion of this NP, one survey transect (EI-1) was set up in the lower elevation deciduous in an area of reforestation and the other (EI-2) was located in semi-deciduous forest.
+Its elevation ranged from 250 to 1425 m above sea level.
+It was the largest NP in the country, covering 3792 ha.
+The topography was extremely steep and broken, with many cliffs [@Alvarez:2003].
+The area contained deciduous and semi-deciduous forest, secondary growth vegetation on the edges, and former pasture areas that were reforested with native species in 1997.
+We set up one survey route (EI-1) in the western portion of El Imposible in an area of reforestation and at a low elevation.
+We set up the other survey route (EI-2) in an area of semi-deciduous forest.
 
-Montecristo NP is located 125 km northwest of San Salvador (Fig.
+Montecristo NP was located 125 km northwest of San Salvador (Fig.
 1XXX; Table 1XXX).
-Its altitude ranges from 730 to 2418 m above sea level and covers 1973 ha [@MARN:2010].
-The higher elevations contain cloud forest, middle elevations contain pine/oak forest (with some pine and cypress plantations) and the lower elevations contain a semi-deciduous tropical forest [@MARN:2015b].
-This NP is part of the Montecristo Tri-national Protected Area which includes extensive adjoining natural areas in Guatemala and Honduras [@Komar:2010].
-In this NP, one transect (M-1) was set up in a cloud forest and another (M-2) was located at a lower elevation in a pine/oak forest.
+Its altitude ranged from 730 to 2418 m above sea level, and the NP covered 1973 ha [@MARN:2010].
+The higher elevations contained cloud forest, middle elevations contained pine/oak forest (with some pine and cypress plantations) and the lower elevations contained a semi-deciduous tropical forest [@MARN:2015b].
+This NP was part of the Montecristo Tri-national Protected Area which included extensive adjoining natural areas in Guatemala and Honduras [@Komar:2010].
+In this NP, one survey route (M-1) was set up in a cloud forest and another (M-2) was located at a lower elevation in a pine/oak forest.
 
-Nancuchiname Forest, a Protected Natural Area, is located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
+Nancuchiname Forest, a Protected Natural Area, was located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
 1XXX).
-Its elevation ranges from 5 to 12 m above sea level and covers 797 ha.
-Nancuchiname Forest represents the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001]xxx **we can't site a geocities website**.
-The area contains deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike.
-In Nancuchiname Forest, one survey transect (N-1) was set up in alluvial forest along a dike surrounded by second growth vegetation and another (N-2) in a semi-deciduous alluvial forest.
+Its elevation ranged from 5 to 12 m above sea level and the protected area covered 797 ha.
+Nancuchiname Forest represented the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001]xxx **we can't site a geocities website**.
+The area contained deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike.
+In Nancuchiname Forest, we set up one survey route (N-1) in alluvial forest along the dike, which was surrounded by secondary growth, and we set up the other route (N-2) in a semi-deciduous alluvial forest.
 
@@
 During 1979-1981, in the western portion of El Imposible NP, West (1988) heard or observed six owl species, including two that were not detected during this study (Barn Owls and Black-and-white Owls) in that NP.
-In 1980, Barn Owls nested in a cave of Loma de Paja Mountain (opposite this study's El Imposible transects).
+In 1980, Barn Owls nested in a cave of Loma de Paja Mountain (opposite this study's El Imposible routes).
 An 5 April 2002, Barn Owls were detected in that same location (Vidal Campos Aguirre, personal communication).

### 2021-01-12 manuscript_working/ms_working_RMD.Rmd
beginning to fix survey methods section, do i have coordinates?

@@
 
-*Table 1XX. Route locations and habitats.*
+*Table 1XX.
+Route locations and habitats.
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
+*
 
@@
 #  c(NamePark, NameRoute, Latitude, Longitude, Habitat), byrow = T, nrow = 5, ncol = 7))
-parkTable <- data.frame(Area, Route, Latitude, Longitude, Elevation, Habitat)
+parkTable <- data.frame(Route, Latitude, Longitude, Elevation, Habitat)
 pander(parkTable)
@@
 
-We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and raptor research guidelines [@Fuller:1987].
+We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and other raptor research guidelines [@Fuller:1987].
 This study's survey protocol was an extension of previous research completed in the western half of El Imposible NP [@West:1988].
-For our surveys, we created six 2-km long survey routes (transects) that each included 10 permanent survey stations at 200 m intervals.
+For our surveys, we created six 2-km long survey routes (transects) that each included 10 permanent survey stations set at 200 m intervals.
 Two routes were established in each protected area (Table 1xxx).
@@
 Stygian Owl detected during the course of our surveys, we did also hear a c.f.
-Stygian Owl between the survey points on 25 April XXXXyear.
+Stygian Owl between the survey stations on 25 April XXXXyear.
 We also heard c.f.

### 2021-01-14 manuscript_working/ms_working_RMD.Rmd
updating owl name format in ms and figures

@@
 The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
-*springer and forsman not cited on doc from Jane xxx* Owls vocalize to communicate with the same species and to delimit territory [@Johnsgard:1988].
-Spontaneous and provocation listening are two methods that researchers use with increasing frequency to survey nocturnal birds of prey [e.g., @Baumgardt:2019].
-Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls." One benefit of auditory surveys (including both spontaneous and broadcast) is being able to identify multiple species simultaneously.
+*springer 1978 and forsman 1983 are not cited on previous ms draft xxx* Owls vocalize to communicate with the same species, threaten or provoke other species, or to delimit territory [@Johnsgard:1988].
+Spontaneous and provocation listening are two methods that researchers use with increasing frequency during acoustic surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
+Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls." One benefit of auditory surveys, whether spontaneous or broadcast, is being able to identify multiple species during the same survey period.
 Adding in provocation to auditory surveys can increase the detection probability of owls, although intra- and inter-species interactions can play a role whether owls are more likely or less likely to call in response to a vocalization simulus [@Enriquez:1997;@Baumgardt:2019].
@@
 We set up two routes in each protected area to further represent the habitats diversity of this heterogeneous country.
-We conducted nighttime foot surveys using passive listening and broadcast calls during the breeding season for three weeks each year from from 2003 through 2013.
+We conducted nighttime foot surveys using passive listening and broadcast calls during the breeding season for three weeks each year from 2003 through 2013.
 
@@
 This NP was part of the Montecristo Tri-national Protected Area which included extensive adjoining natural areas in Guatemala and Honduras [@Komar:2010].
-In this NP, one survey route (M-1) was set up in a cloud forest and another (M-2) was located at a lower elevation in a pine/oak forest.
+In this NP, one survey route (M-1) was set up in a cloud forest and another (M-2) was located at a moderate elevation in a pine/oak forest.
 
@@
 The broadcast calls varied between routes, but were consistent at each route and station across years.
-Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible (EI-1 and EI-2) and Nancuchiname (N-1) and N-2).
-We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo (M-1 and M-2).
+Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible NP (EI-1 and EI-2) and Nancuchiname Forest (N-1 and N-2).
+We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo NP (M-1 and M-2).
 Surveys were repeated at each route up to three times a year from 2003 through 2013, depending on site access and weather.
 We did not conduct any surveys in 2006.
-We georeferenced the owl detections for each survey and route  to eliminate double counting of the owl observations.
+We georeferenced the owl detections for each survey and route  to eliminate double counting of individual owl observations.
 
@@
 This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 2XX).
-This upper limit to expected species richness relates directly to the 13 species known to inhabit El Salvador [@Dickey:1938] plus one species we identified (c.f.
+This upper limit to expected species richness related directly to the 13 species known to inhabit El Salvador [@Dickey:1938] plus one species we identified (c.f.
 Stygian owl), which was previously undocumented in El Salvador.
 We thus augmented our owl detection data with 5 additional potential species with all-zero detection records [@Royle:2007].
+
 For each species $s$ (where $s = 1,2, \ldots,15$), $w_{h,s}$ was a binary indicator of whether that owl species was present in each route based on the probability of species presence in that route, $\Omega_h$ [@Broms:2016]:
@@
 We also estimated year-to-year richness ($\text{Richness}^*_{t,h}$) by route by adding up the number of unique species that were present in each route, year, and survey.
-We first determined if a species was present in any of that route and year's set of surveys with a species-specific binary indicator $\vartheta_{t,h,s}$.
+We first determined if a species was present in any of that route and year's set of surveys with a species-specific binary indicator $\vartheta_{t,h,s}$ where $\vartheta_{t,h,s} = 1$ if $z_{t,h,i,s} = 1$ in any of that route's surveys $i$ that year.
 Then, we summed up the number of species present in each year and route:
@@
 We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the Bayesian models.
-We verified that R-hats were lower than XXXX and visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
+We verified that R-hats were lower than 1.3 and visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
 
@@
 No other owl species had enough detections to analyze species-specific occupancy (Table 2XX).
-Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled owls than for Ferruginous Pygmy or Spectacled owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
+Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled Owls than for Ferruginous Pygmy-owls or Spectacled Owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
 1X).
@@
 2XX).
-We found that the probability of occupancy for Mottled owls was relatively consistent across time in the El Imposible routes and in N-2 (Fig.
-3XX), but that it seemed to increase in the last ~6 years in M-2 and N-1 (Fig.
+We found that the probability of occupancy for Mottled Owls was relatively consistent across time in the El Imposible routes and in N-2 (Fig.
+3XX), but that it seemed to increase in the last approximately 6 years in M-2 and N-1 (Fig.
 3XX).
-The probability of occupancy in M-1 seemed to have decreased through time for Mottled owls (Fig.
+The probability of occupancy in M-1 seemed to have decreased through time for Mottled Owls (Fig.
 3XX).
 
-For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy and Spectacled stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy in EI-1 in the early years of the surveys (Fig.
+For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy-owls and Spectacled Owls stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy-owl in EI-1 in the early years of the surveys (Fig.
 3XX).
-The probability of occupancy for Ferruginous Pygmy owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
-3XX), and the probability of occupancy for the Spectacled owl  fluctuated slightly from year to year in the Nancuchaname routes (Fig.
+The probability of occupancy for Ferruginous Pygmy-owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
+3XX), and the probability of occupancy for the Spectacled Owl fluctuated slightly from year to year in the Nancuchaname routes (Fig.
 3XX).
@@
 Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
-These were equivalent to the number of species detected in each route (Table 2XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
+These median values were equivalent to the number of species detected in each route (Table 2XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
 
@@
 5XXX).
-When incorporating all owl species, broadcasting calls from Mottled, Pacific, Crested, Black and White, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned owl calls decreased detection (Fig.
+When incorporating all owl species, broadcasting calls from Mottled Owls, Pacific Screech-owls, Crested Owls, Black-and-white Owls, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned Owl calls decreased detection (Fig.
 5XXX).
 
-The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled, Ferruginous Pygmy, and Spectacled owls (Fig.
+The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled Owls, Ferruginous Pygmy-owls, and Spectacled Owls (Fig.
 5XX).
-The probability of detecting Mottled Owls increased after broadcasting Mottled, Pacific, Crested, Black and White, or Spectacled owl calls.
-The probability of detecting Ferruginous Pygmy owls increased after broadcasting Mottled and Pacific owl calls, and the probability of detecting Spectacled owls increased after broadcasting Pacific, Black and White, and Spectacled owl calls (Fig.
+The probability of detecting Mottled Owls increased after broadcasting Mottled Owl, Pacific Screech-owl, Crested Owl, Black-and-white Owl, or Spectacled Owl calls.
+The probability of detecting Ferruginous Pygmy-owls increased after broadcasting Mottled Owl and Pacific Owl calls, and the probability of detecting Spectacled Owls increased after broadcasting Pacific Screech-owl, Black-and-white Owl, and Spectacled Owl calls (Fig.
 5XX).

### 2021-01-14 programs/e01_processing_results_ferpy.R

@@ -199,10 +199,12 @@ p.det.post.ferpy$Broadcast <- factor(p.det.post.ferpy$broadcast.param,
                                              "beta.bw", "beta.spectacled",
                                              "beta.whiskered", "beta.gbarred",
                                              "beta.stygian", "beta.ghorned"),
-                                  labels = c("Pre-broadcast", "Mottled",
-                                             "Pacific Screech", "Crested", "Black-and-white",
-                                             "Spectacled", "Whiskered Screech", "Fulvous",
-                                             "Stygian", "Great Horned"))
+                                  labels = c("Pre-broadcast", "Mottled Owl",
+                                             "Pacific Screech-owl", "Crested Owl", 
+                                             "Black-and-white Owl",
+                                             "Spectacled Owl", 
+                                             "Whiskered Screech-owl", "Fulvous Owl",
+                                             "Stygian Owl", "Great Horned Owl"))
 
 #' _____________________________________________________________________________
 #' ## Save files

### 2021-01-14 programs/e02_processing_results_Mottd.R

@@ -204,10 +204,12 @@ p.det.post.mottd$Broadcast <- factor(p.det.post.mottd$broadcast.param,
                                              "beta.bw", "beta.spectacled",
                                              "beta.whiskered", "beta.gbarred",
                                              "beta.stygian", "beta.ghorned"),
-                                  labels = c("Pre-broadcast", "Mottled",
-                                             "Pacific Screech", "Crested", "Black-and-white",
-                                             "Spectacled", "Whiskered Screech", "Fulvous",
-                                             "Stygian", "Great Horned"))
+                                  labels = c("Pre-broadcast", "Mottled Owl",
+                                             "Pacific Screech-owl", "Crested Owl", 
+                                             "Black-and-white Owl",
+                                             "Spectacled Owl", 
+                                             "Whiskered Screech-owl", "Fulvous Owl",
+                                             "Stygian Owl", "Great Horned Owl"))
 
 #' _____________________________________________________________________________
 #' ## Save files

### 2021-01-14 programs/e03_processing_results_specd.R

@@ -199,10 +199,12 @@ p.det.post.specd$Broadcast <- factor(p.det.post.specd$broadcast.param,
                                              "beta.bw", "beta.spectacled",
                                              "beta.whiskered", "beta.gbarred",
                                              "beta.stygian", "beta.ghorned"),
-                                  labels = c("Pre-broadcast", "Mottled",
-                                             "Pacific Screech", "Crested", "Black-and-white",
-                                             "Spectacled", "Whiskered Screech", "Fulvous",
-                                             "Stygian", "Great Horned"))
+                                  labels = c("Pre-broadcast", "Mottled Owl",
+                                             "Pacific Screech-owl", "Crested Owl", 
+                                             "Black-and-white Owl",
+                                             "Spectacled Owl", 
+                                             "Whiskered Screech-owl", "Fulvous Owl",
+                                             "Stygian Owl", "Great Horned Owl"))
 
 #' _____________________________________________________________________________
 #' ## Save files

### 2021-01-14 programs/e04_processing_results_richness.R

@@ -195,10 +195,12 @@ p.det.post.richness$Broadcast <- factor(p.det.post.richness$broadcast.param,
                                                    "beta.bw", "beta.spectacled",
                                                    "beta.whiskered", "beta.gbarred",
                                                    "beta.stygian", "beta.ghorned"),
-                                        labels = c("Pre-broadcast", "Mottled",
-                                                   "Pacific Screech", "Crested", "Black-and-white",
-                                                   "Spectacled", "Whiskered Screech", "Fulvous",
-                                                   "Stygian", "Great Horned"))
+                                        labels = c("Pre-broadcast", "Mottled Owl",
+                                                   "Pacific Screech-owl", "Crested Owl", 
+                                                   "Black-and-white Owl",
+                                                   "Spectacled Owl", 
+                                                   "Whiskered Screech-owl", "Fulvous Owl",
+                                                   "Stygian Owl", "Great Horned Owl"))
 
 
 #' Species presence by route

### 2021-01-14 programs/g_graphing_results_global.R

@@ -158,7 +158,8 @@ ggplot(data = psi.post.all,
 #' Merge data
 p.det.post.richness$Species <- "All Species"
 #' 
-p.det.post <- rbind(p.det.post.ferpy, p.det.post.mottd, p.det.post.specd,p.det.post.richness)
+p.det.post <- rbind(p.det.post.ferpy, p.det.post.mottd, 
+                    p.det.post.specd,p.det.post.richness)
 table(p.det.post$Broadcast)
 
 table(p.det.post$Species)

### 2021-01-14 programs/d04_richness_analysis.R
updated figs for richness

@@ -178,6 +178,7 @@ richness.jagsout <- jags(data = richness.jags.data,
                       n.iter = 20000, #10000
                       n.burnin = 2000, #1000
                       n.thin = 2)
+# began 2021-01-13 22:39
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy

### 2021-01-14 programs/f01_graphing_results_ferpy.R
ferpy, rerun

@@ -30,14 +30,14 @@ set.seed(587453)
 #' 
 #' 
 #' Psi Posteriors by year and route
-load(file = "data/output_data/ferpy_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/ferpy_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
-load(file = "data/output_data/ferpy_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/ferpy_psi_posteriors_RtSpp.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
 #' 
-load(file = "data/output_data/ferpy_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/ferpy_p_detection_posteriors.Rdata")
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy

### 2021-01-14 programs/e03_processing_results_specd.R
specd rerun

@@ -210,14 +210,14 @@ p.det.post.specd$Broadcast <- factor(p.det.post.specd$broadcast.param,
 #' ## Save files
 #' 
 #' Psi Posteriors by year and route
-save(psi.post.specd, file = "data/output_data/specd_psi_posteriors_RtYr.Rdata")
+save(psi.post.specd, file = "data/plotting_data/specd_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
-save(psi.means.specd, file = "data/output_data/specd_psi_posteriors_RtSpp.Rdata")
+save(psi.means.specd, file = "data/plotting_data/specd_psi_posteriors_RtSpp.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
 #' 
-save(p.det.post.specd, file = "data/output_data/specd_p_detection_posteriors.Rdata")
+save(p.det.post.specd, file = "data/plotting_data/specd_p_detection_posteriors.Rdata")
 
 
 #' _____________________________________________________________________________

### 2021-01-14 programs/f03_graphing_results_specd.R

@@ -30,14 +30,14 @@ set.seed(587453)
 #' 
 #' 
 #' Psi Posteriors by year and route
-load(file = "data/output_data/specd_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/specd_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
-load(file = "data/output_data/specd_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/specd_psi_posteriors_RtSpp.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
 #' 
-load(file = "data/output_data/specd_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/specd_p_detection_posteriors.Rdata")
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy

### 2021-01-14 manuscript_working/ms_working_RMD.Rmd
updated richness results

@@
 
-Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 5 (4,5), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
+Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 3 (3,3), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
 These median values were equivalent to the number of species detected in each route (Table 2XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.

### 2021-01-14 programs/e04_processing_results_richness.R
reran models

@@ -115,6 +115,7 @@ tempYear <- tempYear[include.byrow]
 tempRoute <- 
   rep(c("EI1", "EI2", "M1", "M2", "N1", "N2"), length(year.index))
 tempRoute <- tempRoute[include.byrow]
+richness.RtYr.post <- richness.RtYr.post[include.byrow,]
 richness.RtYr.post$Year <- tempYear
 richness.RtYr.post$Route <- tempRoute
 
@@ -131,7 +132,7 @@ colnames(richness.RtYr.post)
 colnames(richness.RtYr.post) <- c("Richness.mean", "Richness.sd", "Richness.LL05", 
                              "Richness.median",
                              "Richness.UL95", "Richness.Rhat", "Richness.neff", 
-                             "Year", "Route", "Region", "richness.d", "richness.detected")
+                             "Year", "Route", "Region")#, "richness.d", "richness.detected")
 
 
 
@@ -178,8 +179,8 @@ for(rr in 1:nrow(richness.RtYr.post)){
 #' Merge data together and change column names
 #' 
 colnames(p.det.post.richness)
-#colnames(p.det.post.richness) <- c("mean", "sd", "LL05", "median", "UL95", "Rhat",
-#                                   "n.eff", "broadcast.param")
+colnames(p.det.post.richness) <- c("mean", "sd", "LL05", "median", "UL95", "Rhat",
+                                   "n.eff", "broadcast.param")
 
 #' Transform to probability scale
 #' 

### 2021-01-14 programs/f02_graphing_results_Mottd.R

@@ -30,14 +30,14 @@ set.seed(587453)
 #' 
 #' 
 #' Psi Posteriors by year and route
-load(file = "data/output_data/mottd_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/mottd_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
-load(file = "data/output_data/mottd_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/mottd_psi_posteriors_RtSpp.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
 #' 
-load(file = "data/output_data/mottd_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/mottd_p_detection_posteriors.Rdata")
 
 #' _____________________________________________________________________________
 #' ## Psi = Probability of occupancy

### 2021-01-14 programs/g_graphing_results_global.R

@@ -30,21 +30,21 @@ set.seed(587453)
 #' 
 #' 
 #' Psi Posteriors by year and route
-load(file = "data/output_data/specd_psi_posteriors_RtYr.Rdata")
-load(file = "data/output_data/ferpy_psi_posteriors_RtYr.Rdata")
-load(file = "data/output_data/mottd_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/specd_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/ferpy_psi_posteriors_RtYr.Rdata")
+load(file = "data/plotting_data/mottd_psi_posteriors_RtYr.Rdata")
 load(file = "data/plotting_data/richness_psi_posteriors_RtYr.Rdata")
 
 #' Psi posteriors across years by species and route
-load(file = "data/output_data/specd_psi_posteriors_RtSpp.Rdata")
-load(file = "data/output_data/ferpy_psi_posteriors_RtSpp.Rdata")
-load(file = "data/output_data/mottd_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/specd_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/ferpy_psi_posteriors_RtSpp.Rdata")
+load(file = "data/plotting_data/mottd_psi_posteriors_RtSpp.Rdata")
 
 #' Probability of detection by broadcast species and species of analysis
 #' 
-load(file = "data/output_data/specd_p_detection_posteriors.Rdata")
-load(file = "data/output_data/ferpy_p_detection_posteriors.Rdata")
-load(file = "data/output_data/mottd_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/specd_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/ferpy_p_detection_posteriors.Rdata")
+load(file = "data/plotting_data/mottd_p_detection_posteriors.Rdata")
 load(file = "data/plotting_data/richness_p_detection_posteriors.Rdata")
 
 #' Species accounts
@@ -177,14 +177,7 @@ p.det.post$Species[p.det.post$Species == "Specd"] <- "SPEO"
 #' 
 #+ p_detection, fig.width = 2.8346, dpi = 600, fig.height = 3
 ggplot(data = p.det.post, aes(y = median.plogis, x = Broadcast, group = Species))+
-  #geom_bar(stat = "identity", position = position_dodge(), 
[…]

### 2021-01-14 programs/d02_occupancy_analysis_Mottd.R
reran mottled model

@@ -22,7 +22,7 @@ library(R2jags)
 #' to be loaded, and setting the seed ensures that your analysis is 
 #' repeatable*
 remove(list = ls())
-set.seed(1770562)
+set.seed(64781)
 
 #' _____________________________________________________________________________
 #' ## Load Data

### 2021-01-14 programs/g_graphing_results_global.R
axes formatted to lowercase except first word

@@ -91,7 +91,7 @@ ggplot(data = psi.means, aes(x = Route, y = Psi.median, group = code))+
   ylim(c(0,1))+
   scale_fill_manual(values = c("#cccccc", "#969696", "#525252"))+
   theme_minimal()+
-  ylab("Probability of Occupancy")+
+  ylab("Probability of occupancy")+
   xlab("Route")+
   theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=6), 
@@ -140,7 +140,7 @@ ggplot(data = psi.post.all,
   facet_grid(Route~code)+
   theme_minimal()+
   xlab("Year")+
-  ylab("Probability of Occupancy")+
+  ylab("Probability of occupancy")+
   scale_x_continuous(breaks = 2003:2013, limits = c(2003,2013))+
   theme(axis.text.x = element_text(size=6, angle = 90),
         axis.text.y = element_text(size=6),
@@ -186,8 +186,8 @@ ggplot(data = p.det.post, aes(y = median.plogis, x = Broadcast, group = Species)
   geom_linerange(aes(ymin = LL05.plogis, ymax = UL95.plogis), 
                  position = position_dodge(0.9))+
   facet_wrap(~Species, nrow = 2)+
-  ylab("Probability of Detection")+
-  xlab("Broadcast Species")+
+  ylab("Probability of detection")+
+  xlab("Broadcast species")+
   theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5),
         legend.position = "none",
@@ -226,7 +226,7 @@ ggplot(data = richness.RtYr.post,
   facet_wrap(~Route, nrow = 3)+
   theme_minimal()+
   xlab("Year")+
-  ylab("Species Richness")+
+  ylab("Species richness")+
   scale_x_continuous(breaks = 2003:2013)+
   scale_y_continuous(breaks = seq(0,14,by=2), limits = c(0,6))+
   theme(axis.text.x = element_text(size=6, angle = 90),

### 2021-01-20 manuscript_working/ms_working_RMD.Rmd
version sent to MRE

@@
 author: "Jane West and Althea Archer"
-date: "1/11/2021"
+date: "1/19/2021"
 output:

### 2021-01-28 manuscript_working/ms_working_RMD.Rmd
making map for ms

@@
              "Alluvial Forest")
-Latitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
-Longitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
-Elevation <- c("xxx", "xxx", "xxx", "xxx", "xxx", "xxx")
+Latitude <- c("13xx 50.75'", "13xx 51.38'", 
+              "14xx 24.88'", "14xx 23.19'", 
+              "13xx 19.95'", "13xx 22.03'")
+Longitude <- c("89xx 59.74'", "89xx 58.24'", 
+               "89xx 21.63'", "89xx 22.33'", 
+               "88xx 43.63'", "88xx 43.30'")
+Elevation <- c("609", "958", "2111", "1755", "3", "2")
 Route <- c("EI-1", "EI-2", "M-1", "M-2", "N-1", "N-2")

### 2021-01-28 programs/g_graphing_results_global.R

@@ -15,6 +15,11 @@ library(devtools) # documentation-related
 # analysis-related
 library(ggplot2)
 library(ggthemes)
+library(maptools)
+library(plyr)
+library(rgdal)
+library(raster)
+library(mapproj)
 
 #' Clear environment and set seed
 #' 
@@ -238,6 +243,52 @@ ggplot(data = richness.RtYr.post,
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank())
 
+
+#' _____________________________________________________________________________
+#' ## Location map
+#' 
+#' Extract country data
+ElSalvador <- getData("GADM", country = "SV", level = 0)
+Guatemala <- getData("GADM", country = "GT", level = 0)
+Honduras <- getData("GADM", country = "HN", level = 0)
+
+# Process into df for ggplot
+ElSalvador <- fortify(ElSalvador)
+Guatemala <- fortify(Guatemala)
+Honduras <- fortify(Honduras)
+
+# Range for mapping
+range(ElSalvador$long)
+range(ElSalvador$lat)
+
+#' 
+#' Map of protected areas
+#+ map, fig.width = 5.67, dpi = 600, fig.height = 4.4
+ggplot(data = ElSalvador, aes(x = long, y = lat, group = group))+
+  geom_polygon(fill = "grey", color = "darkgrey")+
+  geom_polygon(data = Honduras, aes(x = long, y = lat, group = group),
+               fill = "lightgrey", color = "darkgrey")+
+  geom_polygon(data = Guatemala, aes(x = long, y = lat, group = group),
+               fill = "lightgrey", color = "darkgrey")+
+  coord_map(xlim = c(-90.2, -87.2), ylim = c(13.0, 14.5))+
+  # El Imposible
+  annotate(geom = "text", y = 13.65, x = -89.6, label = "EINP", size = 4)+
+  annotate(geom = "segment", y = 13.85, yend = 13.75,
+           x = -89.9, xend = -89.7, color = "#999999")+
+  annotate(geom = "text", y = 13.85, x = -89.97, label = "*", size = 8)+
+  # Montecristo
+  annotate(geom = "text", y = 14.2, x = -89.2, label = "MNP", size = 4)+
+  annotate(geom = "text", y = 14.33, x = -89.36, label = "*", size = 8)+
+  annotate(geom = "segment", y = 14.38, yend = 14.3,
+           x = -89.36, xend = -89.3, color = "#999999")+
[…]

### 2021-01-28 programs/g_graphing_results_global.R
map with correct EINP layout

@@ -272,15 +272,16 @@ ggplot(data = ElSalvador, aes(x = long, y = lat, group = group))+
                fill = "lightgrey", color = "darkgrey")+
   coord_map(xlim = c(-90.2, -87.2), ylim = c(13.0, 14.5))+
   # El Imposible
-  annotate(geom = "text", y = 13.65, x = -89.6, label = "EINP", size = 4)+
-  annotate(geom = "segment", y = 13.85, yend = 13.75,
-           x = -89.9, xend = -89.7, color = "#999999")+
-  annotate(geom = "text", y = 13.85, x = -89.97, label = "*", size = 8)+
+  annotate(geom = "text", y = 13.74, x = -89.8, label = "EINP", size = 4)+
+  #annotate(geom = "segment", y = 13.85, yend = 13.75,
+   #        x = -89.9, xend = -89.7, color = "#999999")+
+  #annotate(geom = "text", y = 13.85, x = -89.97, label = "*", size = 8)+
+  annotate(geom = "point", y = 13.83, x = -89.97, size = 5, shape = 21, fill = "lightgrey")+
   # Montecristo
   annotate(geom = "text", y = 14.2, x = -89.2, label = "MNP", size = 4)+
   annotate(geom = "text", y = 14.33, x = -89.36, label = "*", size = 8)+
-  annotate(geom = "segment", y = 14.38, yend = 14.3,
-           x = -89.36, xend = -89.3, color = "#999999")+
+  annotate(geom = "segment", y = 14.33, yend = 14.27,
+           x = -89.31, xend = -89.26, color = "#999999")+
   # Nancuchiname
   annotate(geom = "text", y = 13.5, x = -88.7, label = "NF", size = 4)+
   annotate(geom = "text", y = 13.35, x = -88.72, label = "*", size = 8)+

### 2021-01-28 programs/g_graphing_results_global.R
protected areas & labels placed on map

@@ -273,20 +273,13 @@ ggplot(data = ElSalvador, aes(x = long, y = lat, group = group))+
   coord_map(xlim = c(-90.2, -87.2), ylim = c(13.0, 14.5))+
   # El Imposible
   annotate(geom = "text", y = 13.74, x = -89.8, label = "EINP", size = 4)+
-  #annotate(geom = "segment", y = 13.85, yend = 13.75,
-   #        x = -89.9, xend = -89.7, color = "#999999")+
-  #annotate(geom = "text", y = 13.85, x = -89.97, label = "*", size = 8)+
   annotate(geom = "point", y = 13.83, x = -89.97, size = 5, shape = 21, fill = "lightgrey")+
   # Montecristo
-  annotate(geom = "text", y = 14.2, x = -89.2, label = "MNP", size = 4)+
-  annotate(geom = "text", y = 14.33, x = -89.36, label = "*", size = 8)+
-  annotate(geom = "segment", y = 14.33, yend = 14.27,
-           x = -89.31, xend = -89.26, color = "#999999")+
+  annotate(geom = "text", y = 14.23, x = -89.2, label = "MNP", size = 4)+
+  annotate(geom = "point", y = 14.33, x = -89.36, size = 5, shape = 21, fill = "lightgrey")+
   # Nancuchiname
-  annotate(geom = "text", y = 13.5, x = -88.7, label = "NF", size = 4)+
-  annotate(geom = "text", y = 13.35, x = -88.72, label = "*", size = 8)+
-  annotate(geom = "segment", y = 13.35, yend = 13.5,
-           x = -88.72, xend = -88.7)+
+  annotate(geom = "text", y = 13.45, x = -88.64, label = "NF", size = 4)+
+  annotate(geom = "point", y = 13.33, x = -88.72, size = 5, shape = 21, fill = "lightgrey")+
   ylab("Latitude")+
   xlab("Longitude")+
   theme_bw()

### 2021-01-28 programs/g_graphing_results_global.R
map drafted

@@ -265,11 +265,11 @@ range(ElSalvador$lat)
 #' Map of protected areas
 #+ map, fig.width = 5.67, dpi = 600, fig.height = 4.4
 ggplot(data = ElSalvador, aes(x = long, y = lat, group = group))+
-  geom_polygon(fill = "grey", color = "darkgrey")+
+  geom_polygon(fill = "grey", color = "#848484")+
   geom_polygon(data = Honduras, aes(x = long, y = lat, group = group),
-               fill = "lightgrey", color = "darkgrey")+
+               fill = "lightgrey", color = "#848484")+
   geom_polygon(data = Guatemala, aes(x = long, y = lat, group = group),
-               fill = "lightgrey", color = "darkgrey")+
+               fill = "lightgrey", color = "#848484")+
   coord_map(xlim = c(-90.2, -87.2), ylim = c(13.0, 14.5))+
   # El Imposible
   annotate(geom = "text", y = 13.74, x = -89.8, label = "EINP", size = 4)+
@@ -280,6 +280,12 @@ ggplot(data = ElSalvador, aes(x = long, y = lat, group = group))+
   # Nancuchiname
   annotate(geom = "text", y = 13.45, x = -88.64, label = "NF", size = 4)+
   annotate(geom = "point", y = 13.33, x = -88.72, size = 5, shape = 21, fill = "lightgrey")+
+  # Country & Ocean names
+  annotate(geom = "text", y = 13.8, x = -89.0, label = "EL SALVADOR", size = 3)+
+  annotate(geom = "text", y = 14.2, x = -90.0, label = "GUATEMALA", size = 3)+
+  annotate(geom = "text", y = 14.2, x = -87.8, label = "HONDURAS", size = 3)+
+  annotate(geom = "text", y = 13.3, x = -89.5, label = "Pacific Ocean", size = 2.5,
+           fontface = "italic")+
   ylab("Latitude")+
   xlab("Longitude")+
   theme_bw()

### 2021-01-31 manuscript_working/ms_working_RMD.Rmd
put map into manuscript
updated map legend in ms
update format: table 1

--- 2021-01-20 manuscript_working/ms_working_RMD.Rmd
@@
 library(pander)
-Habitat <- c("Reforested Deciduous Forest",
-             "Semi-Deciduous Forest",
-             "Cloud Forest",
-             "Pine/Oak Forest",
-             "Alluvial Forest with Secondary Growth",
-             "Alluvial Forest")
-Latitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
-Longitude <- c("xxx", "XXX", "XXX", "XXX", "XXX", "XXX")
-Elevation <- c("xxx", "xxx", "xxx", "xxx", "xxx", "xxx")
+Habitat <- c("Reforested deciduous forest",
+             "Semi-deciduous forest",
+             "Cloud forest",
+             "Pine/oak forest",
+             "Alluvial forest with secondary growth",
+             "Alluvial forest")
+Latitude <- c("13xx 50.75'", "13xx 51.38'", 
+              "14xx 24.88'", "14xx 23.19'", 
+              "13xx 19.95'", "13xx 22.03'")
+Longitude <- c("89xx 59.74'", "89xx 58.24'", 
+               "89xx 21.63'", "89xx 22.33'", 
+               "88xx 43.63'", "88xx 43.30'")
+Elevation <- c("609", "958", "2111", "1755", "3", "2")
 Route <- c("EI-1", "EI-2", "M-1", "M-2", "N-1", "N-2")
@@
 Map of El Salvador.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
-\label{psi_means}](../output/figures/psi_means-12.png)
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EINP), Montecristo National Park (MNP) and Nancuchiname Forest (NF).
+\label{map}](../output/figures/map-1.png)
 

### 2021-02-27 manuscript_working/ms_working_RMD.Rmd
incorporated MRE/AAA edits

@@
 --- 
-title: "Owl occupancy in three of El Salvador's protected areas from 2003 to 2013"
+title: "Owl Occupancy in Three of El Salvador's Protected Areas from 2003 through 2013"
 author: "Jane West and Althea Archer"
-date: "1/19/2021"
+date: "2/27/2021"
 output:
@@
 
-Owls are a keystone species in many ecosystems, and as such are an important bioindicator of ecosystem health.
-Top predators such as owls affect communities through their top-down control on biodiversity and their presence or absence can cause trophic cascades.
-However, owls are also cryptic, nocturnal predators that are difficult to accurately census.
-The Mexican Spotted Owl (*Strix occidentalis*) and its sub-species have long been studied as an indicator species for the interactions between anthropogenic and environmental forces, demonstrating that even various sub-species of spotted owls have complex and often contrasting responses to environmental land use and management [e.g., @White:2013; @Wan:2018].
+As apex predators, owls are an important bioindicator of ecosystem health, though much of our understanding comes from temperate systems [@White:2013; @Wan:2018; @Buechley:2019].
+In the tropics, particularly in rain and cloud forests, owl studies are more difficult than in temperate regions, regardless of potentially denser populations [@Konig:1999], and neotropical owls are less well-studied than their temperate counterparts [Buechley:2019].
+Further, the neotropics were identified by @Buechley:2019 as a priority area for owl research, based on an extensive literature review that identified a relative lack of research attention and a high conservation risk for neotropical owls.
+As a result, we have limited understanding of neotropical owls’ distribution, ecological requirements, population dynamics, and reproductive behavior [@Clark:1978; @Enriquez:2006; @Perez-Leon:2017; @Rangel-Salazar:2017].
+This lack of information poses a challenge to proactive conservation, even as neotropical habitat experiences tremendous anthropogenic pressure and rapid, unpredictable responses to climate change [@Corlett:2012].
+In general, while our knowledge about the status and trends of neotropical owl populations is limited, we do know that owl abundance and distribution is decreasing in several regions where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
 
-Neotropical owls are less well-studied than temperate species, and we have limited understanding of their ecological requirements, population dynamics, reproductive behavior, or conservation status [@Enriquez:2006; @Clark:1978; @Perez-Leon:2017; @Rangel-Salazar:2017].
-Yet, neotropical owls are reliant upon ecosystems under tremendous anthropogenic pressure and rapid and unpredictable responses to climate change [@Corlett:2012].
-In the tropics, particularly in rain and cloud forests, owl studies are even more difficult than in temperate regions, regardless of denser populations [@Konig:1999].
-In general, while our knowledge about the conservation status of neotropical populations is limited, we do know that owl abundance and distribution is decreasing in several regions where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
+Like many owl species, neotropical owls are difficult to accurately census due to their cryptic nature and nocturnal habits.
+Vocalizations of crepuscular and cryptic animals such as owls are commonly leveraged in survey methodology [@Konig:1999].
+Owl vocalizations can be diagnostic and are often considered more practically attainable than visual identification.
+Owls vocalize to communicate with the same species, to threaten or provoke other species, and to delimit territory [@Johnsgard:1988].
 
-The vocalization of crepuscular and cryptic animals such as owls are often leveraged for surveying, and vocal patterns are often considered more important than coloration as an aid to identification [@Springer:1978; @Forsman:1983; @Konig:1999].
-*springer 1978 and forsman 1983 are not cited on previous ms draft xxx* Owls vocalize to communicate with the same species, threaten or provoke other species, or to delimit territory [@Johnsgard:1988].
-Spontaneous and provocation listening are two methods that researchers use with increasing frequency during acoustic surveys of nocturnal birds of prey [e.g., @Baumgardt:2019].
-Spontaneous or passive surveys are based off of listening for animal calls with no prior stimulus whereas provocation surveys are conducted by listening after a period of playing animal vocalizations or "broadcast calls." One benefit of auditory surveys, whether spontaneous or broadcast, is being able to identify multiple species during the same survey period.
-Adding in provocation to auditory surveys can increase the detection probability of owls, although intra- and inter-species interactions can play a role whether owls are more likely or less likely to call in response to a vocalization simulus [@Enriquez:1997;@Baumgardt:2019].
+Vocalization surveys for nocturnal birds of prey fall generally into two categories.
+Spontaneous or “passive” surveys involve listening for animal calls with no prior stimulus, whereas provocation or “broadcast” surveys are conducted by listening after a period of playing a recording of animal vocalizations [e.g., @Baumgardt:2019].
+The inclusion of broadcast calls in auditory surveys can increase the detection probability of owls, although intra- and inter-species interactions can play a role whether owls are more likely or less likely to call in response to a vocalization simulus [@Enriquez:1997; @Borges:2004; @Flesch:2007; @Baumgardt:2019].
 
-The objective of our study was to monitor population dynamics of El Salvador owls through time to determine both individual species' population dynamics and community composition.
-El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot [@Myers:2000], and we selected the three survey areas to represent ecosystem diversity across El Salvador.
-We set up two routes in each protected area to further represent the habitats diversity of this heterogeneous country.
-We conducted nighttime foot surveys using passive listening and broadcast calls during the breeding season for three weeks each year from 2003 through 2013.
+The objective of our study was to collect occupancy data on El Salvador owls through time to determine both the population dynamics of individual owl species and the overall community composition of owls in three targeted protected areas, El Imposible National Park, Montecristo National Park, and Nancuchiname Forest.
+El Salvador is located in the heart of the Mesoamerican Biodiversity Hotspot [@Myers:2000], and we selected three survey areas in protected landscapes to represent the breadth of ecosystem diversity across the country.
+We set up two routes in each protected area and conducted nighttime foot surveys using passive listening and broadcast calls during the breeding season for three weeks during each year from 2003 through 2013.
 
@@
 
-El Salvador is the smallest and most densely populated Central American country with 13.6% of its land forested [@CIA:2020].
+El Salvador is located on the western side of the Central American isthmus and is the smallest and most densely populated Central American country  [@CIA:2020].
 It is bordered by the Pacific Ocean, Guatemala and Honduras (Fig.
 1XXX).
-We conducted surveys in three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that are representative of the country's diverse ecosystems (Fig.
+Overall, 13.6% of El Salvador's 20,720 km$^2$ of land is forest and 82% is used for agricultural purposes [@CIA:2020].
+We conducted surveys in three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that represent a portion of the country's diverse ecosystems (Fig.
 1xxx, Table 1XXX).
@@
 1XXX, Table 1XXX).
-Its elevation ranged from 250 to 1425 m above sea level.
-It was the largest NP in the country, covering 3792 ha.
-The topography was extremely steep and broken, with many cliffs [@Alvarez:2003].
+Its elevation ranged from 250 to 1425 m above sea level (a.s.l.).
+It was the largest NP in the country, covering 3792 ha, and its elevation ranged from 250 to 1425 m a.s.l..
+The topography of El Imposible was extremely steep and broken, with many cliffs [@Alvarez:2003].
 The area contained deciduous and semi-deciduous forest, secondary growth vegetation on the edges, and former pasture areas that were reforested with native species in 1997.
-We set up one survey route (EI-1) in the western portion of El Imposible in an area of reforestation and at a low elevation.
-We set up the other survey route (EI-2) in an area of semi-deciduous forest.
+We established one survey route (EI-1) in the western portion of El Imposible in an area of reforestation and at a low elevation and a second survey route (EI-2) in an area of semi-deciduous forest.
 
-Montecristo NP was located 125 km northwest of San Salvador (Fig.
+Montecristo NP (1973 ha) was located 125 km northwest of San Salvador (Fig.
 1XXX; Table 1XXX).
-Its altitude ranged from 730 to 2418 m above sea level, and the NP covered 1973 ha [@MARN:2010].
-The higher elevations contained cloud forest, middle elevations contained pine/oak forest (with some pine and cypress plantations) and the lower elevations contained a semi-deciduous tropical forest [@MARN:2015b].
+Its altitude ranged from 730 to 2418 m a.s.l.
+[@MARN:2010].
+High elevations of Montecristo contained cloud forest, middle elevations contained pine/oak forest (with some pine and cypress plantations) and the lower elevations contained a semi-deciduous tropical forest [@MARN:2015b].
 This NP was part of the Montecristo Tri-national Protected Area which included extensive adjoining natural areas in Guatemala and Honduras [@Komar:2010].
-In this NP, one survey route (M-1) was set up in a cloud forest and another (M-2) was located at a moderate elevation in a pine/oak forest.
+Within Montecristo, we established one survey route (M-1) in a cloud forest and another (M-2) at a moderate elevation in a pine/oak forest.
 
-Nancuchiname Forest, a Protected Natural Area, was located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
+Nancuchiname Forest (797 ha) is a Protected Natural Area located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
 1XXX).
-Its elevation ranged from 5 to 12 m above sea level and the protected area covered 797 ha.
-Nancuchiname Forest represented the largest remaining patch of tropical forest on the alluvial plain of the country [@Restrepo:2001]xxx **we can't site a geocities website**.
-The area contained deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike.
-In Nancuchiname Forest, we set up one survey route (N-1) in alluvial forest along the dike, which was surrounded by secondary growth, and we set up the other route (N-2) in a semi-deciduous alluvial forest.
+Its elevation ranged from sea level to 12 m a.s.l.
+The area contained deciduous and semi-deciduous forest, second growth vegetation, native species reforestation and a dike surrounded by secondary growth.
+In Nancuchiname Forest, we established one survey route (N-1) in alluvial forest along the dike and the other route (N-2) in a semi-deciduous alluvial forest.
 
@@
 
-We conducted foot surveys following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and other raptor research guidelines [@Fuller:1987].
+We conducted acoustic surveys on foot following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and other raptor research guidelines [@Fuller:1987].
 This study's survey protocol was an extension of previous research completed in the western half of El Imposible NP [@West:1988].
-For our surveys, we created six 2-km long survey routes (transects) that each included 10 permanent survey stations set at 200 m intervals.
+We created six 2-km long survey routes (transects) that each included 10 permanent survey stations set at 200 m intervals.
 Two routes were established in each protected area (Table 1xxx).
@@
 At each station, we measured and recorded environmental conditions (Brunton Sherpa, Brunton Incorporated) including temperature, precipitation, cloud cover, fog cover, wind speed, barometric pressure, moon phase, and noise level.
-The survey then began with passive listening for two minutes, followed by a three-minute broadcast call, and then seven minutes of silent listening.
+The acoustic survey then began with passive listening for two minutes, followed by a three-minute broadcast call, and ending with seven minutes of silent listening.
 We recorded local owl vocalizations and used them for broadcast calls whenever possible.
 If an owl vocalized during broadcast playback, we stopped the broadcast and recorded the owl's vocalization.
-When we detected an owl, we noted its  location and angle relative to the station, and we tracked when the same owl was identified at consecutive stations.
+When we detected an owl, we noted its  location and bearing relative to the station, and we tracked when the same owl was identified at consecutive stations.
+
 The broadcast calls varied between routes, but were consistent at each route and station across years.
-Specifically, we broadcasted the vocalizations of Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, Spectacled Owls, Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible NP (EI-1 and EI-2) and Nancuchiname Forest (N-1 and N-2).
-We broadcasted the vocalizations of Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, Great Horned Owls, Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo NP (M-1 and M-2).
+Specifically, we broadcast the vocalizations of Pacific Screech-Owls (*Megascops cooperi*), Mottled Owls (*Ciccaba virgata*), Crested Owls (*Lophostrix cristata*), Black-and-white Owls (*C.
+nigrolineata*), Spectacled Owls (*Pulsatrix perspicillata*), Pacific Screech-Owls, Mottled Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at stations 1 through 10, respectively, in all routes in El Imposible NP (EI-1 and EI-2) and Nancuchiname Forest (N-1 and N-2).
+We broadcast the vocalizations of Whiskered Owls (*M.
+trichopsis*), Mottled Owls, Fulvous Owls (*Strix fulvescens*), Stygian Owls (*Asio stygius*), Great Horned Owls (*Bubo virginianus*), Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo NP (M-1 and M-2).
+
 Surveys were repeated at each route up to three times a year from 2003 through 2013, depending on site access and weather.
@@
 2XX).
-We found that the probability of occupancy for Mottled Owls was relatively consistent across time in the El Imposible routes and in N-2 (Fig.
+We found that the probability of occupancy for Mottled Owls was relatively consistent across time in both El Imposible routes and in N-2 (Fig.
 3XX), but that it seemed to increase in the last approximately 6 years in M-2 and N-1 (Fig.
@@
 
-When analyzed by year, median species richness ($\text{Richness}^*_{t,h}$) varied from 0 to 5, depending on route and year (Fig.
+When analyzed by year and route, median species richness ($\text{Richness}^*_{t,h}$) varied from 0 to 5 (Fig.
 4XXX).
@@
 
-Including all detected species into one richness model increased the precision of the estimates of the effect of broadcast calls on detection probability (i.e., compare All Species with individual occupancy results in Fig.
+Including all detected species in one richness model increased the precision of the estimates of the effect of broadcast calls on detection probability (i.e., compare All Species with individual occupancy results in Fig.
 5XXX).
@@
 
-The goal of this long-term project was to complete two survey passes at each route for each year from 2003 through 2013.
-However, that was not always possible due to weather conditions, such as strong winds or rain.
-Access to the routes also precluded us from surveys in some instances.
-Rain and high winds diminish the vocalizations of owls as well as the ability for surveyors to detect any owl calls [@Takats:2001; @Andersen:2007].
-The survey was initially set up for 10 years based on recommendations from Mexican Spotted Owls researchers who stated that a monitoring period of 10 years would provide adequate information about population trends [@USFWS:1995].
+The goal of this project was to gather baseline data on owl abundance, distribution, and population trends in El Salvador.
+We established survey routes within three protected natural areas in disparate ecological regions (alluvial plain, forested, cloud forests) and attempted to compete two surveys of each route for each year from 2003 through 2013.
+Our goal of ten years’ worth of repeated surveys was based on similar studies that have previously been conducted on Mexican spotted owls (*Strix occidentalis*) [@USFWS:1995].
+However, our surveys were not always completed twice per year due to poor weather, issues with site access, or other complications.
+We did not survey during significant weather events, as rain and high winds diminish owl vocalizations and the ability for surveyors to detect owl calls [@Takats:2001; @Andersen:2007].
+Overall, our surveys yielded sufficient data to analyze occupancy patters for three species of interest; the Mottled Owl, Ferruginous Pygmy-owl, and Spectacled Owl.
 
-An estimated 75% of the nearly 250 surviving species of owls are associated with dense and undisturbed forests [@Mikkola:2012].
-@Perez-Leon:2017 stated that within El Salvador, natural ecosystems were the most diverse, specifically natural habitats of cloud, deciduous, riparian and pine-oak forests.
-In El Salvador, 13.6% of its land is forested [@CIA:2020].
-The country has continued with the traditional shade-growing coffee farming, which helps maintain forest cover.
-Coffee plantations account for around 7% of its forested land because of the intensive use of shade-cover farming practices [@Silva:2016].
-@Silva:2016 also stated that almost all primary or closed forest is surrounded by coffee farms, which act as a buffer zone against land use change.
-@Perez-Leon:2017 stated that coffee plantations are the most important agroecosystem for owl communities’ survival in El Salvador.
-
-In Central America, the Mottled Owl is the most common and widespread owl [@Vallely:2018].
-@Perez-Leon:2017 stated that habitat loss has restricted the distribution of several owl species, but that populations of Barn Owls, Pacific Screech-owls, Mottled Owls, and Ferruginous Pygmy-owls have survived the fragmentation processes in El Salvador relative to other Neotropical species.
-Contrarily, @Enriquez:2001 stated that Mottled Owl populations in La Selva may have decreased during the past 30 years.
-The results of this study indicated that Mottled Owls are still fairly common at each of the protected areas, although there seemed to be a downward trend in occupancy in M-2 (Fig 3XX).
+The Mottled Owl is the most common and widespread owl in South America [@Vallely:2018] and they have been documented in a wide range of habitat types and elevations in Mexico [@Enriquez-Rocha:1993].
+Mottled Owls were the most frequently detected owl during our surveys.
+A recent study indicated that Mottled Owl populations in El Salvador have persisted despite habitat loss that has restricted the distribution of several other owl species [@Perez-Leon:2017].
+In contrast, @Enriquez:2001 stated that Mottled Owl populations in La Selva may have decreased during the past 30 years.
+Our results indicated that Mottled Owls are still relatively common at each of our study areas, although there seemed to be a downward trend in occupancy in M-2 that could warrant further monitoring.
 
 Ferruginous Pygmy-owls were the second most-often detected owls in this study.
-Ferruginous Pygmy-owls were often detected at the beginning of survey routes, often closer to sunset than later at night, and Ferruginous Pygmy-owls demonstrated fairly low occupancy rates in El Imposible routes, other than in the first couple years of the survey in EI-1.
-Ferruginous Pygmy-owls prefer open habitats with a lot of edge.
-@West:1988 observed that the lower elevation areas of El Imposible had been grazed and dotted with houses.
-The activities and vegetation structure of the area has changed overtime, and native species re-vegetation efforts began in that area in 1997.
-It is possible that Ferruginous Pygmy-owls are responding to the decrease in edge habitat in this park, and we recommend further surveys to determine the risk to Ferruginous Pygmy-owls in response to revegetation management.
+This species demonstrated fairly low occupancy rates in El Imposible routes, other than in the first two years in EI-1.
+Ferruginous Pygmy-owls tend to prefer open habitats with a lot of edge, such as the flooded igap\'o forests of Amazonia [@Borges:2004] and heterogenous forests in central Argentina [@Campioni:2013], and their dietary composition is dependent on forest density [@Sarasola:2014].
+Lower elevation areas of El Imposible were grazed and dotted with houses in the late 1980s [@West:1988], and this mixed agrarian landscape may have provided suitable habitat which has been reduced by subsequent reforestation efforts that began in 1997.
+It is possible that Ferruginous Pygmy-owls are responding to a decrease in edge habitat within El Salvador, and we recommend further surveys to determine the risk to Ferruginous Pygmy-owls in response to revegetation management.
 
-Stygian Owl was listed as an expected owl in El Salvador [@Perez-Leon:2017).
+Changes in habitat availability are not the only challenge facing El Salvador's owls.
+@Perez-Leon:2017 stated that there are human activities affecting owl populations in El Salvador, including illegal hunting, trapping, persecution, killing, and wildlife trade.
+They also stated that, of eight species of owls confiscated from local markets in 1995 to 2008, the most frequent were Mottled Owls and Ferruginous Pygmy-owls.
+If poaching of these species continues, their populations may be at more risk than our results indicated.
+
+Spectacled Owls followed similar occupancy patterns to that of the Ferruginous Pygmy-owls, with highest levels of occupancy in the riparian Nancuchiname Forest.
+Spectacled Owls can be found in a range of habitats including coffee plantations and urban parks [@Marin-Gomez:2017].
+Spectacled Owls tend to prefer lowland habitat, and they were reported to live in elevations below 900 m a.s.l.
+in Mexico [@Enriquez-Rocha:1993] and in lowland habitats in Ecuador below 1000 m a.s.l.
+[@Orihuela-Torres:2018]; however, they have been reportedly able to live up to 2600 m a.s.l.
+[@Marin-Gomez:2017].
+We detected Spectacled Owls in Montecristo National Park, transect M-2, which was on average 1755 m a.s.l., which is potentially near their preferred elevation range.
+Spectacled Owls have been reported to prefer mammalian prey over reptilian or arthropods [@Orihuela-Torres:2018] and preferentially hunt near fallen logs where small mammals often seek refuge [@Esclarski:2014], but we were unable to test for micro-habitat occupancy in our present study.
+Thus, further refinement over Spectacled Owl habitat selection could be explored in the future to better understand their ecology and distribution in the Neotropics and specifically El Salvador.
+
+Nancuchiname Forest was the richest in diversity of our three protected areas; however, many of the habitat specialist species were only detected in Montecristo NP (e.g., Whiskered Owl, Great Horned Owl, Fulvous Owl, and c.f.
+Stygian Owl).
+A Nancuchiname Forest Management Plan only listed one owl, Ferruginous Pygmy-owl, as being present [@Zepeda:1995], but we confirmed the presence of four additional owl species in that NP.
+Fulvous Owl is one of El Salvador's 18 endemic bird species [@Komar:2002], and it was detected numerous times in the cloud forest of Montecristo NP (M-1).
+Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
+
+Stygian Owl, which often inhabit dense cloud forests [@Enriquez-Rocha:1993], were listed as an owl species expected to reside in El Salvador [@Perez-Leon:2017), but has been previously undetected in the country.
 On 26 March 2002, after listening to recordings and viewing photographs, the Montecristo NP guards stated that they thought c.f.
@@
 While there was only one c.f.
-Stygian Owl detected during the course of our surveys, we did also hear a c.f.
+Stygian Owl detected on our survey routes, we did also hear a c.f.
 Stygian Owl between the survey stations on 25 April XXXXyear.
 We also heard c.f.
-Stygian Owl twice on M-2 between station points on 20 March 2005 and once between stations in 2007.
+Stygian Owl vocalizations twice on M-2 between station points on 20 March 2005 and once between stations in 2007.
 We did not record any of these vocalizations by c.f.
-Stygian Owls.
-Detections that occurred between stations or away from survey routes were not included in our results.
+Stygian Owls and detections that occurred between stations or away from survey routes were not included in our results.
 So while we are confident that c.f.
@@
 
-Nancuchiname Forest was the richest in diversity of our three protected areas; however, many of the more specialist species were only detected in Montecristo NP (e.g., Whiskered Owl, Great Horned Owl, Fulvous Owl, and c.f.
-Stygian Owl).
-The 2002 Nancuchiname Forest Management Plan only listed one owl, Ferruginous Pygmy-owl, as being present [@Zepeda:1995]xx *why does year not match**, but we confirmed the presence of four additional owl species in that NP.
-Fulvous Owl is one of El Salvador's 18 endemic bird species [@Komar:2002], and it was detected numerous times in the cloud forest of Montecristo NP (M-1).
-Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
+Distinguishing individuals by non-invasive means, such as vocal traits, is preferable when species are rare, sensitive to handling, elusive, or when other techniques are unfeasible [@Terry:2005].
+Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally-recorded owl vocalizations when possible.
+In Guatemala, @Gerhardt:1989 found that a series of Mottled Owl vocalizations obtained from a non-local ornithological laboratory could not generate a response from Guatemalan owls.
+We found that the probability of detecting most owls increased after we broadcasted the vocalizations of Mottled Owls, Pacific Screech-owls, Crested Owls, and Spectacled Owls.
+Conversely, owls were less likely to respond after we played Great Horned Owl calls.
+Because owls, like many birds of prey, are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019] and may change their response to broadcast calls depending on their life stage (e.g., nesting versus courtship) [@Flesch:2007], we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
 
-Distinguishing individuals by non-invasive means, such as vocal traits, is preferable when species are rare, sensitive to handling, elusive, or when other techniques are unfeasible [@Terry:2005].
-Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally recorded owl vocalizations when possible.
-In Guatemala, @Gerhardt:1989 found that a series of vocalizations obtained from a non-local ornithological laboratory, which were identified as vocalizations of the Mottled Owl, could not generate a response from Guatemalan owls.
-We found that the probability of detecting most owls increased after we broadcasted the vocalizations of Mottled Owls, Pacific Screech-owls, Crested Owls, and Spectacled Owls.
-On the other hand, owls were less likely to call or be detected after we played Great Horned Owl calls.
-Because owls, like many birds of prey, are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019], we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
+Globally, an estimated 75% of the nearly 250 species of owls are associated with dense and undisturbed forests [@Mikkola:2012].
+@Perez-Leon:2017 stated that within El Salvador, natural ecosystems had the most diverse owl populations, specifically habitats comprised of cloud, deciduous, riparian and pine-oak forests.
+In addition to areas that are preserved in a natural state, shade-grown coffee production may be providing valuable owl habitat.
+Coffee plantations account for around 7% of El Salvador's forested land due to a prevalence of shade-cover farming practices  [@Silva:2016].
+@Silva:2016 also concluded that almost all primary or closed forest is surrounded by coffee farms, which act as a buffer zone against land use change.
+Likewise, @Perez-Leon:2017 stated that coffee plantations are the most important agroecosystem for owl communities’ survival in El Salvador.
 
-We did not use spectogram analysis to determine any of the specific individuals of owls that we detected.
-Spectograms of acoustic signals, such as owl vocalizations, are capable of not only recognizing differences between species, but even potentially individual owls of the same species or owl gender [@Eakle:1989; @Farquahar:1993; @Galeotii:1993; @Rognan:2009].
-*xx can't find Eakle 1989, Farquahar 1993, Galeotii 1993* Identifying individuals within a population can improve census estimates and provide useful information on demographics, life history traits, and behavior--all of which often influence management decisions [@Terry:2005].
-Spectogram analysis could help future owl researchers in El Salvadar determine more accurate population sizes of the more rare species we detected, such as Whiskered Screech-owls, c.f.
-Stygian Owls, and Great Horned Owls.
+This research study was critical because of its long-term study of multiple owl species in an rapidly urbanizing and population-dense country with diverse habitats and relatively little previous research.
+As we continue to identify areas of the world that are owl research priorities based on the lack of current information leveraged against the potential conservation risk, such as in @Buechley:2019, studies such as ours will be important and hopefully increasing in frequency.
 
-@Perez-Leon:2017 stated that there are human activities affecting owl populations in El Salvador, including illegal hunting, trapping, persecution, killing, and wildlife trade.
-They also stated that, of eight species of owls confiscated from local markets in 1995 to 2008, the most frequent were Mottled Owls and Ferruginous Pygmy-owls.
+# ACKNOWLEDGEMENTS
+
+Thank you to the Ministerio de Medio Ambiente y Recursos Naturales (MARN) for their authorization to conduct scientific research in El Salvador’s protected natural areas (permits were obtained annually starting in 2002: AIMA MARN-DGPN 006-2002 and ending in 2013: MARN AIMA-005-2013).
+Thank you to: MARN and Salva Natura for their employee’s assistance in San Salvador and in the field; to the MARN for logistics support and the GIS layers; to Dr.
+David Anderson for his technical review; especially thanks to Alfonso Sermeño Martínez, Vidal Campos Aguirre, Charles Starner, our families, and the US Fish and Wildlife Service for their assistance throughout the years of this project.
+Author contributions: JNW conceived of, developed, and conducted the owl surveys; AAA conducted the statistical analyses.
+AAA wrote the manuscript with support from JNW.
 
@@
 ![2XX.
-Probability of occupancy ($\psi_{t,h}$) averaged by route for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
@@
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
-Posteriors were summarized with median (black dot) $\pm$ 90% credible intervals (dark grey shading and vertical black lines).
-The number of owl species detected at each route and year is indicated with the light grey area.
+Posteriors of Richness* were summarized with median (black dot) $\pm$ 90% credible intervals (dark grey shading and vertical black lines).
+The number of owl species detected at each route and year was indicated with the light grey area.
 \label{figure5}](../output/figures/richness_byRtYr-1.png)
@@
 Posteriors were summarized with median $\pm$ 90% credible intervals.
-\label{figure6}](../output/figures/p_detection-1.png)
+The pre-broadcase effect size's credible intervals are represented with two grey horizontal lines on each graph \label{figure6}](../output/figures/p_detection-1.png)
 

### 2021-02-27 manuscript_working/ms_working_RMD.tex
incorporated MRE/AAA edits

@@
-\documentclass[]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Owl Occupancy in Three of El Salvador's Protected Areas from 2003 through 2013},
+  pdfauthor={Jane West and Althea Archer},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
-            pdftitle={working owls methods},
-            pdfauthor={Althea Archer},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 
-\title{working owls methods}
-\author{Althea Archer}
-\date{11/3/2020}
+\title{Owl Occupancy in Three of El Salvador's Protected Areas from 2003
+through 2013}
+\author{Jane West and Althea Archer}
+\date{2/27/2021}
 
@@
 
-\section{Methods}\label{methods}
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
 
-\subsection{Occupancy Model}\label{occupancy-model}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Let
+As apex predators, owls are an important bioindicator of ecosystem
+health, though much of our understanding comes from temperate systems
+(White et al. 2013, Wan et al. 2018, Buechley et al. 2019). In the
+tropics, particularly in rain and cloud forests, owl studies are more
+difficult than in temperate regions, regardless of potentially denser
+populations (König and Weick 1999), and neotropical owls are less
+well-studied than their temperate counterparts {[}Buechley:2019{]}.
+Further, the neotropics were identified by Buechley et al. (2019) as a
+priority area for owl research, based on an extensive literature review
+that identified a relative lack of research attention and a high
+conservation risk for neotropical owls. As a result, we have limited
+understanding of neotropical owls' distribution, ecological
+requirements, population dynamics, and reproductive behavior (Clark et
+al. 1978, Enríquez et al. 2006, Pérez-Léon et al. 2017, Rangel-Salazar
+and Enríquez 2017). This lack of information poses a challenge to
+proactive conservation, even as neotropical habitat experiences
+tremendous anthropogenic pressure and rapid, unpredictable responses to
+climate change (Corlett 2012). In general, while our knowledge about the
+status and trends of neotropical owl populations is limited, we do know
+that owl abundance and distribution is decreasing in several regions
+where species have been added to endangered species lists or have become
+locally extirpated (Enríquez et al. 2006).
+
+Like many owl species, neotropical owls are difficult to accurately
+census due to their cryptic nature and nocturnal habits. Vocalizations
+of crepuscular and cryptic animals such as owls are commonly leveraged
+in survey methodology (König and Weick 1999). Owl vocalizations can be
+diagnostic and are often considered more practically attainable than
+visual identification. Owls vocalize to communicate with the same
+species, to threaten or provoke other species, and to delimit territory
+(Johnsgard 1988).
+
+Vocalization surveys for nocturnal birds of prey fall generally into two
+categories. Spontaneous or ``passive'' surveys involve listening for
+animal calls with no prior stimulus, whereas provocation or
+``broadcast'' surveys are conducted by listening after a period of
+playing a recording of animal vocalizations (e.g., Baumgardt et al.
+2019). The inclusion of broadcast calls in auditory surveys can increase
+the detection probability of owls, although intra- and inter-species
+interactions can play a role whether owls are more likely or less likely
+to call in response to a vocalization simulus (Enrı́quez and Salazar
+1997, Borges et al. 2004, Flesch and Steidl 2007, Baumgardt et al.
+2019).
+
+The objective of our study was to collect occupancy data on El Salvador
+owls through time to determine both the population dynamics of
+individual owl species and the overall community composition of owls in
+three targeted protected areas, El Imposible National Park, Montecristo
+National Park, and Nancuchiname Forest. El Salvador is located in the
+heart of the Mesoamerican Biodiversity Hotspot (Myers and Mittermeier
+2000), and we selected three survey areas in protected landscapes to
+represent the breadth of ecosystem diversity across the country. We set
+up two routes in each protected area and conducted nighttime foot
+surveys using passive listening and broadcast calls during the breeding
+season for three weeks during each year from 2003 through 2013.
+
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{study-area}{%
+\subsection{Study area}\label{study-area}}
+
+El Salvador is located on the western side of the Central American
+isthmus and is the smallest and most densely populated Central American
+country (Central Intelligence Agency 2020). It is bordered by the
+Pacific Ocean, Guatemala and Honduras (Fig. 1XXX). Overall, 13.6\% of El
+Salvador's 20,720 km\(^2\) of land is forest and 82\% is used for
+agricultural purposes (Central Intelligence Agency 2020). We conducted
+surveys in three protected natural areas located at different elevations
+(high, mid and low elevations) and in different types of forest
+vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud
+forest) that represent a portion of the country's diverse ecosystems
+(Fig. 1xxx, Table 1XXX).
+
+El Imposible National Park (NP) was located 119 km southwest of San
+Salvador (Fig. 1XXX, Table 1XXX). Its elevation ranged from 250 to 1425
+m above sea level (a.s.l.). It was the largest NP in the country,
+covering 3792 ha, and its elevation ranged from 250 to 1425 m a.s.l..
+The topography of El Imposible was extremely steep and broken, with many
+cliffs (Álvarez and Komar 2003). The area contained deciduous and
+semi-deciduous forest, secondary growth vegetation on the edges, and
+former pasture areas that were reforested with native species in 1997.
+We established one survey route (EI-1) in the western portion of El
+Imposible in an area of reforestation and at a low elevation and a
+second survey route (EI-2) in an area of semi-deciduous forest.
+
+Montecristo NP (1973 ha) was located 125 km northwest of San Salvador
+(Fig. 1XXX; Table 1XXX). Its altitude ranged from 730 to 2418 m a.s.l.
+(Ministerio de Medio Ambiente y Recursos Naturales 2010). High
+elevations of Montecristo contained cloud forest, middle elevations
+contained pine/oak forest (with some pine and cypress plantations) and
+the lower elevations contained a semi-deciduous tropical forest
+(Ministerio de Medio Ambiente y Recursos Naturales 2015a). This NP was
+part of the Montecristo Tri-national Protected Area which included
+extensive adjoining natural areas in Guatemala and Honduras (Komar
+2010). Within Montecristo, we established one survey route (M-1) in a
+cloud forest and another (M-2) at a moderate elevation in a pine/oak
+forest.
+
+Nancuchiname Forest (797 ha) is a Protected Natural Area located 95 km
+southeast of San Salvador in the coastal alluvial plain along the Lempa
+River (Fig. 1XXX). Its elevation ranged from sea level to 12 m a.s.l.
+The area contained deciduous and semi-deciduous forest, second growth
+vegetation, native species reforestation and a dike surrounded by
+secondary growth. In Nancuchiname Forest, we established one survey
+route (N-1) in alluvial forest along the dike and the other route (N-2)
+in a semi-deciduous alluvial forest.
+
+\emph{Table 1XX. Route locations and habitats. Owl surveys were
+conducted from 2003 to 2013 in three different protected areas within El
+Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo
+National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2). }
+
+\begin{longtable}[]{@{}ccccc@{}}
+\toprule
+\begin{minipage}[b]{0.08\columnwidth}\centering
+Route\strut
+\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\centering
+Latitude\strut
+\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\centering
+Longitude\strut
+\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\centering
+Elevation\strut
+\end{minipage} & \begin{minipage}[b]{0.35\columnwidth}\centering
+Habitat\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.08\columnwidth}\centering
+EI-1\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+13xx 50.75'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+89xx 59.74'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+609\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Reforested deciduous forest\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\centering
+EI-2\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+13xx 51.38'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+89xx 58.24'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+958\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Semi-deciduous forest\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\centering
+M-1\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+14xx 24.88'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+89xx 21.63'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+2111\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Cloud forest\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\centering
+M-2\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+14xx 23.19'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+89xx 22.33'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+1755\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Pine/oak forest\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\centering
+N-1\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+13xx 19.95'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+88xx 43.63'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+3\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Alluvial forest with secondary growth\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.08\columnwidth}\centering
+N-2\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+13xx 22.03'\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\centering
+88xx 43.30'\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\centering
+2\strut
+\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\centering
+Alluvial forest\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{survey-methods}{%
+\subsection{Survey methods}\label{survey-methods}}
+
+We conducted acoustic surveys on foot following the Guidelines for
+Nocturnaal Owl Monitoring (Takats et al. 2001), a similar protocol
+developed for Costa Rica (Enrı́quez and Rangel-Salazar 2001), and other
+raptor research guidelines (Fuller and Mosher 1987). This study's survey
+protocol was an extension of previous research completed in the western
+half of El Imposible NP (West 1988). We created six 2-km long survey
+routes (transects) that each included 10 permanent survey stations set
+at 200 m intervals. Two routes were established in each protected area
+(Table 1xxx).
+
+Surveys began at local twilight and took approximately five hours to
+complete. At each station, we measured and recorded environmental
+conditions (Brunton Sherpa, Brunton Incorporated) including temperature,
+precipitation, cloud cover, fog cover, wind speed, barometric pressure,
+moon phase, and noise level. The acoustic survey then began with passive
+listening for two minutes, followed by a three-minute broadcast call,
+and ending with seven minutes of silent listening. We recorded local owl
+vocalizations and used them for broadcast calls whenever possible. If an
+owl vocalized during broadcast playback, we stopped the broadcast and
+recorded the owl's vocalization. When we detected an owl, we noted its
+location and bearing relative to the station, and we tracked when the
+same owl was identified at consecutive stations.
+
+The broadcast calls varied between routes, but were consistent at each
+route and station across years. Specifically, we broadcast the
+vocalizations of Pacific Screech-Owls (\emph{Megascops cooperi}),
+Mottled Owls (\emph{Ciccaba virgata}), Crested Owls (\emph{Lophostrix
+cristata}), Black-and-white Owls (\emph{C. nigrolineata}), Spectacled
+Owls (\emph{Pulsatrix perspicillata}), Pacific Screech-Owls, Mottled
+Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls at
+stations 1 through 10, respectively, in all routes in El Imposible NP
+(EI-1 and EI-2) and Nancuchiname Forest (N-1 and N-2). We broadcast the
+vocalizations of Whiskered Owls (\emph{M. trichopsis}), Mottled Owls,
+Fulvous Owls (\emph{Strix fulvescens}), Stygian Owls (\emph{Asio
+stygius}), Great Horned Owls (\emph{Bubo virginianus}), Whiskered Owls,
+Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at
+stations 1 through 10, respectively, in both routes of Montecristo NP
+(M-1 and M-2).
+
+Surveys were repeated at each route up to three times a year from 2003
+through 2013, depending on site access and weather. We did not conduct
+any surveys in 2006. We georeferenced the owl detections for each survey
+and route to eliminate double counting of individual owl observations.
+
+\emph{Table 2XX. Species detection records by route. Owl surveys were
+conducted from 2003 to 2013 in three different protected areas within El
+Salvador, although specific survey years varied by route.}
+
+\begin{longtable}[]{@{}ccccccc@{}}
+\toprule
+\begin{minipage}[b]{0.31\columnwidth}\centering
+Name\strut
+\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\centering
+EI-1\strut
+\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\centering
+EI-2\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
+M-1\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
+M-2\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
+N-1\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
+N-2\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Barn Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+8\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+2\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Tyto alba)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Whiskered Screech-Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+2\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Megascops trichopsis)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Pacific Screech-Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+11\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+6\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+9\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Megascops cooperi)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Crested Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Lophostrix cristata)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Spectacled Owl**\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+59\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+75\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Pulsatrix perspicillata)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Great Horned Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+3\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Bubo virginianus)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Ferruginous Pygmy-Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+4\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+82\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+98\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Glaucidium brasilianum)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Burrowing Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Athene cunicularia)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Mottled Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+214\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+117\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+22\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+62\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+114\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Ciccaba virgata)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Black-and-white Owl**\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Ciccaba nigrolineata)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Fulvous Owl*\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+17\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Strix fulvescens)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+c.f. Stygian Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+1\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Asio stygius)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Striped Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Pseudoscops clamator)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+Unspotted Saw-whet Owl\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+0\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.31\columnwidth}\centering
+(Aegolius ridgwayi)\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
+\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\emph{Note: ** = Endangered and * = Threatened (Ministerio de Medio
+Ambiente y Recursos Naturales 2015b)}
+
+\hypertarget{single-species-occupancy-model-framework}{%
+\subsection{Single-species occupancy model
+framework}\label{single-species-occupancy-model-framework}}
+
+We modeled occupancy within each route, year, and survey assuming that
+the probability of occupancy would be closed across surveys of a given
+route in a given year. In other words, for any single species of owl, we
+assumed that the probability of a route being occupied would not change
+between surveys within a year but could vary between years. Let:
 
@@
 \begin{aligned}
-t  &=  \text{the years of surveys (here }t = 1,2,\ldots,11)\\
-h  &=  \text{the individual routes (here }h = 1,2,\ldots,6)\\
-i  &=  \text{the individual surveys conducted in each year and route (here }i = 1,2,3)\\
-j  &=  \text{the individual stations along each route (here }j = 1,2,\ldots,10)\\
-k  &=  \text{the broadcast species (here }k = 1,2,\ldots,10)\\
-z_{t,h,i} &= \text{a random variable equal to 1 when a survey was occupied and 0 otherwise}\\
+t  &=  \text{the year each survey was conducted (}t = 1,2,\ldots,11)\text{ representing 2003-2013}\\
+h  &=  \text{the individual route (}h = 1,2,\ldots,6)\text{ representing EI-1, EI-2, M-1, M-2, N-1, N-2}\\
+i  &=  \text{the individual surveys conducted in each year and route (}i = 1,2,3)\\
+j  &=  \text{the individual stations along each route (}j = 1,2,\ldots,10)\\
+k  &=  \text{the broadcast call species (}k = 1,2,\ldots,10)\\
+z_{t,h,i} &= \text{a random variable equal to 1 when a survey route was occupied and 0 otherwise}\\
 y_{t,h,i,j,k} &= \text{a random variable equal to 1 when an owl was detected and 0 otherwise}
@@
 
-We assumed that survey occupancy (\(z_{t,h,i}\)) was the outcome of
-Bernoulli trials with probability \(\psi\), which we allowed to vary by
-year and route:
+We assumed that occupancy in each individual survey (\(z_{t,h,i}\)) was
+the outcome of Bernoulli trials with a probability of occupancy,
+\(\psi_{t,h}\), which we allowed to vary by year and route:
 
@@
 We assumed that detecting an owl depended on an owl being present during
-that survey (\(z_{t,h,i} = 1\)) and the probability of detection, which
-was related to the broadcast species:
+that survey (\(z_{t,h,i} = 1\)) and the probability of detection,
+\(p_{t,h,i,j,k}\), which was related to the species-specific broadcast
+call (\(k\)):
 
@@
 
-where generally \(\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k\), or more
+where, generally, \(\text{logit}(p_{t,h,i,j,k}) = \beta_kX_k\), or more
 specifically:
@@
 
-This model provides a consistent probability of detection for all
-surveys in the first two minutes of each survey before the broadcast owl
-recording was played. Then, the probability of detection for each
-post-broadcast time period would depend on the species of broadcast owl.
-This allowed species-specific behavior in response to the different
-broadcast owl species (Baumgardt et al. 2019). We used means
-parameterization such that the coefficients were interpretable as the
-effect of that specific broadcast species \(k\) (including the
-pre-broadcast time frame). The broadcast species were consistent at each
-station across years, but varied by route:
-
-\begin{table}[]
-\centering
-\caption{Transformations Associated with the Johnson System}
-\begin{tabular}{|l|l|l|l|}
-\hline
-Johnson Family & Transformation & Parameter Conditions & X Condition \\ \hline
-$S_B$ & $Z=\gamma + \eta ln(\frac {X - \epsilon} {\lambda + \epsilon - X})$ & $\eta, \lambda >0, -\infty < \gamma, \epsilon < \infty$ & $\epsilon < X < \epsilon + \lambda$ \\ \hline
-$S_L$ & $Z=\gamma + \eta ln(X - \epsilon)$ & $\eta >0, -\infty < \gamma, \epsilon < \infty$ & $X > \epsilon$ \\ \hline
-$S_U$ & $Z=\gamma + \eta \sinh^{-1}(\frac {X - \epsilon} {\lambda})$ & $\eta, \lambda >0, -\infty < \gamma, \epsilon < \infty$ & $-\infty < X < \infty$ \\ \hline
-\end{tabular}
-\end{table}
-
-The priors for every logistic model coefficient \(\beta_k\) were chosen
-to be near uniform as recommended in Gelman et al. (2008). Specifically,
-we used the Cauchy prior as such:
+This model provided a consistent probability of detection for all
+surveys in the first two minutes of each observation period, before the
+broadcast call was played (\(k = \text{Pre-broadcast}\)). Then, the
+probability of detection for each post-broadcast time period depended
+which call had been played. This allowed species-specific behavior in
+response to the different broadcast calls (Baumgardt et al. 2019). We
+used means parameterization for the logistic regression model such that
+the coefficients were interpretable as the effect of that specific
+broadcast call \(k\) (including the pre-broadcast time frame). The
+priors for every logistic model coefficient \(\beta_k\) were chosen to
+be near uniform as recommended in Gelman et al. (2008):
 \(\beta_k \sim \text{Cauchy}(\text{precision} = 0.16)\).
 
-\subsubsection{Implementation of Occupany
-Model}\label{implementation-of-occupany-model}
+\hypertarget{richness-model-framework}{%
+\subsection{Richness model framework}\label{richness-model-framework}}
 
-We used the R2jags package in R (R Development Core Team 2014,Plummer
-(2013)) to implement the occupancy model for three owl species: Mottled,
-Spectacled, and FerPy owls. These owls had enough positive detections to
-analyze occupancy, as Mottled, Spectacled, and FerPy owls had 542, 137,
-and 187 positive detections over the 11 year period, respectively. Based
-on our understanding of owl ecology, we assumed that Spectacled and
-FerPy owls would not occupy route M1 in Montecristo, so we removed that
-route from those occupancy models.
+To model species richness at each route in each year, we assumed that
+there were 14 possible species present in El Salvador. This included the
+9 species observed during our surveys and an additional 5 species that
+we never observed (Table 2XX). This upper limit to expected species
+richness related directly to the 13 species known to inhabit El Salvador
+(Dickey and van Rossem 1938) plus one species we identified (c.f.
+Stygian owl), which was previously undocumented in El Salvador. We thus
+augmented our owl detection data with 5 additional potential species
+with all-zero detection records (Royle 2007).
 
-For all three species' occupancy models, we ran 3 chains for 10,000
-iterations and 1000 iterations discarded as burn-in, for a total of
-27,000 iterations comprising the posterior distributions for each model
-parameter. We visually inspected traceplots to verify that chains mixed
-well (Supplemental information S1-3xx).
+For each species \(s\) (where \(s = 1,2, \ldots,15\)), \(w_{h,s}\) was a
+binary indicator of whether that owl species was present in each route
+based on the probability of species presence in that route, \(\Omega_h\)
+(Broms et al. 2016):
 
-\section{Results}\label{results}
+\[
+w_{h,s} \sim \text{Bernoulli}(\Omega_h)
+\] We used prior distributions for \(\Omega_h\):
+\(\Omega_h \sim \text{Beta}(0.001, 1)\) (Broms et al. 2016,
+Guillera-Arroita et al. 2019).
 
-\subsection{Results sub-section}\label{results-sub-section}
+In the richness model, the presence of an owl species in any year,
+route, and survey, \(z_{t,h,i,s}\), was conditional on the probability
+of occupancy for that route, year, and species \(\psi_{t,h,s}\)
+\emph{and} that species belonging to that route's community that year
+(\(w_{h,s}\)):
 
-\section{Discussion}\label{discussion}
+\[
+z_{t,h,i,s} \sim \text{Bernoulli}(w_{h,s}*\psi_{t,h,s})
+\]
 
-You can cite references using the \textbf{citr} add-in. With
-\textbf{citr} you can cite papers in line like ({\textbf{???}}) is fond
-of doing, or parenthetically (R Development Core Team 2014). You can
-also cite multiple sources at once ({\textbf{???}}, {\textbf{???}},
-Plummer 2013). References can be added into the myrefs.bib file
-directly, or created with BibTeX through LaTeX.
+To share information about occupancy amongst species, we incorporated
+random effects into the parameterization of \(\psi_{t,h,s}\):
 
-\section{Figures}\label{figures}
+\[
+\begin{aligned}
+\psi_{t,h,s} &\sim \text{Beta}(a^\psi_{t,h,s},b^\psi_{t,h,s})\\ 
+a^\psi_{t,h,s} &= \mu^\psi_{t,h}\rho^\psi\\
+b^\psi_{t,h,s} &= \rho^\psi - \mu^\psi_{t,h}\rho^\psi\\
+\text{logit}(\mu^\psi_{t,h}) &\sim N(0,2.25)\\
+\ln(\rho^\psi) &\sim N(5,1)\text{, truncated to (0.01,10)}\\
+\end{aligned}
+\]
 
-Figures can be included with the following code, but they must be in
-.png format to show up in both Word and PDF. They can be referenced for
-the PDF with Fig. \ref{figure1}. \emph{Unfortunately, I don't yet know
-how to make the reference number show up in Word.}
+This parameterization allowed us to borrow strength across species in
+determining the mean probability of occupancy (\(\mu^\psi_{t,h}\)) for
+each route and year (Guillera-Arroita et al. 2019).
 
-\begin{figure}[htbp]
+Richness of each route was conditional on each species' presence and was
+estimated as the sum of owl species present in that route:
+
+\[
+\text{Richness}_{h} = \sum_{s=1}^{15} w_{h,s} 
+\]
+
+We also estimated year-to-year richness (\(\text{Richness}^*_{t,h}\)) by
+route by adding up the number of unique species that were present in
+each route, year, and survey. We first determined if a species was
+present in any of that route and year's set of surveys with a
+species-specific binary indicator \(\vartheta_{t,h,s}\) where
+\(\vartheta_{t,h,s} = 1\) if \(z_{t,h,i,s} = 1\) in any of that route's
+surveys \(i\) that year. Then, we summed up the number of species
+present in each year and route: \[
+\text{Richness}^*_{t,h} = \sum_{s=1}^{15} \vartheta_{t,h,s} 
+\]
+
+Consistent with our approach for the occupancy models, we assumed that
+detecting an owl depended on an individual of that species being present
+during that survey (\(z_{t,h,i,s} = 1\)) and the species-specific
+probability of detection \(p_{t,h,i,j,k,s}\), which was related to the
+broadcast call:
+
+\[
+y_{t,h,i,j,k,s} \sim \text{Bernoulli}(z_{t,h,i,s}*p_{t,h,i,j,k,s})
+\]
+
+where generally \(\text{logit}(p_{t,h,i,j,k,s}) = \beta_kX_k\) for each
+broadcast call \(k\) as described above. The coefficients of the
+detection model were thus assumed to be consistent for all owl species,
+which allowed us to borrow information about the detection process for
+undetected owl species from those that were detected.
+
+\hypertarget{model-implementation}{%
+\subsubsection{Model implementation}\label{model-implementation}}
+
+We used the R2jags package in R (R Development Core Team 2014) (Plummer
+2013) to implement the Bayesian models. We verified that R-hats were
+lower than 1.3 and visually inspected traceplots to verify that chains
+mixed well (Supplemental information S1-3xx).
+
+We applied the single-species occupancy model for three owl species that
+had sufficient positive detections for analysis: Mottled Owl, Spectacled
+Owl, and Ferruginous Pygmy-Owl (Table 2XX). Based on our understanding
+of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls
+would not occupy route M-1 in Montecristo, so we removed M-1 from those
+species' occupancy analyses. For all three of the single-species
+occupancy models, we ran 3 chains for 10,000 iterations and 1000
+iterations discarded as burn-in, for a total of 27,000 iterations
+comprising the posterior distributions for each model parameter. Initial
+values for \(z_{t,h,i}\) were equivalent to one if an owl was detected
+in that survey \(i\) and otherwise zero.
+
+To implement the richness model, we ran 3 chains for 20,000 iterations
+with 2000 iterations discarded as burn-in and a thinning rate of 2, for
+a total of 27,000 iterations comprising the posterior distributions for
+each model parameter. Initial values for \(z_{t,h,i,s}\) were equivalent
+to one if that species of owl was detected in that survey and otherwise
+zero. Similarly, initial values for whether a species belonged to a
+route's community (\(w_{h,s}\)) were equivalent to one if that species
+of owl was ever detected in that route and otherwise zero.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+A total of 86 surveys were conducted between March and May from 2003
+through 2005 and 2007 through 2013. No surveys were conducted in 2006.
+
+We detected nine species of owls during our surveys: Barn Owl, Whiskered
+Screech-Owl, Pacific Screech-Owl, Spectacled Owl, Great Horned Owl,
+Ferruginous Pygmy-Owl, Mottled Owl, Fulvous Owl, and c.f. Stygian Owl
+(Table 2xXX). We identified Whiskered Screech-Owls in Montecristo NP,
+which had previously been undocumented in that protected area. We also
+identified a c.f. Stygian Owl in Montecristo NP, which was a species not
+previously documented in the entire country of El Salvador. The c.f.
+Stygian Owl requires further investigation to determine true presence of
+that owl in El Salvador.
+
+\hypertarget{single-species-occupancy-model-results}{%
+\subsection{Single-species occupancy model
+results}\label{single-species-occupancy-model-results}}
+
+Mottled Owls, Spectacled Owls, and Ferruginous Pygmy-Owls had 542, 137,
+and 187 positive detections over the 11 year period, respectively (Table
+2XX). No other owl species had enough detections to analyze
+species-specific occupancy (Table 2XX). Averaged across years, we found
+that the probability of occupancy was somewhat higher for Mottled Owls
+than for Ferruginous Pygmy-owls or Spectacled Owls in both routes of El
+Imposible and in M-2 route of Montecristo (Fig. 1X). Occupancy
+probabilities were relatively high across all three species for both
+routes of Nancuchiname (Fig. 2XX). We found that the probability of
+occupancy for Mottled Owls was relatively consistent across time in both
+El Imposible routes and in N-2 (Fig. 3XX), but that it seemed to
+increase in the last approximately 6 years in M-2 and N-1 (Fig. 3XX).
+The probability of occupancy in M-1 seemed to have decreased through
+time for Mottled Owls (Fig. 3XX).
+
+For routes EI-1, EI-2, and M-2, the probability of occupancy for
+Ferruginous Pygmy-owls and Spectacled Owls stayed relatively low across
+time other than a higher probability of occupancy for Ferruginous
+Pygmy-owl in EI-1 in the early years of the surveys (Fig. 3XX). The
+probability of occupancy for Ferruginous Pygmy-owls was relatively high
+in both routes of Nancuchaname except in 2008 and 2009 (Fig. 3XX), and
+the probability of occupancy for the Spectacled Owl fluctuated slightly
+from year to year in the Nancuchaname routes (Fig. 3XX).
+
+\hypertarget{richness-model-results}{%
+\subsection{Richness model results}\label{richness-model-results}}
+
+Our estimates of each route's community richness
+(\(\text{Richness}_{h}\)) and 90\% credible intervals were 4 (4,4), 3
+(3,3), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2,
+N-1, and N-2, respectively. These median values were equivalent to the
+number of species detected in each route (Table 2XX), indicating that
+the probability of occupancy for undetected species was low and no
+additional species were predicted to be present in a route's community,
+other than those that were detected.
+
+When analyzed by year and route, median species richness
+(\(\text{Richness}^*_{t,h}\)) varied from 0 to 5 (Fig. 4XXX). At most,
+the median species richness was one species higher than the number of
+species detected at any single route and year; however, the 90\%
+credible intervals of richness added up to 4 new undetected species
+(e.g., in M-2 during 2004 where we detected zero species (Fig. 4XXX).
+
+\hypertarget{detection-results}{%
+\subsection{Detection results}\label{detection-results}}
+
+Including all detected species in one richness model increased the
+precision of the estimates of the effect of broadcast calls on detection
+probability (i.e., compare All Species with individual occupancy results
+in Fig. 5XXX). When incorporating all owl species, broadcasting calls
+from Mottled Owls, Pacific Screech-owls, Crested Owls, Black-and-white
+Owls, and Spectacled Owls increased detection probability; whereas
+broadcasting the Great Horned Owl calls decreased detection (Fig. 5XXX).
+
+The probability of detection when analyzed with single-species occupancy
+models varied little during the timeframe before broadcast recordings
+were played (Pre-broadcast); however, the probability of detecting owls
+after broadcast calls were played varied by call and between Mottled
+Owls, Ferruginous Pygmy-owls, and Spectacled Owls (Fig. 5XX). The
+probability of detecting Mottled Owls increased after broadcasting
+Mottled Owl, Pacific Screech-owl, Crested Owl, Black-and-white Owl, or
+Spectacled Owl calls. The probability of detecting Ferruginous
+Pygmy-owls increased after broadcasting Mottled Owl and Pacific Owl
+calls, and the probability of detecting Spectacled Owls increased after
+broadcasting Pacific Screech-owl, Black-and-white Owl, and Spectacled
+Owl calls (Fig. 5XX).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+The goal of this project was to gather baseline data on owl abundance,
+distribution, and population trends in El Salvador. We established
+survey routes within three protected natural areas in disparate
+ecological regions (alluvial plain, forested, cloud forests) and
+attempted to compete two surveys of each route for each year from 2003
+through 2013. Our goal of ten years' worth of repeated surveys was based
+on similar studies that have previously been conducted on Mexican
+spotted owls (\emph{Strix occidentalis}) (United States Fish and
+Wildlife Service 1995). However, our surveys were not always completed
+twice per year due to poor weather, issues with site access, or other
+complications. We did not survey during significant weather events, as
+rain and high winds diminish owl vocalizations and the ability for
+surveyors to detect owl calls (Takats et al. 2001, Andersen 2007).
+Overall, our surveys yielded sufficient data to analyze occupancy
+patters for three species of interest; the Mottled Owl, Ferruginous
+Pygmy-owl, and Spectacled Owl.
+
+The Mottled Owl is the most common and widespread owl in South America
+(Vallely and Dyer 2018) and they have been documented in a wide range of
+habitat types and elevations in Mexico (Enríquez-Rocha et al. 1993).
+Mottled Owls were the most frequently detected owl during our surveys. A
+recent study indicated that Mottled Owl populations in El Salvador have
+persisted despite habitat loss that has restricted the distribution of
+several other owl species (Pérez-Léon et al. 2017). In contrast,
+Enrı́quez and Rangel-Salazar (2001) stated that Mottled Owl populations
+in La Selva may have decreased during the past 30 years. Our results
+indicated that Mottled Owls are still relatively common at each of our
+study areas, although there seemed to be a downward trend in occupancy
+in M-2 that could warrant further monitoring.
+
+Ferruginous Pygmy-owls were the second most-often detected owls in this
+study. This species demonstrated fairly low occupancy rates in El
+Imposible routes, other than in the first two years in EI-1. Ferruginous
+Pygmy-owls tend to prefer open habitats with a lot of edge, such as the
+flooded igap'o forests of Amazonia (Borges et al. 2004) and heterogenous
+forests in central Argentina (Campioni et al. 2013), and their dietary
+composition is dependent on forest density (Sarasola and Santillán
+2014). Lower elevation areas of El Imposible were grazed and dotted with
+houses in the late 1980s (West 1988), and this mixed agrarian landscape
+may have provided suitable habitat which has been reduced by subsequent
+reforestation efforts that began in 1997. It is possible that
+Ferruginous Pygmy-owls are responding to a decrease in edge habitat
+within El Salvador, and we recommend further surveys to determine the
+risk to Ferruginous Pygmy-owls in response to revegetation management.
+
+Changes in habitat availability are not the only challenge facing El
+Salvador's owls. Pérez-Léon et al. (2017) stated that there are human
+activities affecting owl populations in El Salvador, including illegal
+hunting, trapping, persecution, killing, and wildlife trade. They also
+stated that, of eight species of owls confiscated from local markets in
+1995 to 2008, the most frequent were Mottled Owls and Ferruginous
+Pygmy-owls. If poaching of these species continues, their populations
+may be at more risk than our results indicated.
+
+Spectacled Owls followed similar occupancy patterns to that of the
+Ferruginous Pygmy-owls, with highest levels of occupancy in the riparian
+Nancuchiname Forest. Spectacled Owls can be found in a range of habitats
+including coffee plantations and urban parks (Marín-Gómez et al. 2017).
+Spectacled Owls tend to prefer lowland habitat, and they were reported
+to live in elevations below 900 m a.s.l. in Mexico (Enríquez-Rocha et
+al. 1993) and in lowland habitats in Ecuador below 1000 m a.s.l.
+(Orihuela-Torres et al. 2018); however, they have been reportedly able
+to live up to 2600 m a.s.l. (Marín-Gómez et al. 2017). We detected
+Spectacled Owls in Montecristo National Park, transect M-2, which was on
+average 1755 m a.s.l., which is potentially near their preferred
+elevation range. Spectacled Owls have been reported to prefer mammalian
+prey over reptilian or arthropods (Orihuela-Torres et al. 2018) and
+preferentially hunt near fallen logs where small mammals often seek
+refuge (Esclarski and Cintra 2014), but we were unable to test for
+micro-habitat occupancy in our present study. Thus, further refinement
+over Spectacled Owl habitat selection could be explored in the future to
+better understand their ecology and distribution in the Neotropics and
+specifically El Salvador.
+
+Nancuchiname Forest was the richest in diversity of our three protected
+areas; however, many of the habitat specialist species were only
+detected in Montecristo NP (e.g., Whiskered Owl, Great Horned Owl,
+Fulvous Owl, and c.f. Stygian Owl). A Nancuchiname Forest Management
+Plan only listed one owl, Ferruginous Pygmy-owl, as being present
+(Zepeda 1995), but we confirmed the presence of four additional owl
+species in that NP. Fulvous Owl is one of El Salvador's 18 endemic bird
+species (Komar 2002), and it was detected numerous times in the cloud
+forest of Montecristo NP (M-1). Fulvous Owl has also been recently
+photographed in Montecristo NP (Gonzalez et al. 2017).
+
+Stygian Owl, which often inhabit dense cloud forests (Enríquez-Rocha et
+al. 1993), were listed as an owl species expected to reside in El
+Salvador {[}Pérez-Léon et al. (2017)), but has been previously
+undetected in the country. On 26 March 2002, after listening to
+recordings and viewing photographs, the Montecristo NP guards stated
+that they thought c.f. Stygian owl was present in the park (personal
+communication). While there was only one c.f. Stygian Owl detected on
+our survey routes, we did also hear a c.f. Stygian Owl between the
+survey stations on 25 April XXXXyear. We also heard c.f. Stygian Owl
+vocalizations twice on M-2 between station points on 20 March 2005 and
+once between stations in 2007. We did not record any of these
+vocalizations by c.f. Stygian Owls and detections that occurred between
+stations or away from survey routes were not included in our results. So
+while we are confident that c.f. Stygian Owls are present in Montecristo
+NP, we believe that further documentation of the species would be
+beneficial for confirmation of the species in El Salvador.
+
+During 1979-1981, in the western portion of El Imposible NP, West (1988)
+heard or observed six owl species, including two that were not detected
+during this study (Barn Owls and Black-and-white Owls) in that NP. In
+1980, Barn Owls nested in a cave of Loma de Paja Mountain (opposite this
+study's El Imposible routes). An 5 April 2002, Barn Owls were detected
+in that same location (Vidal Campos Aguirre, personal communication).
+Vidal Campos Aguirre also stated that Black-and-white Owls were observed
+in two areas of the Mistepe River Valley in that same timeframe. It is
+likely, then, that these two owls still occur in El Imposible, though we
+were not able to detect them. We also observed a Barn Owl nesting near
+the N-1 route; however, again, we did not include any visual or auditory
+detections of owls outside of the survey stations. Barn Owls were
+detected in both Nancuchiname routes, but Black-and-white Owls were not
+detected in any routes during our surveys.
+
+Distinguishing individuals by non-invasive means, such as vocal traits,
+is preferable when species are rare, sensitive to handling, elusive, or
+when other techniques are unfeasible (Terry et al. 2005). Overall, the
+probability of detection was fairly low for all species (median values
+\(\le 0.50\), even though we used locally-recorded owl vocalizations
+when possible. In Guatemala, Gerhardt (1989) found that a series of
+Mottled Owl vocalizations obtained from a non-local ornithological
+laboratory could not generate a response from Guatemalan owls. We found
+that the probability of detecting most owls increased after we
+broadcasted the vocalizations of Mottled Owls, Pacific Screech-owls,
+Crested Owls, and Spectacled Owls. Conversely, owls were less likely to
+respond after we played Great Horned Owl calls. Because owls, like many
+birds of prey, are sensitive to interactions with other owls (Enrı́quez
+and Salazar 1997, Baumgardt et al. 2019) and may change their response
+to broadcast calls depending on their life stage (e.g., nesting versus
+courtship) (Flesch and Steidl 2007), we recommend that surveyors
+carefully consider the intra- and inter-species interactions that may
+affect detectability after playing specific bird vocalizations.
+
+Globally, an estimated 75\% of the nearly 250 species of owls are
+associated with dense and undisturbed forests (Mikkola 2012). Pérez-Léon
+et al. (2017) stated that within El Salvador, natural ecosystems had the
+most diverse owl populations, specifically habitats comprised of cloud,
+deciduous, riparian and pine-oak forests. In addition to areas that are
+preserved in a natural state, shade-grown coffee production may be
+providing valuable owl habitat. Coffee plantations account for around
+7\% of El Salvador's forested land due to a prevalence of shade-cover
+farming practices (Silva 2016). Silva (2016) also concluded that almost
+all primary or closed forest is surrounded by coffee farms, which act as
+a buffer zone against land use change. Likewise, Pérez-Léon et al.
+(2017) stated that coffee plantations are the most important
+agroecosystem for owl communities' survival in El Salvador.
+
+This research study was critical because of its long-term study of
+multiple owl species in an rapidly urbanizing and population-dense
+country with diverse habitats and relatively little previous research.
+As we continue to identify areas of the world that are owl research
+priorities based on the lack of current information leveraged against
+the potential conservation risk, such as in Buechley et al. (2019),
+studies such as ours will be important and hopefully increasing in
+frequency.
+
+\hypertarget{acknowledgements}{%
+\section{ACKNOWLEDGEMENTS}\label{acknowledgements}}
+
+Thank you to the Ministerio de Medio Ambiente y Recursos Naturales
+(MARN) for their authorization to conduct scientific research in El
+Salvador's protected natural areas (permits were obtained annually
+starting in 2002: AIMA MARN-DGPN 006-2002 and ending in 2013: MARN
+AIMA-005-2013). Thank you to: MARN and Salva Natura for their employee's
+assistance in San Salvador and in the field; to the MARN for logistics
+support and the GIS layers; to Dr.~David Anderson for his technical
+review; especially thanks to Alfonso Sermeño Martínez, Vidal Campos
+Aguirre, Charles Starner, our families, and the US Fish and Wildlife
+Service for their assistance throughout the years of this project.
+Author contributions: JNW conceived of, developed, and conducted the owl
+surveys; AAA conducted the statistical analyses. AAA wrote the
+manuscript with support from JNW.
+
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
+
+\begin{figure}
 \centering
-\includegraphics{../output/figures/ferpy_psi_byYr-1.png}
-\caption{1. Example figure caption, which can include in-line equations
-like \(y = \alpha + \beta x\) and references like ({\textbf{???}}).
-\label{figure1}}
+\includegraphics{../output/figures/map-1.png}
+\caption{1XX. Map of El Salvador. Owl surveys were conducted from 2003
+to 2013 in three different protected areas within El Salvador: El
+Imposible National Park (EINP), Montecristo National Park (MNP) and
+Nancuchiname Forest (NF). \label{map}}
 \end{figure}
 
-\section*{References}\label{references}
+\begin{figure}
+\centering
+\includegraphics{../output/figures/psi_means-1.png}
+\caption{2XX. Probability of occupancy (\(\psi_{t,h}\)) summarized by
+route for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and
+Spectacled Owl (SPEO). Owl surveys were conducted from 2003 to 2013 in
+three different protected areas within El Salvador: El Imposible
+National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2)
+and Nancuchiname Forest (N-1 and N-2). Posteriors were summarized with
+median \(\pm\) 90\% credible intervals. \label{psi_means}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../output/figures/psi_byYr-1.png}
+\caption{3XX. Probability of occupancy by route and year for Ferruginous
+Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO). Owl
+surveys were conducted from 2003 to 2013 in three different protected
+areas within El Salvador: El Imposible National Park (EI-1 and EI-2),
+Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and
+N-2). Posteriors were summarized with median \(\pm\) 90\% credible
+intervals. \label{figure2}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../output/figures/richness_byRtYr-1.png}
+\caption{4XX. Species richness (Richness\emph{) by route and year. Owl
+surveys were conducted from 2003 to 2013 in three different protected
+areas within El Salvador: El Imposible National Park (EI-1 and EI-2),
+Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and
+N-2). Posteriors of Richness} were summarized with median (black dot)
+\(\pm\) 90\% credible intervals (dark grey shading and vertical black
+lines). The number of owl species detected at each route and year was
+indicated with the light grey area. \label{figure5}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{../output/figures/p_detection-1.png}
+\caption{5XX. The probabiliity of detecting owls during spontaneous
+surveys (i.e., pre-broadcast) and after broadcast, depending on the
+species of owl used for broadcast vocalization. These logistic
+regression coefficients were evaluated from the richness model, which
+combined all detected species (All Species) and the single-species
+occupancy models for Ferruginous Pygmy-owls (FEPO), Mottled Owls (MOOW),
+and Spectacled Owls (SPEO). Owl surveys were conducted from 2003 to 2013
+in three different protected areas within El Salvador. Posteriors were
+summarized with median \(\pm\) 90\% credible intervals. The
+pre-broadcase effect size's credible intervals are represented with two
+grey horizontal lines on each graph \label{figure6}}
+\end{figure}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Baumgardt:2019}{}
+\leavevmode\hypertarget{ref-Andersen:2007}{}%
+Andersen, D. E. 2007. Survey techniques. Pages 89--100 \emph{in} D. M.
+Bird and K. L. Bildstein, editors. Raptor research and management
+techniques. Hancock House Publishers, Blaine, WA USA.
+
+\leavevmode\hypertarget{ref-Alvarez:2003}{}%
+Álvarez, J. M., and O. Komar. 2003. El parque nacional el imposible y su
+vida silvestre. Salva Natura y Ministerio de Medio Ambiente y Recursos
+Naturales.
+
+\leavevmode\hypertarget{ref-Baumgardt:2019}{}%
 Baumgardt, J. A., M. L. Morrison, L. A. Brennan, and T. A. Campbell.
@@
 
-\hypertarget{ref-Gelman:2008}{}
+\leavevmode\hypertarget{ref-Borges:2004}{}%
+Borges, S. H., L. M. Henriques, and A. Carvalhaes. 2004. Density and
+habitat use by owls in two Amazonian forest types. Journal of Field
+Ornithology 75:176--182.
+
+\leavevmode\hypertarget{ref-Broms:2016}{}%
+Broms, K. M., M. B. Hooten, and R. M. Fitzpatrick. 2016. Model selection
+and assessment for multi-species occupancy models. Ecology
+97:1759--1770.
+
+\leavevmode\hypertarget{ref-Buechley:2019}{}%
+Buechley, E. R., A. Santangeli, M. Girardello, M. H. C. Neate-Clegg, D.
+Oleyar, C. J. W. McClure, and Ç. H. Şekercioğlu. 2019. Global raptor
+research and conservation priorities: Tropical raptors fall prey to
+knowledge gaps. Diversity and Distributions 25:856--869.
+
+\leavevmode\hypertarget{ref-Campioni:2013}{}%
+Campioni, L., J. H. Sarasola, M. Santillán, and M. M. Reyes. 2013.
+Breeding season habitat selection by Ferruginous Pygmy Owls
+\emph{Glaucidium brasilianum} in central Argentina. Bird Study
+60:35--43.
+
+\leavevmode\hypertarget{ref-CIA:2020}{}%
+Central Intelligence Agency. 2020. The world factbook.
+www.cia.gov/the-world-factbook.
+
+\leavevmode\hypertarget{ref-Clark:1978}{}%
+Clark, R. J., D. G. Smith, and L. H. Kelso. 1978. Working bibliography
+of owls of the world. National Wildlife Federation Scientific and
+Technical Series.
+
+\leavevmode\hypertarget{ref-Corlett:2012}{}%
+Corlett, R. T. 2012. Climate change in the tropics: The end of the world
+as we know it? Biological Conservation 151:22--25.
+
+\leavevmode\hypertarget{ref-Dickey:1938}{}%
+Dickey, D. R., and A. J. van Rossem. 1938. The birds of El Salvador.
+Field Museum of Natural History.
+
+\leavevmode\hypertarget{ref-Enriquez:1997}{}%
+Enrı́quez, P. L., and J. L. R. Salazar. 1997. Intra-and interspecific
+calling in a tropical owl community. \emph{in} In: Duncan, james r.;
+johnson, david h.; nicholls, thomas h., eds. Biology and conservation of
+owls of the northern hemisphere: 2nd international symposium. Gen. Tech.
+Rep. NC-190. St. Paul, mn: US dept. Of agriculture, forest service,
+north central forest experiment station. 525-532.
+
+\leavevmode\hypertarget{ref-Enriquez:2001}{}%
+Enrı́quez, P., and J. Rangel-Salazar. 2001. Owl occurrence and calling
+behavior in a tropical rain forest. Journal of Raptor Research
+35:107--114.
+
+\leavevmode\hypertarget{ref-Enriquez:2006}{}%
+Enríquez, P. L., D. H. Johnson, and J. L. Rangel-Salazar. 2006.
+Taxonomy, distribution and conservation of owls in the Neotropics: A
+review. Pages 254--307 \emph{in} R. Rodríguez-Estrella, editor. Current
+Raptor Studies in México. Comisión Nacional para el Conocimiento y Uso
+de la Biodiversidad.
+
+\leavevmode\hypertarget{ref-Enriquez-Rocha:1993}{}%
+Enríquez-Rocha, P., J. L. Rangel-Salazar, and D. W. Holt. 1993. Presence
+and distribution of Mexican owls: a review. Journal of Raptor Research
+27:154--160.
+
+\leavevmode\hypertarget{ref-Esclarski:2014}{}%
+Esclarski, P., and R. Cintra. 2014. Effects of terra firm-forest
+structure on habitat use by owls (Aves: Strigiformes) in central
+Brazilian Amazonia. Ornitological Neotropical 25:433--458.
+
+\leavevmode\hypertarget{ref-Flesch:2007}{}%
+Flesch, A. D., and R. J. Steidl. 2007. Detectability and response rates
+of Ferruginous Pygmy-Owls. The Journal of Wildlife Management
+71:981--990.
+
+\leavevmode\hypertarget{ref-Fuller:1987}{}%
+Fuller, M. R., and J. A. Mosher. 1987. Raptor survey techniques.
+\emph{in} Raptor research techniques manual. US Fish and Wildlife
+Service.
+
+\leavevmode\hypertarget{ref-Gelman:2008}{}%
 Gelman, A., A. Jakulin, M. G. Pittau, and Y.-S. Su. 2008. A weakly
@@
 
-\hypertarget{ref-Plummer:2013}{}
+\leavevmode\hypertarget{ref-Gerhardt:1989}{}%
+Gerhardt, R. P. 1989. Response of the Mottled Owl (\emph{Ciccaba
+virgata}) to broadcast of conspecific call. \emph{in} W. A. Burnham, J.
+P. Jenny, and C. W. Turley, editors. Maya project: Use of raptors as
+environmental indices for design and management of protected areas and
+for building local capacity for conservation in latin america. The
+Peregrine Fund, Inc., World Center for Birds of Prey.
+
+\leavevmode\hypertarget{ref-Gonzalez:2017}{}%
+Gonzalez, R., R. Pineda, and L. Pineda. 2017. Photographic documentation
+of Fulvous Owl (\emph{Strix fulvescens} in Montecristo National Park, El
+Salvador. Spizaetus Neotropical Raptor Network Newsletter. 24:16--21.
+
+\leavevmode\hypertarget{ref-Guillera-Arroita:2019}{}%
+Guillera-Arroita, G., M. Kéry, and J. J. Lohoz-Monfort. 2019. Inferring
+species richness using multispecies occupancy modeling: Estimation
+performance and interpretation. Ecology and Evolution 9:780--792.
+
+\leavevmode\hypertarget{ref-Johnsgard:1988}{}%
+Johnsgard, P. A. 1988. North american owls: Biology and natural history.
+Yale University Press, New Haven, CT.
+
+\leavevmode\hypertarget{ref-Komar:2002}{}%
+Komar, O. 2002. Priority conservation areas for birds in El Salvador.
+Animal Conservation 5:173--183.
+
+\leavevmode\hypertarget{ref-Komar:2010}{}%
+Komar, O. 2010. Research stewardship strategy for Montecristo National
+Park, El Salvador. Technical report, USAID.
+
+\leavevmode\hypertarget{ref-Konig:1999}{}%
+König, C., and F. Weick. 1999. Owls: A guide to owls of the world. Yale
+University Press, New Haven, CT.
+
+\leavevmode\hypertarget{ref-Marin-Gomez:2017}{}%
+Marín-Gómez, O. H., Y. Toro-López, M. M. López-García, J. I. Garzón
+Zuluaga, and D. M. Santa-Aristizabal. 2017. First records of the
+Spectacled Owl (\emph{Pulsatrix perspicillata}) in urban areas, with
+notes on reproduction. North-Western Journal of Zoology 13:2.
+
+\leavevmode\hypertarget{ref-Mikkola:2012}{}%
+Mikkola, H. 2012. Owls of the world-a photographic guide. Firefly Books,
+Buffalo, New York, NY, USA.
+
+\leavevmode\hypertarget{ref-MARN:2010}{}%
+Ministerio de Medio Ambiente y Recursos Naturales. 2010. Guía de Aves
+Parque Nacional Montecristo. Guías de la Biodiversidad del Parque
+Nacíonal Montecristo. San Salvador, El Salvador.
+
+\leavevmode\hypertarget{ref-MARN:2015b}{}%
+Ministerio de Medio Ambiente y Recursos Naturales. 2015a. La guía de
+aves del Parque Nacíonal Montecristo. Parte de la serie de guías de la
+biodiversidad del Parque Nacíonal Montecristo. San Salvador, El
+Salvador.
+
+\leavevmode\hypertarget{ref-MARN:2015}{}%
+Ministerio de Medio Ambiente y Recursos Naturales. 2015b. Acuerdo No.~74
+Listado oficial de especíes de vida silvestre amenazadas o en peligro de
+extinción. San Salvador, El Salvador.
+
+\leavevmode\hypertarget{ref-Myers:2000}{}%
+Myers, N., and R. A. Mittermeier. 2000. Biodiversity hotspots for
+conservation priorities. Nature 403:853.
+
+\leavevmode\hypertarget{ref-Orihuela-Torres:2018}{}%
+Orihuela-Torres, A., L. Ordóñez-Delgado, A. Verdezoto-Celi, and J.
+Brito. 2018. Diet of Spectacled Owl (\emph{Pulsatrix perspicillata}) in
+Zapotillo, southwestern Ecuador. Revista Brasileira de Ornitologia
+26:52--56.
+
+\leavevmode\hypertarget{ref-Perez-Leon:2017}{}%
+Pérez-Léon, R., I. Vega, and N. Herrera. 2017. The owls of El Salvador.
+Pages 397--418 \emph{in} P. L. Enríquez, editor. Neotropical owls:
+Diversity and Conservation. Springer International Publishing AG, Cham,
+Switzerland.
+
+\leavevmode\hypertarget{ref-Plummer:2013}{}%
 Plummer, M. 2013. JAGS Version 3.4.0 user manual.
 
-\hypertarget{ref-R:2014}{}
+\leavevmode\hypertarget{ref-Rangel-Salazar:2017}{}%
+Rangel-Salazar, J. L., and P. L. Enríquez. 2017. Introduction: The birds
+in the neotropical region. Pages 1--6 \emph{in} Neotropical owls.
+Springer.
+
+\leavevmode\hypertarget{ref-R:2014}{}%
 R Development Core Team. 2014. R: A Language and Environment for
@@
 
+\leavevmode\hypertarget{ref-Royle:2007}{}%
+Royle, J. A. 2007. Analysis of multinomial models with unknown index
+using data augmentation. Journal of Computational and Graphical
+Statistics 16:67--85.
+
+\leavevmode\hypertarget{ref-Sarasola:2014}{}%
+Sarasola, J. H., and M. Á. Santillán. 2014. Spatial and temporal
+variation in the feeding ecology of ferruginous pygmy-owls
+(\emph{Glaucidium brasilianum}) in semiarid forests of central
+Argentina. Journal of Arid Environments 109:39--43.
+
+\leavevmode\hypertarget{ref-Silva:2016}{}%
+Silva, R. O. 2016. Country coffee profile: El Salvador. Technical
+report, International Coffee Organization.
+
+\leavevmode\hypertarget{ref-Takats:2001}{}%
+Takats, D. L., C. M. Francis, G. L. Holroyd, J. R. Duncan, K. M. Mazur,
+R. J. Cannings, W. Harris, and D. Holt. 2001. Guidelines for nocturnal
+owl monitoring in North America. Beaverhill Bird Observatory and Bird
+Studies Canada.
+
+\leavevmode\hypertarget{ref-Terry:2005}{}%
+Terry, A. M., T. M. Peake, and P. K. McGregor. 2005. The role of vocal
+individuality in conservation. Frontiers in Zoology 2:10.
+
+\leavevmode\hypertarget{ref-USFWS:1995}{}%
+United States Fish and Wildlife Service. 1995. Recovery plan for the
+Mexican Spotted Owl. Technical report, United States Department of the
+Interior.
+
+\leavevmode\hypertarget{ref-Vallely:2018}{}%
+Vallely, A. C., and D. Dyer. 2018. Birds Of Central America: BELIZE,
+GUATEMALA, HONDURAS, EL SALVADOR, NICARAGUA, COSTA RICA, and PANAMA.
+Princeton University Press, Princeton, NJ, USA.
+
+\leavevmode\hypertarget{ref-Wan:2018}{}%
+Wan, H. Y., J. L. Ganey, C. D. Vojta, and S. A. Cushman. 2018. Managing
+emerging threats to spotted owls. The Journal of Wildlife Management
+82:682--697.
+
+\leavevmode\hypertarget{ref-West:1988}{}%
+West, J. N. 1988. Raptors of El Imposible Forest, El Salvador. Masters
+thesis, Central Washington University, Ellensburg, WA.
+
+\leavevmode\hypertarget{ref-White:2013}{}%
+White, A. M., E. F. Zipkin, P. N. Manley, and M. D. Schlesinger. 2013.
+Conservation of avian diversity in the Sierra Nevada: Moving beyond a
+single-species management focus. PLoS ONE 8:e63088.
+
+\leavevmode\hypertarget{ref-Zepeda:1995}{}%
+Zepeda, L. 1995. Plan de manejo del área natural silvestre Nancuchiname.
+Anexo 4 - listado preliminar de las aves del área, San Salvador, El
+Salvador.
+
 \end{document}

### 2021-02-28 manuscript_working/ms_working_RMD.Rmd
updated organization, intro/discussion

@@
 As apex predators, owls are an important bioindicator of ecosystem health, though much of our understanding comes from temperate systems [@White:2013; @Wan:2018; @Buechley:2019].
-In the tropics, particularly in rain and cloud forests, owl studies are more difficult than in temperate regions, regardless of potentially denser populations [@Konig:1999], and neotropical owls are less well-studied than their temperate counterparts [Buechley:2019].
-Further, the neotropics were identified by @Buechley:2019 as a priority area for owl research, based on an extensive literature review that identified a relative lack of research attention and a high conservation risk for neotropical owls.
+The Neotropics were recently identified as a priority area for owl research because the high conservation risk for neotropical owls has not been supported by enough research attention [@Buechley:2019], but in tropical rain and cloud forests, owl studies are much more difficult than in temperate regions, regardless of potentially denser populations [@Konig:1999].
 As a result, we have limited understanding of neotropical owls’ distribution, ecological requirements, population dynamics, and reproductive behavior [@Clark:1978; @Enriquez:2006; @Perez-Leon:2017; @Rangel-Salazar:2017].
-This lack of information poses a challenge to proactive conservation, even as neotropical habitat experiences tremendous anthropogenic pressure and rapid, unpredictable responses to climate change [@Corlett:2012].
+This lack of information poses a challenge for proactive conservation, even as neotropical habitat experiences tremendous anthropogenic pressure and rapid, unpredictable responses to climate change [@Corlett:2012].
 In general, while our knowledge about the status and trends of neotropical owl populations is limited, we do know that owl abundance and distribution is decreasing in several regions where species have been added to endangered species lists or have become locally extirpated [@Enriquez:2006].
 
-Like many owl species, neotropical owls are difficult to accurately census due to their cryptic nature and nocturnal habits.
-Vocalizations of crepuscular and cryptic animals such as owls are commonly leveraged in survey methodology [@Konig:1999].
-Owl vocalizations can be diagnostic and are often considered more practically attainable than visual identification.
-Owls vocalize to communicate with the same species, to threaten or provoke other species, and to delimit territory [@Johnsgard:1988].
-
-Vocalization surveys for nocturnal birds of prey fall generally into two categories.
-Spontaneous or “passive” surveys involve listening for animal calls with no prior stimulus, whereas provocation or “broadcast” surveys are conducted by listening after a period of playing a recording of animal vocalizations [e.g., @Baumgardt:2019].
-The inclusion of broadcast calls in auditory surveys can increase the detection probability of owls, although intra- and inter-species interactions can play a role whether owls are more likely or less likely to call in response to a vocalization simulus [@Enriquez:1997; @Borges:2004; @Flesch:2007; @Baumgardt:2019].
+El Salvador is located on the western side of the Central American isthmus and is the smallest and most densely populated Central American country [Fig.
+1XXX; @CIA:2020].
+Overall, 13.6% of El Salvador's 20,720 km$^2$ of land is forested, and 82% is used for agricultural purposes [@CIA:2020].
+A common agroecosystem in El Salvador is shade-grown coffee, which supports forested cover on otherwise agricultural land.
+Approximately 7% of El Salvador's forested land is comprised of shade-grown coffee, which may be providing an important land-use buffer around El Salvador's protected natural areas [@Silva:2016; @Perez-Leon:2017].
+Although shade-grown coffee systems are potentially reducing the impacts of land-use and climate change on the forest-dependent owls of El Salvador [e.g., @Perez-Leon:2017], there is a lack of information about the long-term  dynamics of these populations.
+Concerns about habitat availability are not the only challenge facing El Salvador's owls.
+@Perez-Leon:2017 stated that there are human activities affecting owl populations in El Salvador, including illegal hunting, trapping, persecution, killing, and wildlife trade.
 
@@
 
-El Salvador is located on the western side of the Central American isthmus and is the smallest and most densely populated Central American country  [@CIA:2020].
-It is bordered by the Pacific Ocean, Guatemala and Honduras (Fig.
-1XXX).
-Overall, 13.6% of El Salvador's 20,720 km$^2$ of land is forest and 82% is used for agricultural purposes [@CIA:2020].
-We conducted surveys in three protected natural areas located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that represent a portion of the country's diverse ecosystems (Fig.
+We conducted surveys in three protected natural areas in El Salvador (El Imposible National Park, Montecristo National Park, and Nancuchiname Forest) located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that represent a portion of the country's diverse ecosystems (Fig.
 1xxx, Table 1XXX).
-
 El Imposible National Park (NP) was located 119 km southwest of San Salvador (Fig.
-1XXX, Table 1XXX).
+1XXX; Table 1XXX).
 Its elevation ranged from 250 to 1425 m above sea level (a.s.l.).
@@
 Nancuchiname Forest (797 ha) is a Protected Natural Area located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
-1XXX).
+1XXX; Table 1XXX).
 Its elevation ranged from sea level to 12 m a.s.l.
@@
 Route locations and habitats.
-Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
+Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2), and Nancuchiname Forest (N-1 and N-2).
 *
@@
 
-We conducted acoustic surveys on foot following the Guidelines for Nocturnaal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and other raptor research guidelines [@Fuller:1987].
+We conducted acoustic surveys on foot following the Guidelines for Nocturnal Owl Monitoring [@Takats:2001], a similar protocol developed for Costa Rica [@Enriquez:2001], and other raptor research guidelines [@Fuller:1987].
 This study's survey protocol was an extension of previous research completed in the western half of El Imposible NP [@West:1988].
@@
 No other owl species had enough detections to analyze species-specific occupancy (Table 2XX).
-Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled Owls than for Ferruginous Pygmy-owls or Spectacled Owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
+Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled Owls than for Ferruginous Pygmy-Owls or Spectacled Owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
 1X).
@@
 
-For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy-owls and Spectacled Owls stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy-owl in EI-1 in the early years of the surveys (Fig.
+For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy-Owls and Spectacled Owls stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy-Owl in EI-1 in the early years of the surveys (Fig.
 3XX).
-The probability of occupancy for Ferruginous Pygmy-owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
+The probability of occupancy for Ferruginous Pygmy-Owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
 3XX), and the probability of occupancy for the Spectacled Owl fluctuated slightly from year to year in the Nancuchaname routes (Fig.
@@
 4XXX).
-At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species (e.g., in M-2 during 2004 where we detected zero species (Fig.
+At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species (e.g., in M-2 during 2004 where we detected zero species; Fig.
 4XXX).
@@
 5XXX).
-When incorporating all owl species, broadcasting calls from Mottled Owls, Pacific Screech-owls, Crested Owls, Black-and-white Owls, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned Owl calls decreased detection (Fig.
+When incorporating all owl species, broadcasting calls from Mottled Owls, Pacific Screech-Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned Owl calls decreased detection (Fig.
 5XXX).
 
-The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled Owls, Ferruginous Pygmy-owls, and Spectacled Owls (Fig.
+The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled Owls, Ferruginous Pygmy-Owls, and Spectacled Owls (Fig.
 5XX).
-The probability of detecting Mottled Owls increased after broadcasting Mottled Owl, Pacific Screech-owl, Crested Owl, Black-and-white Owl, or Spectacled Owl calls.
-The probability of detecting Ferruginous Pygmy-owls increased after broadcasting Mottled Owl and Pacific Owl calls, and the probability of detecting Spectacled Owls increased after broadcasting Pacific Screech-owl, Black-and-white Owl, and Spectacled Owl calls (Fig.
+The probability of detecting Mottled Owls increased after broadcasting Mottled Owl, Pacific Screech-Owl, Crested Owl, Black-and-white Owl, or Spectacled Owl calls.
+The probability of detecting Ferruginous Pygmy-Owls increased after broadcasting Mottled Owl and Pacific Owl calls, and the probability of detecting Spectacled Owls increased after broadcasting Pacific Screech-Owl, Black-and-white Owl, and Spectacled Owl calls (Fig.
 5XX).
@@
 The goal of this project was to gather baseline data on owl abundance, distribution, and population trends in El Salvador.
-We established survey routes within three protected natural areas in disparate ecological regions (alluvial plain, forested, cloud forests) and attempted to compete two surveys of each route for each year from 2003 through 2013.
-Our goal of ten years’ worth of repeated surveys was based on similar studies that have previously been conducted on Mexican spotted owls (*Strix occidentalis*) [@USFWS:1995].
-However, our surveys were not always completed twice per year due to poor weather, issues with site access, or other complications.
-We did not survey during significant weather events, as rain and high winds diminish owl vocalizations and the ability for surveyors to detect owl calls [@Takats:2001; @Andersen:2007].
-Overall, our surveys yielded sufficient data to analyze occupancy patters for three species of interest; the Mottled Owl, Ferruginous Pygmy-owl, and Spectacled Owl.
+We established survey routes within three protected natural areas in disparate ecological regions (alluvial plain, forested, cloud forests) and attempted to complete two surveys passes per route each year from 2003 through 2013.
+Our surveys were not always completed twice per year in each route because of isses with site access or poor weather, which decreases owl vocalizations and detection [@Takats:2001; @Andersen:2007].
+Overall, our surveys yielded sufficient data to analyze occupancy patters for three species of interest; the Mottled Owl, Ferruginous Pygmy-Owl, and Spectacled Owl.
 
-The Mottled Owl is the most common and widespread owl in South America [@Vallely:2018] and they have been documented in a wide range of habitat types and elevations in Mexico [@Enriquez-Rocha:1993].
 Mottled Owls were the most frequently detected owl during our surveys.
+Mottled Owls are the most common and widespread owl in South America [@Vallely:2018], and they have been documented in a wide range of habitat types and elevations in Mexico [@Enriquez-Rocha:1993].
 A recent study indicated that Mottled Owl populations in El Salvador have persisted despite habitat loss that has restricted the distribution of several other owl species [@Perez-Leon:2017].
@@
 
-Ferruginous Pygmy-owls were the second most-often detected owls in this study.
+Ferruginous Pygmy-Owls were the second most-often detected owls in this study.
 This species demonstrated fairly low occupancy rates in El Imposible routes, other than in the first two years in EI-1.
-Ferruginous Pygmy-owls tend to prefer open habitats with a lot of edge, such as the flooded igap\'o forests of Amazonia [@Borges:2004] and heterogenous forests in central Argentina [@Campioni:2013], and their dietary composition is dependent on forest density [@Sarasola:2014].
+Ferruginous Pygmy-Owls tend to prefer open habitats with a lot of edge, such as the flooded $\text{igap\'o}$ forests of Amazonia [@Borges:2004] and heterogenous forests in central Argentina [@Campioni:2013].
+Further, the dietary composition of Ferruginous Pygmy-Owls varies based on forest density [@Sarasola:2014].
 Lower elevation areas of El Imposible were grazed and dotted with houses in the late 1980s [@West:1988], and this mixed agrarian landscape may have provided suitable habitat which has been reduced by subsequent reforestation efforts that began in 1997.
-It is possible that Ferruginous Pygmy-owls are responding to a decrease in edge habitat within El Salvador, and we recommend further surveys to determine the risk to Ferruginous Pygmy-owls in response to revegetation management.
+It is possible that Ferruginous Pygmy-Owls are responding to a decrease in edge habitat within El Salvador, and we recommend further surveys to determine the risk to Ferruginous Pygmy-Owls in response to revegetation management.
+Poaching may also be a risk for these owls: of eight species of owls confiscated from local markets in 1995 to 2008, the most frequent were Mottled Owls and Ferruginous Pygmy-Owls [@Perez-Leon:2017].
 
-Changes in habitat availability are not the only challenge facing El Salvador's owls.
-@Perez-Leon:2017 stated that there are human activities affecting owl populations in El Salvador, including illegal hunting, trapping, persecution, killing, and wildlife trade.
-They also stated that, of eight species of owls confiscated from local markets in 1995 to 2008, the most frequent were Mottled Owls and Ferruginous Pygmy-owls.
-If poaching of these species continues, their populations may be at more risk than our results indicated.
-
-Spectacled Owls followed similar occupancy patterns to that of the Ferruginous Pygmy-owls, with highest levels of occupancy in the riparian Nancuchiname Forest.
-Spectacled Owls can be found in a range of habitats including coffee plantations and urban parks [@Marin-Gomez:2017].
-Spectacled Owls tend to prefer lowland habitat, and they were reported to live in elevations below 900 m a.s.l.
-in Mexico [@Enriquez-Rocha:1993] and in lowland habitats in Ecuador below 1000 m a.s.l.
-[@Orihuela-Torres:2018]; however, they have been reportedly able to live up to 2600 m a.s.l.
-[@Marin-Gomez:2017].
-We detected Spectacled Owls in Montecristo National Park, transect M-2, which was on average 1755 m a.s.l., which is potentially near their preferred elevation range.
+Spectacled Owls followed similar occupancy patterns to that of the Ferruginous Pygmy-Owls, with highest levels of occupancy in the alluvial Nancuchiname Forest.
+Spectacled Owls can be found in a range of habitats including coffee plantations and urban parks [@Marin-Gomez:2017], and they tend to prefer lowland habitats below 1000 m a.s.l.
+[@Enriquez-Rocha:1993; @Orihuela-Torres:2018].
+Spectacled owls were detected in each of the routes located at low to moderate elevation; however, we also detected Spectacled Owls in Montecristo National Park, transect M-2 (1755 m a.s.l.), which is potentially near the maximum of their elevation tolerance range.
 Spectacled Owls have been reported to prefer mammalian prey over reptilian or arthropods [@Orihuela-Torres:2018] and preferentially hunt near fallen logs where small mammals often seek refuge [@Esclarski:2014], but we were unable to test for micro-habitat occupancy in our present study.
-Thus, further refinement over Spectacled Owl habitat selection could be explored in the future to better understand their ecology and distribution in the Neotropics and specifically El Salvador.
+Further refinement over Spectacled Owl habitat selection could be explored in the future to better understand their ecology and distribution in the Neotropics, and specifically, El Salvador.
 
@@
 Stygian Owl).
-A Nancuchiname Forest Management Plan only listed one owl, Ferruginous Pygmy-owl, as being present [@Zepeda:1995], but we confirmed the presence of four additional owl species in that NP.
 Fulvous Owl is one of El Salvador's 18 endemic bird species [@Komar:2002], and it was detected numerous times in the cloud forest of Montecristo NP (M-1).
-Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
+A Fulvous Owl has also been recently photographed in Montecristo NP [@Gonzalez:2017].
 
-Stygian Owl, which often inhabit dense cloud forests [@Enriquez-Rocha:1993], were listed as an owl species expected to reside in El Salvador [@Perez-Leon:2017), but has been previously undetected in the country.
+Stygian Owl, which often inhabit dense cloud forests [@Enriquez-Rocha:1993], were listed as an owl species expected to reside in El Salvador [@Perez-Leon:2017], but has been previously undetected in the country.
 On 26 March 2002, after listening to recordings and viewing photographs, the Montecristo NP guards stated that they thought c.f.
@@
 We did not record any of these vocalizations by c.f.
-Stygian Owls and detections that occurred between stations or away from survey routes were not included in our results.
+Stygian Owls, and detections that occurred between stations or away from survey routes were not included in our results.
 So while we are confident that c.f.
@@
 
-During 1979-1981, in the western portion of El Imposible NP, West (1988) heard or observed six owl species, including two that were not detected during this study (Barn Owls and Black-and-white Owls) in that NP.
-In 1980, Barn Owls nested in a cave of Loma de Paja Mountain (opposite this study's El Imposible routes).
-An 5 April 2002, Barn Owls were detected in that same location (Vidal Campos Aguirre, personal communication).
+During 1979-1981, in the western portion of El Imposible NP, West (1988) heard or observed six owl species, including two that were not detected during this study in El Imposible (Barn Owls and Black-and-white Owls).
+Barn Owls have been detected in a cave of Loma de Paja Mountain, which was opposite this study's El Imposible routes (Vidal Campos Aguirre, personal communication).
 Vidal Campos Aguirre also stated that Black-and-white Owls were observed in two areas of the Mistepe River Valley in that same timeframe.
-It is likely, then, that these two owls still occur in El Imposible, though we were not able to detect them.
-We also observed a Barn Owl nesting near the N-1 route; however, again, we did not include any visual or auditory detections of owls outside of the survey stations.
-Barn Owls were detected in both Nancuchiname routes, but Black-and-white Owls were not detected in any routes during our surveys.
+It is likely, then, that these two owls still occur in El Imposible, though we did not detect them.
 
+Like many owl species, neotropical owls are difficult to accurately census due to their cryptic nature and nocturnal habits.
+Owls vocalize to communicate with the same species, to threaten or provoke other species, and to delimit territory [@Johnsgard:1988].
 Distinguishing individuals by non-invasive means, such as vocal traits, is preferable when species are rare, sensitive to handling, elusive, or when other techniques are unfeasible [@Terry:2005].
-Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally-recorded owl vocalizations when possible.
+Thus, the vocalizations of crepuscular and cryptic animals such as owls are commonly leveraged in field surveys because they can be diagnostic and are often considered more practically attainable than visual identification [@Konig:1999].
+
+Vocalization surveys for nocturnal birds of prey fall generally into two categories.
+Spontaneous or “passive” surveys involve listening for animal calls with no prior stimulus, whereas provocation or “broadcast” surveys are conducted by listening after a period of playing a recording of animal vocalizations [e.g., @Baumgardt:2019].
+Our surveys coupled both approaches, with spontaneous surveys preceeding broadcast surveys.
+Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally-recorded owl vocalizations for our broadcast surveys when possible.
 In Guatemala, @Gerhardt:1989 found that a series of Mottled Owl vocalizations obtained from a non-local ornithological laboratory could not generate a response from Guatemalan owls.
-We found that the probability of detecting most owls increased after we broadcasted the vocalizations of Mottled Owls, Pacific Screech-owls, Crested Owls, and Spectacled Owls.
+We found that the probability of detecting most owls increased after we broadcast the vocalizations of Mottled Owls, Pacific Screech-Owls, Crested Owls, and Spectacled Owls.
 Conversely, owls were less likely to respond after we played Great Horned Owl calls.
-Because owls, like many birds of prey, are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019] and may change their response to broadcast calls depending on their life stage (e.g., nesting versus courtship) [@Flesch:2007], we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
+Owls are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019] and may change their response to broadcast calls depending on their life stage (e.g., nesting versus courtship) [@Flesch:2007], so we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
 
 Globally, an estimated 75% of the nearly 250 species of owls are associated with dense and undisturbed forests [@Mikkola:2012].
-@Perez-Leon:2017 stated that within El Salvador, natural ecosystems had the most diverse owl populations, specifically habitats comprised of cloud, deciduous, riparian and pine-oak forests.
-In addition to areas that are preserved in a natural state, shade-grown coffee production may be providing valuable owl habitat.
-Coffee plantations account for around 7% of El Salvador's forested land due to a prevalence of shade-cover farming practices  [@Silva:2016].
-@Silva:2016 also concluded that almost all primary or closed forest is surrounded by coffee farms, which act as a buffer zone against land use change.
-Likewise, @Perez-Leon:2017 stated that coffee plantations are the most important agroecosystem for owl communities’ survival in El Salvador.
-
-This research study was critical because of its long-term study of multiple owl species in an rapidly urbanizing and population-dense country with diverse habitats and relatively little previous research.
-As we continue to identify areas of the world that are owl research priorities based on the lack of current information leveraged against the potential conservation risk, such as in @Buechley:2019, studies such as ours will be important and hopefully increasing in frequency.
+@Perez-Leon:2017 stated that within El Salvador, natural ecosystems such as cloud, deciduous, riparian, and pine-oak forests had the most diverse owl populations, and our findings supported the importance of maintaining protected areas of El Salvador.
+We found that over the 10-year period of our study, owl species richness and single-species occupancy for three owls were all generally stable in the protected natural areas we investigated.
+Our study demonstrates the importance of maintaining natural forested areas and diverse habitat heterogeneity in a country under immense climate and population pressures.
 
@@
 ![2XX.
-Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
@@
 ![3XX.
-Probability of occupancy by route and year for Ferruginous Pygmy-owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Probability of occupancy by route and year for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
@@
 The probabiliity of detecting owls during spontaneous surveys (i.e., pre-broadcast) and after broadcast, depending on the species of owl used for broadcast vocalization.
-These logistic regression coefficients were evaluated from the richness model, which combined all detected species (All Species) and the single-species occupancy models for Ferruginous Pygmy-owls (FEPO), Mottled Owls (MOOW), and Spectacled Owls (SPEO).
+These logistic regression coefficients were evaluated from the richness model, which combined all detected species (All Species) and the single-species occupancy models for Ferruginous Pygmy-Owls (FEPO), Mottled Owls (MOOW), and Spectacled Owls (SPEO).
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2021-02-28 manuscript_working/ms_working_RMD.Rmd
updated figure numbers, XX removed

@@
 El Salvador is located on the western side of the Central American isthmus and is the smallest and most densely populated Central American country [Fig.
-1XXX; @CIA:2020].
+1; @CIA:2020].
 Overall, 13.6% of El Salvador's 20,720 km$^2$ of land is forested, and 82% is used for agricultural purposes [@CIA:2020].
@@
 We conducted surveys in three protected natural areas in El Salvador (El Imposible National Park, Montecristo National Park, and Nancuchiname Forest) located at different elevations (high, mid and low elevations) and in different types of forest vegetation (alluvial, deciduous, semi-deciduous, pine-oak and cloud forest) that represent a portion of the country's diverse ecosystems (Fig.
-1xxx, Table 1XXX).
+1, Table 1).
 El Imposible National Park (NP) was located 119 km southwest of San Salvador (Fig.
-1XXX; Table 1XXX).
+1; Table 1).
 Its elevation ranged from 250 to 1425 m above sea level (a.s.l.).
@@
 Montecristo NP (1973 ha) was located 125 km northwest of San Salvador (Fig.
-1XXX; Table 1XXX).
+1; Table 1).
 Its altitude ranged from 730 to 2418 m a.s.l.
@@
 Nancuchiname Forest (797 ha) is a Protected Natural Area located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
-1XXX; Table 1XXX).
+1; Table 1).
 Its elevation ranged from sea level to 12 m a.s.l.
@@
 
-*Table 1XX.
+*Table 1.
 Route locations and habitats.
@@
 We created six 2-km long survey routes (transects) that each included 10 permanent survey stations set at 200 m intervals.
-Two routes were established in each protected area (Table 1xxx).
+Two routes were established in each protected area (Table 1).
 
@@
 
-*Table 2XX. Species detection records by route. Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route.*
+*Table 2. Species detection records by route. Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador, although specific survey years varied by route.*
 
@@
 To model species richness at each route in each year, we assumed that there were 14 possible species present in El Salvador.
-This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 2XX).
+This included the 9 species observed during our surveys and an additional 5 species that we never observed (Table 2).
 This upper limit to expected species richness related directly to the 13 species known to inhabit El Salvador [@Dickey:1938] plus one species we identified (c.f.
@@
 We used the R2jags package in R [@R:2014] [@Plummer:2013] to implement the Bayesian models.
-We verified that R-hats were lower than 1.3 and visually inspected traceplots to verify that chains mixed well (Supplemental information S1-3xx).
+We verified that R-hats were lower than 1.3 and visually inspected traceplots to verify that chains mixed well (Supplemental information S1xx).
 
-We applied the single-species occupancy model for three owl species that had sufficient positive detections for analysis: Mottled Owl, Spectacled Owl, and Ferruginous Pygmy-Owl (Table 2XX).
+We applied the single-species occupancy model for three owl species that had sufficient positive detections for analysis: Mottled Owl, Spectacled Owl, and Ferruginous Pygmy-Owl (Table 2).
 Based on our understanding of owl ecology, we assumed that Spectacled and Ferruginous Pygmy owls would not occupy route M-1 in Montecristo, so we removed M-1 from those species' occupancy analyses.
@@
 We detected nine species of owls during our surveys: Barn Owl, Whiskered Screech-Owl, Pacific Screech-Owl, Spectacled Owl, Great Horned Owl, Ferruginous Pygmy-Owl, Mottled Owl, Fulvous Owl, and c.f.
-Stygian Owl (Table 2xXX).
+Stygian Owl (Table 2).
 We identified Whiskered Screech-Owls in Montecristo NP, which had previously been undocumented in that protected area.
@@
 
-Mottled Owls, Spectacled Owls, and Ferruginous Pygmy-Owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 2XX).
-No other owl species had enough detections to analyze species-specific occupancy (Table 2XX).
+Mottled Owls, Spectacled Owls, and Ferruginous Pygmy-Owls had 542, 137, and 187 positive detections over the 11 year period, respectively (Table 2).
+No other owl species had enough detections to analyze species-specific occupancy.
 Averaged across years, we found that the probability of occupancy was somewhat higher for Mottled Owls than for Ferruginous Pygmy-Owls or Spectacled Owls in both routes of El Imposible and in M-2 route of Montecristo (Fig.
-1X).
+1).
 Occupancy probabilities were relatively high across all three species for both routes of Nancuchiname (Fig.
-2XX).
+2).
 We found that the probability of occupancy for Mottled Owls was relatively consistent across time in both El Imposible routes and in N-2 (Fig.
-3XX), but that it seemed to increase in the last approximately 6 years in M-2 and N-1 (Fig.
-3XX).
+3), but that it seemed to increase in the last approximately 6 years in M-2 and N-1 (Fig.
+3).
 The probability of occupancy in M-1 seemed to have decreased through time for Mottled Owls (Fig.
-3XX).
+3).
 
 For routes EI-1, EI-2, and M-2, the probability of occupancy for Ferruginous Pygmy-Owls and Spectacled Owls stayed relatively low across time other than a higher probability of occupancy for Ferruginous Pygmy-Owl in EI-1 in the early years of the surveys (Fig.
-3XX).
+3).
 The probability of occupancy for Ferruginous Pygmy-Owls was relatively high in both routes of Nancuchaname except in 2008 and 2009 (Fig.
-3XX), and the probability of occupancy for the Spectacled Owl fluctuated slightly from year to year in the Nancuchaname routes (Fig.
-3XX).
+3), and the probability of occupancy for the Spectacled Owl fluctuated slightly from year to year in the Nancuchaname routes (Fig.
+3).
 
@@
 Our estimates of each route's community richness ($\text{Richness}_{h}$) and 90% credible intervals were 4 (4,4), 3 (3,3), 3 (3,4), 4 (4,5), 5 (5,5), and 5 (5,5) for EI-1, EI-2, M-1, M-2, N-1, and N-2, respectively.
-These median values were equivalent to the number of species detected in each route (Table 2XX), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
+These median values were equivalent to the number of species detected in each route (Table 2), indicating that the probability of occupancy for undetected species was low and no additional species were predicted to be present in a route's community, other than those that were detected.
 
 When analyzed by year and route, median species richness ($\text{Richness}^*_{t,h}$) varied from 0 to 5 (Fig.
-4XXX).
+4).
 At most, the median species richness was one species higher than the number of species detected at any single route and year; however, the 90% credible intervals of richness added up to 4 new undetected species (e.g., in M-2 during 2004 where we detected zero species; Fig.
-4XXX).
+4).
 
@@
 Including all detected species in one richness model increased the precision of the estimates of the effect of broadcast calls on detection probability (i.e., compare All Species with individual occupancy results in Fig.
-5XXX).
+5).
 When incorporating all owl species, broadcasting calls from Mottled Owls, Pacific Screech-Owls, Crested Owls, Black-and-white Owls, and Spectacled Owls increased detection probability; whereas broadcasting the Great Horned Owl calls decreased detection (Fig.
-5XXX).
+5).
 
 The probability of detection when analyzed with single-species occupancy models varied little during the timeframe before broadcast recordings were played (Pre-broadcast); however, the probability of detecting owls after broadcast calls were played varied by call and between Mottled Owls, Ferruginous Pygmy-Owls, and Spectacled Owls (Fig.
-5XX).
+5).
 The probability of detecting Mottled Owls increased after broadcasting Mottled Owl, Pacific Screech-Owl, Crested Owl, Black-and-white Owl, or Spectacled Owl calls.
 The probability of detecting Ferruginous Pygmy-Owls increased after broadcasting Mottled Owl and Pacific Owl calls, and the probability of detecting Spectacled Owls increased after broadcasting Pacific Screech-Owl, Black-and-white Owl, and Spectacled Owl calls (Fig.
-5XX).
+5).
 
@@
 
-![1XX.
+![1.
 Map of El Salvador.
@@
 
-![2XX.
+![2.
 Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
@@
 
-![3XX.
+![3.
 Probability of occupancy by route and year for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
@@
 
-![4XX.
+![4.
 Species richness (Richness*) by route and year.
@@
 
-![5XX.
+![5.
 The probabiliity of detecting owls during spontaneous surveys (i.e., pre-broadcast) and after broadcast, depending on the species of owl used for broadcast vocalization.

### 2021-02-28 manuscript_working/ms_working_RMD.Rmd
removed abbreviations from figures (eg MOOW)

@@
 ![2.
-Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Probability of occupancy ($\psi_{t,h}$) summarized by route for Ferruginous Pygmy-Owl, Mottled Owl, and Spectacled Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
@@
 ![3.
-Probability of occupancy by route and year for Ferruginous Pygmy-Owl (FEPO), Mottled Owl (MOOW), and Spectacled Owl (SPEO).
+Probability of occupancy by route and year for Ferruginous Pygmy-Owl, Mottled Owl, and Spectacled Owl.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
@@
 ![4.
-Species richness (Richness*) by route and year.
+Species richness (Richness$^*$) by route and year.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador: El Imposible National Park (EI-1 and EI-2), Montecristo National Park (M-1 and M-2) and Nancuchiname Forest (N-1 and N-2).
-Posteriors of Richness* were summarized with median (black dot) $\pm$ 90% credible intervals (dark grey shading and vertical black lines).
+Posteriors of Richness$^*$ were summarized with median (black dot) $\pm$ 90% credible intervals (dark grey shading and vertical black lines).
 The number of owl species detected at each route and year was indicated with the light grey area.
@@
 The probabiliity of detecting owls during spontaneous surveys (i.e., pre-broadcast) and after broadcast, depending on the species of owl used for broadcast vocalization.
-These logistic regression coefficients were evaluated from the richness model, which combined all detected species (All Species) and the single-species occupancy models for Ferruginous Pygmy-Owls (FEPO), Mottled Owls (MOOW), and Spectacled Owls (SPEO).
+These logistic regression coefficients were evaluated from the richness model, which combined all detected species (All Species) and the single-species occupancy models for Ferruginous Pygmy-Owls, Mottled Owls, and Spectacled Owls.
 Owl surveys were conducted from 2003 to 2013 in three different protected areas within El Salvador.

### 2021-02-28 programs/g_graphing_results_global.R

@@ -20,6 +20,7 @@ library(plyr)
 library(rgdal)
 library(raster)
 library(mapproj)
+library(stringr)
 
 #' Clear environment and set seed
 #' 
@@ -74,8 +75,8 @@ psi.means.ferpy$Species[max(nrow(psi.means.ferpy))] <- "FerPy"
 #'
 #' Merge data
 psi.means <- rbind(psi.means.ferpy, psi.means.specd, psi.means.mottd)
-psi.means$code <- ifelse(psi.means$Species == "FerPy", "FEPO", 
-                         ifelse(psi.means$Species == "Mottd", "MOOW", "SPEO"))
+psi.means$code <- ifelse(psi.means$Species == "FerPy", "Ferruginous Pygmy-Owl", 
+                         ifelse(psi.means$Species == "Mottd", "Mottled Owl", "Spectacled Owl"))
 
 #' Last minute fix to route names
 #' 
@@ -89,8 +90,8 @@ psi.means$Route <- ifelse(psi.means$Route == "EI1", "EI-1",
 
 #'
 #+ psi_means, fig.width = 2.8346, dpi = 600, fig.height = 4.5
-ggplot(data = psi.means, aes(x = Route, y = Psi.median, group = code))+
-  geom_bar(stat = "identity", aes(fill = code), position= position_dodge())+
+ggplot(data = psi.means, aes(x = Route, y = Psi.median, group = str_wrap(code,10)))+
+  geom_bar(stat = "identity", aes(fill = str_wrap(code,10)), position= position_dodge())+
   geom_linerange(aes(ymin = Psi.LL05, ymax = Psi.UL95), position = position_dodge(width = 0.9))+
   facet_wrap(~Route, nrow = 3,scales = "free_x")+
   ylim(c(0,1))+
@@ -123,8 +124,8 @@ psi.post.all$Year[psi.post.all$Psi.mean == 0.00001] <- 2000
 psi.post.all$Species[psi.post.all$Psi.mean == 0.00001] <- "Specd"
 psi.post.all$Route[psi.post.all$Psi.mean == 0.00001] <- "M.1"
 
-psi.post.all$code <- ifelse(psi.post.all$Species == "FerPy", "FEPO", 
-                            ifelse(psi.post.all$Species == "Mottd", "MOOW", "SPEO"))
[…]

### 2021-02-28 manuscript_working/ms_working_RMD.Rmd
Last RMD version, switching to word

@@
 
-Nancuchiname Forest (797 ha) is a Protected Natural Area located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
+Nancuchiname Forest (797 ha) was a Protected Natural Area located 95 km southeast of San Salvador in the coastal alluvial plain along the Lempa River (Fig.
 1; Table 1).
@@
 trichopsis*), Mottled Owls, Fulvous Owls (*Strix fulvescens*), Stygian Owls (*Asio stygius*), Great Horned Owls (*Bubo virginianus*), Whiskered Owls, Mottled Owls, Fulvous Owls, Stygian Owls, and Great Horned Owls at stations 1 through 10, respectively, in both routes of Montecristo NP (M-1 and M-2).
+We used locally-recorded owl vocalizations for our broadcast surveys when possible, because @Gerhardt:1989 found that a series of Mottled Owl vocalizations obtained from a non-local ornithological laboratory could not generate a response from Guatemalan owls.
 
@@
 Distinguishing individuals by non-invasive means, such as vocal traits, is preferable when species are rare, sensitive to handling, elusive, or when other techniques are unfeasible [@Terry:2005].
-Thus, the vocalizations of crepuscular and cryptic animals such as owls are commonly leveraged in field surveys because they can be diagnostic and are often considered more practically attainable than visual identification [@Konig:1999].
-
-Vocalization surveys for nocturnal birds of prey fall generally into two categories.
-Spontaneous or “passive” surveys involve listening for animal calls with no prior stimulus, whereas provocation or “broadcast” surveys are conducted by listening after a period of playing a recording of animal vocalizations [e.g., @Baumgardt:2019].
-Our surveys coupled both approaches, with spontaneous surveys preceeding broadcast surveys.
-Overall, the probability of detection was fairly low for all species (median values $\le 0.50$, even though we used locally-recorded owl vocalizations for our broadcast surveys when possible.
-In Guatemala, @Gerhardt:1989 found that a series of Mottled Owl vocalizations obtained from a non-local ornithological laboratory could not generate a response from Guatemalan owls.
-We found that the probability of detecting most owls increased after we broadcast the vocalizations of Mottled Owls, Pacific Screech-Owls, Crested Owls, and Spectacled Owls.
-Conversely, owls were less likely to respond after we played Great Horned Owl calls.
+Our surveys involved periods of passive listening followed by a broadcast call designed to trigger owl vocalization.
+Overall, the probability of detection was fairly low for all species (median values $\le 0.50$) and was generally lower during the passive listening periods pre-broadcast.
 Owls are sensitive to interactions with other owls [@Enriquez:1997; @Baumgardt:2019] and may change their response to broadcast calls depending on their life stage (e.g., nesting versus courtship) [@Flesch:2007], so we recommend that surveyors carefully consider the intra- and inter-species interactions that may affect detectability after playing specific bird vocalizations.
@@
 @Perez-Leon:2017 stated that within El Salvador, natural ecosystems such as cloud, deciduous, riparian, and pine-oak forests had the most diverse owl populations, and our findings supported the importance of maintaining protected areas of El Salvador.
-We found that over the 10-year period of our study, owl species richness and single-species occupancy for three owls were all generally stable in the protected natural areas we investigated.
-Our study demonstrates the importance of maintaining natural forested areas and diverse habitat heterogeneity in a country under immense climate and population pressures.
+We found that over the 10-year period of our study, overall owl species richness and single-species occupancy for three owls were generally stable in the protected natural areas we investigated.
+We also found that the probability of detecting most owls increased after we broadcast the vocalizations of Mottled Owls, Pacific Screech-Owls, Crested Owls, and Spectacled Owls.
+Conversely, owls were less likely to respond after we played Great Horned Owl calls.
+Our study demonstrates the importance of maintaining natural forested areas and diverse habitat heterogeneity in a country under considerable climate and population pressures.
 
